### 2022-10-27 manuscript/manuscript.Rmd
Add first version of the manuscript.

---
title: Transition from forest to agriculture in the Brazilian Amazon from 1985 to 2021
journal: essd
author:
  - given_name: Hugo
    surname: Tameirão Seixas
    affiliation: "1, 2"
    email: seixas.hugo@protonmail.com
    corresponding: true
  - given_name: Hilton
    surname: Luis Ferraz da Silveira
    affiliation: 2
  - given_name: Alan 
    surname: Falcão
    affiliation: 2
  - given_name: Fabiana
    surname: Da Silva Soares
    affiliation: 2
  - given_name: Ramon
    surname: Bicudo
    affiliation: 1

affiliation:
  - code: 1
    address: Center for Environmental Studies and Research, University of Campinas, Rua dos Flamboyants 155, Brazil
  - code: 2
    address: Embrapa Territorial, Av. Soldado Passarinho 303, Brazil

abstract: |
  A new dataset was created by calculating the time necessary for deforested forests to transition to agriculture in the Brazillian Amazon biome. The new data can be useful in interdiciplinary studies about land cover and land use change in brazil, its drivers and implications. A main inovation is that the dataset links the deforestation year with the year of agriculture establishment, which can provide new information about this process.

bibliography: references.bib

running:
  title: Transition from forest to agriculture in the Brazilian Amazon from 1985 to 2021
  author: Seixas et al.

competinginterests: |
  The authors declare no competing interests.
# See https://publications.copernicus.org/for_authors/licence_and_copyright.html, normally used for transferring the copyright, if needed. 
# Note: additional copyright statements for affiliated software or data need to be placed in the data availability section.

copyrightstatement: |
  The author's copyright for this publication is transferred to institution/company. 
### The following commands are for the statements about the availability of data sets and/or software code corresponding to the manuscript.
### It is strongly recommended to make use of these sections in case data sets and/or software code have been part of your research the article is based on.
### Note: unless stated otherwise, software and data affiliated with the manuscript are assumed to be published under the same licence as the article (currently Creative Commons 4.0)

authorcontribution: |

acknowledgements: |

appendix:

output:
  bookdown::pdf_book:
    base_format: rticles::copernicus_article # for using bookdown features like \@ref()
---

\introduction[Introduction]

In the last decades, the Amazon biome have been submitted to strong changes, the advance of agricultural areas over native forests is shaping a new landscape in the region.
The most common pattern of transitions from natural forest formations to agriculture sites, is the initial process of deforestation, followed by the establishment of pasture, where extensive livestock production takes place, until its replacement to agricultural activities.
This process can take several decades to be accomplished, or even less than one year, in which it may be considered as a direct transition from forest to agriculture.

Land cover transitions can cause important impacts on ecosystem properties [@Nunes2022].

This project aims to characterize and quantify the length of the transitions from forest formations to agriculture in the Amazon.

# Methods

The estimations of transition from forest to agriculture were performed for the Amazon biome region, as defined by the Brazilian Institute of Geography and Statistics (IBGE), in 2019.
Transitions were calculated using land use and land cover classification data from MapBiomas [@Souza2020], which ranges from 1985 to 2021, at a spatial resolution of approximately 30 meters.

Transition length was calculated pixel by pixel, by performing the following steps:

1.  Load raster and extract valid values into a table;

2.  Calculate the year of first occurrence of a non-natural LULC class for each pixel;

3.  Calculate the first year of "Forest Formation" LULC class after the year calculated in step 2, for each pixel;

4.  Classify rows as "before" or "after" the occurrence of the year calculated in step 3, for each pixel;

5.  Calculate the last year of "Forest Formation" within the rows classified as "before", for each pixel;

6.  Calculate the first year of any agriculture type class within the rows classified as "before", for each pixel;

7.  Calculate the difference between years from items 5 and 6 to get the LULC transition length in years, for each pixel;

# Description of data collection

We chose two ways to store and share the data.

-   One is to store the data from all three tables in three sets of Parquet files, in a structure of folders and sub folders;

-   The other is to convert the "trans_length.parquet" tables back to raster files.
    Raster files are useful to efficiently store spatial data, calculate areas, operate in relation to vector objects and perform data summaries, and it is also a very accessible format;

The new raster files are created based on the tiles downloaded from GEE (which are in the path "./data/raw_raster_tiles/"), using their extent values and its number of rows and columns (located on the "./data/trans_tabular_dataset/tiles_metadata.parquet" file).
For each tile file from "./data/raw_raster_tiles/", a set of new raster files are created based in the number of agriculture cycle ("agri_cycle") found in the tables, resulting in one raster for each agriculture cycle.
Each new raster file contains 6 bands:

-   Band 1 = "agri_code" (The agriculture class code that resulted from the transition);

-   Band 2 = "forest_year" (The last forest year before the transition, can be interpreted as the deforestation year);

-   Band 3 = "agri_year" (The first year of the agriculture establishment, being the end of the transition);

-   Band 4 = "trans_length" (The length in years of the transition from forest to agriculture);

-   Band 5 = "trans_cycle" (The number of the transition cycle in which the transition occurred);

-   Band 6 = "forest_type" (The type of the forest that suffered the transition);

The new raster files are saved in the path "./data/trans_raster_tiles/".

To facilitate the use of these new raster files, we also merged all the raster of each agriculture cycle into a single mosaic, which are saved in the path "./data/trans_raster_mosaic/".

# Results and discussion

The transition calculations shows that between 1985 and 2021, 64874 squared kilometers of forests were converted to agriculture, in the Brazilian Amazon biome.
The length of the transitions can go from 1 to 36 years, in which transitions closer to 1 year are considered as fast transitions, and transitions closer to 36 years are considered as slow transitions.

## Transition patterns

Transitions from forest to agriculture can be found in almost every region in the Amazon, but is mostly concentrated in clusters, specially in the south and east of the biome, in the states of Mato Grosso, Pará and Maranhão (Figure \@ref(fig:map-plot)).

```{r map-plot, out.width = "17cm", fig.pos = "h", echo = FALSE, fig.cap = "Map of distribution of transitions from forests to agriculture in the Brazillian Amazon biome. The hexagonal cells represent the most common transition length, and do not reflect the amount of area of transitions inside a cell. Transitions are concentraded in the south (Mato Grosso state), and in the east (Pará and Maranhão states). The transition length ranges from 1 (blue tones) to 36 years (pink tones), and clusters of fast transitions (transitions closer to 1 year) can be discerned from clusters of slow transitions (transitions closer to 36 years). The histogram located in the bottom left shows that fast transitions are more common than slower transitions. The zoomed map in the bottom right shows the results in finer resolution, where it is possible to observe different transition lengths between properties."}

knitr::include_graphics("figs/map.png")

```

Well defined clusters can be observed in the map created with aggregated transition length data (Figure \@ref(fig:map-plot)), slow transition areas tend to concentrate in specific regions in the Amazon biome, while fast transitions seem to have a wider distributions, but also tend to form clusters.
However, this pattern does not hold completely when observing the data at its original scale (Figure \@ref(fig:map-plot)), where areas with different transition lengths are mixed between each other.
When observing at the original scale, we could not spot any well defined pattern or direction of the occurrence of faster to slower transitions (Figure \@ref(fig:map-plot)).

Other studies investigating patterns of transitions in the Amazon also found the formation of clusters of patterns, although there is a big heterogeneity at larger scales [@MullerHansen2017].

The data can be analysed year by year, and also be separated by primary and secondary forests being converted to agriculture (Figure \@ref(fig:transbar-plot)).

```{r transbar-plot, out.width = "17cm", fig.pos = "h", echo = FALSE, fig.cap = "Transition area per year and transition length. The bars represent the total amount of area at some state of the transition for each year. The color gradient in each bar represents the transition length related to a deforestation or a agriculture establishment event. Blue tones represents fast transitions (transitions closer to one year), pink tones represents slow transitions (transitions closer to 36 years). Transition events were separated by deforestation of primary forests (a) and secondary forests (c), and the subsequent  agriculture establishment of primary forests (b) and secondary forests (d). TODO: add labels for important years hwere public policies took place."}

knitr::include_graphics("figs/trans_length_cols.png")

```

The deforestation area of primary forests increased largely from 1985 to 2002, which was followed be a steep decrease until 2007, when the deforested areas being converted to agriculture reached a stable rate.
From 1985 to approximately 1995, most of deforested areas suffered a slow transition, mostly were higher than 10 years, after this period, fast transitions started to become more common, specially from 2001 to 2003.

Agriculture establishment over areas of primary forests peaked in 2005 and 2013.
Despite similar rates between both years, their transition lengths differ greatly, in 2005 most of the transitions were faster than 10 years, while in 2013 the great majority of transitions were slower than 10 years.
The year of 2003 marked a change in the transition length of establishment of agriculture areas, after this year, most of the transitions happened in areas deforested at least 10 years before.
Even after the decrease of deforestation after 2002, agriculture areas are expanding over lands where deforestation happened before 2002.
However, after 2018, a sudden drop of agriculture establishment rate happened.

In 2006, the Brazilian Association of Vegetable Oil Industries (BIOVE) and the National Association of Cereal Exporters (ANEC) committed to avoid commercialization of soy grains harvested from areas deforested after 2006.
Our estimates of transitions show a decrease of deforested areas to be converted to agriculture after 2007, where it reached minimal values (Figure \@ref(fig:transbar-plot).a).
After 2006, agriculture establishment over deforested primary forests suffered a decrease, which stayed relatively stable until 2012, where a steep increase occurred, however, the new areas being occupied by agriculture were mainly over areas that were cleared more than a decade before (therefore, before 2006) (Figure \@ref(fig:transbar-plot).b) This shows that agriculture expansion did not halt after the soy moratorium, producers started expanding in old cleared areas.
Expansion of agriculture areas also expanded over cleared areas of secondary forests in 2012, but with an important amount of fast transitions.
The causes of the increase of agriculture establishment areas can be numerous, one main driver was the approval of a new Forest Code, in 2012, which is considered to have undermined the environmental protection [@Kroger2017; @Pereira2019].

The transition length patterns across years can change significantly between different states (Figure \@ref(fig:transridge-plot)).

```{r transridge-plot, out.width = "17cm", fig.pos = "h", echo = FALSE, fig.cap = "Transition length patterns inside states that belongs to the Amazon biome."}

knitr::include_graphics("figs/trans_ridge.png")

```

The state of Amapá presented fast transitions along all the time series, where slow transitions are not as common.
In contrast, Acre shows a majority of slow transitions, in which only 2021 showed more fast transitions.There are three states where the pattern of transitions length across time are alike, Mato Grosso, Pará and Rondônia presented more fast transitions from 1995 to 2005, after this period, slow transitions became more common with time, until around 2015, where the occurrence of fast transitions were the majority again.

## Validation

When analyzing the errors from the transition length estimates, we observe that the year of deforestation shows the least amount of errors (Figure \@ref(fig:errorbar-plot)).
The MAE of the deforestation year is of 1.42 years, and the error shows a bias towards underestimation.
The year of the agriculture establishment and the transition length estimates showed larger errors when comparing with visual inspection.

```{r errorbar-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = "TODO: Add title."}

knitr::include_graphics("figs/error_bars.png")

```

```{r errorscatter-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = "TODO: Change line showing MAE to shaded areas showing quantiles of the errors. Add title."}

knitr::include_graphics("figs/error_scatter.png")

```

## Qualitative assessment

\conclusions[Conclusions]

### 2022-10-27 manuscript/manuscript.tex
Add first version of the manuscript.

%% Copernicus Publications Manuscript Preparation Template for LaTeX Submissions
%% ---------------------------------
%% This template should be used for copernicus.cls
%% The class file and some style files are bundled in the Copernicus Latex Package, which can be downloaded from the different journal webpages.
%% For further assistance please contact Copernicus Publications at: production@copernicus.org
%% https://publications.copernicus.org/for_authors/manuscript_preparation.html

%% copernicus_rticles_template (flag for rticles template detection - do not remove!)

%% Please use the following documentclass and journal abbreviations for discussion papers and final revised papers.

%% 2-column papers and discussion papers
\documentclass[essd, manuscript]{copernicus}

%% Journal abbreviations (please use the same for preprints and final revised papers)

% Advances in Geosciences (adgeo)
% Advances in Radio Science (ars)
% Advances in Science and Research (asr)
% Advances in Statistical Climatology, Meteorology and Oceanography (ascmo)
% Annales Geophysicae (angeo)
% Archives Animal Breeding (aab)
% Atmospheric Chemistry and Physics (acp)
% Atmospheric Measurement Techniques (amt)
% Biogeosciences (bg)
% Climate of the Past (cp)
% DEUQUA Special Publications (deuquasp)
% Drinking Water Engineering and Science (dwes)
% Earth Surface Dynamics (esurf)
% Earth System Dynamics (esd)
% Earth System Science Data (essd)
% E&G Quaternary Science Journal (egqsj)
% EGUsphere (egusphere) | This is only for EGUsphere preprints submitted without relation to an EGU journal.
% European Journal of Mineralogy (ejm)
% Fossil Record (fr)
% Geochronology (gchron)
% Geographica Helvetica (gh)
% Geoscience Communication (gc)
% Geoscientific Instrumentation, Methods and Data Systems (gi)
% Geoscientific Model Development (gmd)
% History of Geo- and Space Sciences (hgss)
% Hydrology and Earth System Sciences (hess)
% Journal of Bone and Joint Infection (jbji)
% Journal of Micropalaeontology (jm)
% Journal of Sensors and Sensor Systems (jsss)
% Magnetic Resonance (mr)
% Mechanical Sciences (ms)
% Natural Hazards and Earth System Sciences (nhess)
% Nonlinear Processes in Geophysics (npg)
% Ocean Science (os)
% Polarforschung - Journal of the German Society for Polar Research (polf)
% Primate Biology (pb)
% Proceedings of the International Association of Hydrological Sciences (piahs)
% Safety of Nuclear Waste Disposal (sand)
% Scientific Drilling (sd)
% SOIL (soil)
% Solid Earth (se)
% The Cryosphere (tc)
% Weather and Climate Dynamics (wcd)
% Web Ecology (we)
% Wind Energy Science (wes)

% Pandoc citation processing

% The "Technical instructions for LaTex" by Copernicus require _not_ to insert any additional packages.
% % % From pandoc table feature
% \usepackage{longtable,booktabs,array}
% % \usepackage{calc} % for calculating minipage widths
% % Correct order of tables after \paragraph or \subparagraph
% \usepackage{etoolbox}
% \makeatletter
% \patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
% \makeatother
% % Allow footnotes in longtable head/foot
% \IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
% \makesavenoteenv{longtable}
% 
% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%
\begin{document}

\title{Transition from forest to agriculture in the Brazilian Amazon from 1985 to 2021}

\Author[1, 2]{Hugo}{Tameirão Seixas}
\Author[2]{Hilton}{Luis Ferraz da Silveira}
\Author[2]{Alan}{Falcão}
\Author[2]{Fabiana}{Da Silva Soares}
\Author[1]{Ramon}{Bicudo}

\affil[1]{Center for Environmental Studies and Research, University of Campinas, Rua dos Flamboyants 155, Brazil}
\affil[2]{Embrapa Territorial, Av. Soldado Passarinho 303, Brazil}

\runningtitle{Transition from forest to agriculture in the Brazilian Amazon from 1985 to 2021}

\runningauthor{Seixas et al.}

\correspondence{Hugo\ Tameirão Seixas\ (seixas.hugo@protonmail.com)}

\received{}
\pubdiscuss{} %% only important for two-stage journals
\revised{}
\accepted{}
\published{}

%% These dates will be inserted by Copernicus Publications during the typesetting process.

\firstpage{1}

\maketitle

\begin{abstract}
A new dataset was created by calculating the time necessary for deforested forests to transition to agriculture in the Brazillian Amazon biome. The new data can be useful in interdiciplinary studies about land cover and land use change in brazil, its drivers and implications. A main inovation is that the dataset links the deforestation year with the year of agriculture establishment, which can provide new information about this process.
\end{abstract}

\copyrightstatement{The author's copyright for this publication is transferred to institution/company.}

\introduction[Introduction]

In the last decades, the Amazon biome have been submitted to strong changes, the advance of agricultural areas over native forests is shaping a new landscape in the region.
The most common pattern of transitions from natural forest formations to agriculture sites, is the initial process of deforestation, followed by the establishment of pasture, where extensive livestock production takes place, until its replacement to agricultural activities.
This process can take several decades to be accomplished, or even less than one year, in which it may be considered as a direct transition from forest to agriculture.

Land cover transitions can cause important impacts on ecosystem properties \citep{Nunes2022}.

This project aims to characterize and quantify the length of the transitions from forest formations to agriculture in the Amazon.

\section{Methods}

The estimations of transition from forest to agriculture were performed for the Amazon biome region, as defined by the Brazilian Institute of Geography and Statistics (IBGE), in 2019.
Transitions were calculated using land use and land cover classification data from MapBiomas \citep{Souza2020}, which ranges from 1985 to 2021, at a spatial resolution of approximately 30 meters.

Transition length was calculated pixel by pixel, by performing the following steps:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Load raster and extract valid values into a table;
\item
  Calculate the year of first occurrence of a non-natural LULC class for each pixel;
\item
  Calculate the first year of ``Forest Formation'' LULC class after the year calculated in step 2, for each pixel;
\item
  Classify rows as ``before'' or ``after'' the occurrence of the year calculated in step 3, for each pixel;
\item
  Calculate the last year of ``Forest Formation'' within the rows classified as ``before'', for each pixel;
\item
  Calculate the first year of any agriculture type class within the rows classified as ``before'', for each pixel;
\item
  Calculate the difference between years from items 5 and 6 to get the LULC transition length in years, for each pixel;
\end{enumerate}

\section{Description of data collection}

We chose two ways to store and share the data.

\begin{itemize}
\item
  One is to store the data from all three tables in three sets of Parquet files, in a structure of folders and sub folders;
\item
  The other is to convert the ``trans\_length.parquet'' tables back to raster files.
  Raster files are useful to efficiently store spatial data, calculate areas, operate in relation to vector objects and perform data summaries, and it is also a very accessible format;
\end{itemize}

The new raster files are created based on the tiles downloaded from GEE (which are in the path ``./data/raw\_raster\_tiles/''), using their extent values and its number of rows and columns (located on the ``./data/trans\_tabular\_dataset/tiles\_metadata.parquet'' file).
For each tile file from ``./data/raw\_raster\_tiles/'', a set of new raster files are created based in the number of agriculture cycle (``agri\_cycle'') found in the tables, resulting in one raster for each agriculture cycle.
Each new raster file contains 6 bands:

\begin{itemize}
\item
  Band 1 = ``agri\_code'' (The agriculture class code that resulted from the transition);
\item
  Band 2 = ``forest\_year'' (The last forest year before the transition, can be interpreted as the deforestation year);
\item
  Band 3 = ``agri\_year'' (The first year of the agriculture establishment, being the end of the transition);
\item
  Band 4 = ``trans\_length'' (The length in years of the transition from forest to agriculture);
\item
  Band 5 = ``trans\_cycle'' (The number of the transition cycle in which the transition occurred);
\item
  Band 6 = ``forest\_type'' (The type of the forest that suffered the transition);
\end{itemize}

The new raster files are saved in the path ``./data/trans\_raster\_tiles/''.

To facilitate the use of these new raster files, we also merged all the raster of each agriculture cycle into a single mosaic, which are saved in the path ``./data/trans\_raster\_mosaic/''.

\section{Results and discussion}

The transition calculations shows that between 1985 and 2021, 64874 squared kilometers of forests were converted to agriculture, in the Brazilian Amazon biome.
The length of the transitions can go from 1 to 36 years, in which transitions closer to 1 year are considered as fast transitions, and transitions closer to 36 years are considered as slow transitions.

\subsection{Transition patterns}

Transitions from forest to agriculture can be found in almost every region in the Amazon, but is mostly concentrated in clusters, specially in the south and east of the biome, in the states of Mato Grosso, Pará and Maranhão (Figure \ref{fig:map-plot}).

\begin{figure}[h]
\includegraphics[width=17cm]{figs/map} \caption{Map of distribution of transitions from forests to agriculture in the Brazillian Amazon biome. The hexagonal cells represent the most common transition length, and do not reflect the amount of area of transitions inside a cell. Transitions are concentraded in the south (Mato Grosso state), and in the east (Pará and Maranhão states). The transition length ranges from 1 (blue tones) to 36 years (pink tones), and clusters of fast transitions (transitions closer to 1 year) can be discerned from clusters of slow transitions (transitions closer to 36 years). The histogram located in the bottom left shows that fast transitions are more common than slower transitions. The zoomed map in the bottom right shows the results in finer resolution, where it is possible to observe different transition lengths between properties.}\label{fig:map-plot}
\end{figure}

Well defined clusters can be observed in the map created with aggregated transition length data (Figure \ref{fig:map-plot}), slow transition areas tend to concentrate in specific regions in the Amazon biome, while fast transitions seem to have a wider distributions, but also tend to form clusters.
However, this pattern does not hold completely when observing the data at its original scale (Figure \ref{fig:map-plot}), where areas with different transition lengths are mixed between each other.
When observing at the original scale, we could not spot any well defined pattern or direction of the occurrence of faster to slower transitions (Figure \ref{fig:map-plot}).

Other studies investigating patterns of transitions in the Amazon also found the formation of clusters of patterns, although there is a big heterogeneity at larger scales \citep{MullerHansen2017}.

The data can be analysed year by year, and also be separated by primary and secondary forests being converted to agriculture (Figure \ref{fig:transbar-plot}).

\begin{figure}[h]
\includegraphics[width=17cm]{figs/trans_length_cols} \caption{Transition area per year and transition length. The bars represent the total amount of area at some state of the transition for each year. The color gradient in each bar represents the transition length related to a deforestation or a agriculture establishment event. Blue tones represents fast transitions (transitions closer to one year), pink tones represents slow transitions (transitions closer to 36 years). Transition events were separated by deforestation of primary forests (a) and secondary forests (c), and the subsequent  agriculture establishment of primary forests (b) and secondary forests (d). TODO: add labels for important years hwere public policies took place.}\label{fig:transbar-plot}
\end{figure}

The deforestation area of primary forests increased largely from 1985 to 2002, which was followed be a steep decrease until 2007, when the deforested areas being converted to agriculture reached a stable rate.
From 1985 to approximately 1995, most of deforested areas suffered a slow transition, mostly were higher than 10 years, after this period, fast transitions started to become more common, specially from 2001 to 2003.

Agriculture establishment over areas of primary forests peaked in 2005 and 2013.
Despite similar rates between both years, their transition lengths differ greatly, in 2005 most of the transitions were faster than 10 years, while in 2013 the great majority of transitions were slower than 10 years.
The year of 2003 marked a change in the transition length of establishment of agriculture areas, after this year, most of the transitions happened in areas deforested at least 10 years before.
Even after the decrease of deforestation after 2002, agriculture areas are expanding over lands where deforestation happened before 2002.
However, after 2018, a sudden drop of agriculture establishment rate happened.

In 2006, the Brazilian Association of Vegetable Oil Industries (BIOVE) and the National Association of Cereal Exporters (ANEC) committed to avoid commercialization of soy grains harvested from areas deforested after 2006.
Our estimates of transitions show a decrease of deforested areas to be converted to agriculture after 2007, where it reached minimal values (Figure \ref{fig:transbar-plot}.a).
After 2006, agriculture establishment over deforested primary forests suffered a decrease, which stayed relatively stable until 2012, where a steep increase occurred, however, the new areas being occupied by agriculture were mainly over areas that were cleared more than a decade before (therefore, before 2006) (Figure \ref{fig:transbar-plot}.b) This shows that agriculture expansion did not halt after the soy moratorium, producers started expanding in old cleared areas.
Expansion of agriculture areas also expanded over cleared areas of secondary forests in 2012, but with an important amount of fast transitions.
The causes of the increase of agriculture establishment areas can be numerous, one main driver was the approval of a new Forest Code, in 2012, which is considered to have undermined the environmental protection \citep{Kroger2017, Pereira2019}.

The transition length patterns across years can change significantly between different states (Figure \ref{fig:transridge-plot}).

\begin{figure}[h]
\includegraphics[width=17cm]{figs/trans_ridge} \caption{Transition length patterns inside states that belongs to the Amazon biome.}\label{fig:transridge-plot}
\end{figure}

The state of Amapá presented fast transitions along all the time series, where slow transitions are not as common.
In contrast, Acre shows a majority of slow transitions, in which only 2021 showed more fast transitions.There are three states where the pattern of transitions length across time are alike, Mato Grosso, Pará and Rondônia presented more fast transitions from 1995 to 2005, after this period, slow transitions became more common with time, until around 2015, where the occurrence of fast transitions were the majority again.

\subsection{Validation}

When analyzing the errors from the transition length estimates, we observe that the year of deforestation shows the least amount of errors (Figure \ref{fig:errorbar-plot}).
The MAE of the deforestation year is of 1.42 years, and the error shows a bias towards underestimation.
The year of the agriculture establishment and the transition length estimates showed larger errors when comparing with visual inspection.

\begin{figure}[ht]
\includegraphics[width=17cm]{figs/error_bars} \caption{TODO: Add title.}\label{fig:errorbar-plot}
\end{figure}

\begin{figure}[ht]
\includegraphics[width=17cm]{figs/error_scatter} \caption{TODO: Change line showing MAE to shaded areas showing quantiles of the errors. Add title.}\label{fig:errorscatter-plot}
\end{figure}

\subsection{Qualitative assessment}

\conclusions[Conclusions]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\competinginterests{The authors declare no competing interests.} %% this section is mandatory even if you declare that no competing interests are present

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% REFERENCES
%% DN: pre-configured to BibTeX for rticles

%% The reference list is compiled as follows:
%%
%% \begin{thebibliography}{}
%%
%% \bibitem[AUTHOR(YEAR)]{LABEL1}
%% REFERENCE 1
%%
%% \bibitem[AUTHOR(YEAR)]{LABEL2}
%% REFERENCE 2
%%
%% \end{thebibliography}

%% Since the Copernicus LaTeX package includes the BibTeX style file copernicus.bst,
%% authors experienced with BibTeX only have to include the following two lines:
%%
\bibliographystyle{copernicus}
\bibliography{references.bib}
%%
%% URLs and DOIs can be entered in your BibTeX file as:
%%
%% URL = {http://www.xyz.org/~jones/idx_g.htm}
%% DOI = {10.5194/xyz}

%% LITERATURE CITATIONS
%%
%% command                        & example result
%% \citet{jones90}|               & Jones et al. (1990)
%% \citep{jones90}|               & (Jones et al., 1990)
%% \citep{jones90,jones93}|       & (Jones et al., 1990, 1993)
%% \citep[p.~32]{jones90}|        & (Jones et al., 1990, p.~32)
%% \citep[e.g.,][]{jones90}|      & (e.g., Jones et al., 1990)
%% \citep[e.g.,][p.~32]{jones90}| & (e.g., Jones et al., 1990, p.~32)
%% \citeauthor{jones90}|          & Jones et al.
%% \citeyear{jones90}|            & 1990

\end{document}

### 2022-10-27 scripts/01_download.R
Update coment.

@@ -65,7 +65,7 @@ aoi <- sf_as_ee(biomes)
 
 # LOAD LULC DATA --------------------------------------------------------------
 
-## Load MapBiomas collection 6 ----
+## Load MapBiomas collection 7 ----
 mb_img <-
   ee$Image(
     paste(

### 2022-10-27 scripts/06_transition_analysis.R
Format figures. Add code to create map.

@@ -18,6 +18,7 @@
 #
 library(sf)
 library(geobr)
+library(terra)
 library(magrittr)
 library(fs)
 library(lubridate)
@@ -31,6 +32,8 @@ library(readr)
 library(tidyr)
 library(forcats)
 library(ggplot2)
+library(mapsf)
+library(scico)
 #
 # OPTIONS ---------------------------------------------------------------------
 #
@@ -52,7 +55,20 @@ names(palette) <- read_csv('data/mb_class_dictionary.csv')$class_code
 ## Load municipality data ----
 municip <- read_municipality(year = "2019")
 
+amazon <- read_biomes() %>%
+  filter(code_biome == 1) %>%
+  st_transform("EPSG:4326")
+
 # LOAD DATASETS ---------------------------------------------------------------
+
+## Load transition length raster ----
+mosaic <-
+  rast("data/trans_raster_mosaic/mb_mosaic_cycle_1.tif")[["trans_length"]]
+
+mosaic_agg <- terra::aggregate(mosaic, fact = 10, fun = "modal")
+
+## Create hex grid ----
+grid <- vect(st_make_grid(mosaic_agg, cellsize = 0.1, square = FALSE))
 
 ## Connect with Spark -----
 sc <- spark_connect(master = "local", config = config, version = "3.3.0")
@@ -85,6 +101,27 @@ trans_classes <-
   )
 
 # FILTER AND AGGREGATE DATA ---------------------------------------------------
+
+## Extract trans_length values from raster into hex grid ----
+extract_values <- terra::extract(mosaic_agg, grid) %>%
+  drop_na() %>%
+  group_by(ID, trans_length) %>%
+  count() %>%
+  group_by(ID) %>%
+  filter(n == max(n))
+
+trans_hex <- grid %>%
+  st_as_sf() %>%
+  mutate(ID := seq_len(nrow(.))) %>%
+  left_join(., extract_values, by = "ID") %>%
+  filter(n > 20) %>%
+  drop_na()
+
+## Get the frequency of trans_lenght ----
+frequency <- trans_length %>%
+  select(trans_length) %>%
+  sdf_sample(fraction = 0.01, replacement = FALSE, seed = 2) %>%
+  collect()
 
 ## Get LULC percentage during and after transition ----
 trans_time_serie <-
@@ -137,7 +174,10 @@ trans_ridges <-
[…]

### 2022-10-27 scripts/validation_app.R
Add script that generates app to perform visual analysis of satelite images.

@@ -1,0 +1,327 @@
+
+library(sf)
+library(terra)
+library(terrainr)
+library(magrittr)
+library(fs)
+library(lubridate)
+library(glue)
+library(dplyr)
+library(tibble)
+library(readr)
+library(tidyr)
+library(forcats)
+library(purrr)
+library(stringr)
+library(ggplot2)
+library(shiny)
+library(shinythemes)
+library(gridlayout)
+
+proj_path <- here::here()
+
+sample_paths <- dir_ls(glue("{proj_path}/data/validation/landsat/"))
+
+sample_cell <- as.list(as.integer(str_extract(sample_paths, "(\\d)+")))
+names(sample_cell) <- as.character(1:length(sample_cell))
+
+sample_points <- read_sf(glue("{proj_path}/data/validation/sample.fgb"))
+
+results <-
+  tibble(
+    sample_id = NA_real_,
+    cell_id = NA_real_,
+    forest_year = NA_real_,
+    agri_year = NA_real_,
+    trans_length = NA_real_
+  ) %>%
+  drop_na()
+
+if (!file_exists(glue("{proj_path}/data/validation/results.csv"))) {
+
+  write_csv(
+    results,
+    glue("{proj_path}/data/validation/results.csv")
+  )
+
+}
+
+# Define app template
+ui <- grid_page(
+  theme = bslib::bs_theme(bootswatch = "slate"),
+  layout = c(
+    "header  header      header       header     ",
+    "sidebar composite1  composite2   composite3 ",
+    "sidebar forest      agri         length     "
+  ),
+  row_sizes = c(
+    "70px",
+    "1fr",
+    "1fr"
+  ),
+  col_sizes = c(
+    "225px",
+    "1fr",
+    "1fr",
+    "1fr"
+  ),
+  gap_size = "1rem",
+  grid_card(
+    area = "sidebar",
+    item_alignment = "top",
+    title = "Settings",
+    item_gap = "12px",
+    selectInput(
+      inputId = "sampleCell",
+      label = "Sample",
+      choices = sample_cell
+    ),
+    radioButtons(
+      inputId = "seasonOpt",
+      label = "Season",
+      choices = list(
+        Wet = "wet",
+        Dry = "dry"
+      ),
+      width = "100%"
+    ),
+    numericInput(
+      inputId = "imageYear",
+      label = "View Year",
+      value = 1985L,
+      min = 1985L,
+      max = 2020L,
+      step = 1L,
[…]

### 2022-10-28 scripts/06_transition_analysis.R
Add new dependencies. Improve figures.

@@ -32,7 +32,9 @@ library(readr)
 library(tidyr)
 library(forcats)
 library(ggplot2)
-library(mapsf)
+library(ggtext)
+library(ggfx)
+library(patchwork)
 library(scico)
 #
 # OPTIONS ---------------------------------------------------------------------
@@ -240,6 +242,7 @@ ggsave(
 cty <- trans_subset %>%
   group_by(transition, forest_type, trans_length, year) %>%
   summarise(total_area = sum(total_area, na.rm = TRUE), .groups = "drop") %>%
+  mutate(total_area = total_area / 1e6) %>%
   ggplot() +
   facet_grid(
     facets = transition ~ forest_type,
@@ -254,20 +257,53 @@ cty <- trans_subset %>%
       )
     )
   ) +
+  geom_vline(
+    xintercept = ymd("2003-01-01", "2006-01-01", "2012-01-01"),
+    color = "white",
+    linetype = "dashed"
+  ) +
+  geom_richtext(
+    data = tibble(
+      year = ymd("2003-05-01", "2006-05-01", "2012-05-01"),
+      total_area = c(5000, 3500, 2000),
+      forest_type = factor(1),
+      transition = factor("forest_year"),
+      label = c("Deforestation Act", "Soy Moratorium", "Forest Code")
+    ),
+    aes(x = year, y = total_area, label = label),
+    size = 2.5,
+    hjust = "left",
+    label.r = unit(0, "pt"),
+    label.padding = unit(2, "pt"),
+    fill = "#dddddd"
+  ) +
   geom_col(
     aes(
       x = year,
-      y = total_area / 1e6,
+      y = total_area,
       fill = trans_length,
       color = trans_length,
       group = trans_length
     ),
-    size = 0.25
+    lwd = 0.1
+  ) +
+  geom_col(
+    data = . %>%
+      group_by(transition, forest_type, year) %>%
+      summarise(total_area = sum(total_area, na.rm = TRUE), .groups = "drop"),
+    aes(
+      x = year,
+      y = total_area
+    ),
+    fill = "transparent",
+    color = "black",
+    lwd = 0.2
   ) +
   geom_text(
     data = tibble(
       label = c("(a)", "(b)", "(c)", "(d)"),
-      transition = as.factor(c("forest_year", "agri_year", "forest_year", "agri_year")),
+      transition =
[…]

### 2022-10-28 scripts/08_validation_results.R
Add dependencies. Improve figures.

@@ -30,7 +30,9 @@ library(readr)
 library(tidyr)
 library(forcats)
 library(purrr)
+library(scales)
 library(ggplot2)
+library(ggtext)
 #
 # OPTIONS ---------------------------------------------------------------------
 #
@@ -156,23 +158,77 @@ ggsave(
   dpi = 600
 )
 
+quants <- comparison %>%
+  group_by(var) %>%
+  summarise(
+    value = quantile(
+      difference,
+      probs = c(0.05, 0.25, 0.75, 0.95),
+      na.rm = TRUE
+    )
+  ) %>%
+  mutate(
+    lim = c("min", "min", "max", "max"),
+    group_id = c(1, 2, 2, 1)
+  ) %>%
+  pivot_wider(
+    id_cols = c(var, group_id),
+    names_from = lim,
+    values_from = value
+  )
+
+quants <- do.call("rbind", replicate(37, quants, simplify = FALSE)) %>%
+  group_by(var, group_id) %>%
+  mutate(estimated = 0:36, observed = 0:36)
+
 ## Create scatter plot with the errors ----
 scatter_plot <- comparison %>%
   ggplot() +
   facet_wrap( ~ var) +
-  geom_abline(intercept = 0, slope = 1, color = "white") +
-  geom_abline(
-    data = mae,
-    aes(intercept = mae, slope = 1),
-    color = "white", linetype = "dashed"
+  geom_ribbon(
+    data = subset(quants, group_id == 2),
+    aes(
+      ymin = observed + min, ymax = max + observed,
+      xmin = estimated + min, xmax = max + estimated,
+      x = estimated, y = observed
+    ),
+    fill = "white", alpha = 0.7
+  ) +
+  geom_ribbon(
+    data = subset(quants, group_id == 1),
+    aes(
+      ymin = observed + min, ymax = max + observed,
+      xmin = estimated + min, xmax = max + estimated,
+      x = estimated, y = observed
+    ),
+    fill = "white", alpha = 0.3
+  ) +
+  geom_abline(intercept = 0, slope = 1, color = "#6a6a6a") +
+  geom_text(
+    data = tibble(
+      quant = c("5%", "25%", "75%", "95%"),
+      observed = c(17 + 12, 17 + 4.75, 17 - 1.75, 17 - 8.7),
+      estimated = c(16.5, 16.5, 16.5, 16.5),
+      var = factor("agri_year")
+    ),
[…]

### 2022-11-02 scripts/08_validation_results.R
Improve validation plots.

@@ -79,7 +79,13 @@ setdiff(unique(observed$cell_id), unique(estimated$cell_id))
 ## Join tables ----
 comparison <- estimated %>%
   inner_join(observed, by = c("cell_id", "var")) %>%
-  mutate(difference = observed - estimated)
+  mutate(
+    difference = observed - estimated,
+    var = factor(
+      var,
+      levels = c("forest_year", "agri_year", "trans_length")
+    )
+  )
 
 # ANALYSE RESULTS -------------------------------------------------------------
 
@@ -107,6 +113,7 @@ metrics <- mae %>%
 bar_plot <- comparison %>%
   ggplot() +
   facet_wrap( ~ var) +
+  ggtitle("Distribution of the errors of transition estimates.") +
   geom_vline(
     data = mae,
     aes(xintercept = mae),
@@ -153,7 +160,7 @@ ggsave(
   glue("./figs/error_bars.png"),
   bar_plot,
   width = 17,
-  height = 7.5,
+  height = 8.5,
   units = "cm",
   dpi = 600
 )
@@ -185,6 +192,7 @@ quants <- do.call("rbind", replicate(37, quants, simplify = FALSE)) %>%
 scatter_plot <- comparison %>%
   ggplot() +
   facet_wrap( ~ var) +
+  ggtitle("Dispersion between observed and estimated transitions.") +
   geom_ribbon(
     data = subset(quants, group_id == 2),
     aes(
@@ -249,7 +257,7 @@ ggsave(
   glue("./figs/error_scatter.png"),
   scatter_plot,
   width = 17,
-  height = 7,
+  height = 8,
   units = "cm",
-  dpi = 600
+  dpi = 300
 )

### 2022-11-04 scripts/03_extract_transition.R
Convert variables to years. Foest year is now deforestation year.

@@ -259,8 +259,8 @@ walk(
           trans_length %<>%
             rename(
               agri_code = class_code,
-              forest_year = forest_max,
-              agri_year = agri_min,
+              forest_year = forest_max + 1 + 1984L, # Deforestation year
+              agri_year = agri_min + 1984, # Convert to year
             ) %>%
             select(
               cell_id, agri_code, forest_year,

### 2022-11-04 scripts/03_extract_transition.R
Small fix on dates.

@@ -259,8 +259,12 @@ walk(
           trans_length %<>%
             rename(
               agri_code = class_code,
-              forest_year = forest_max + 1 + 1984L, # Deforestation year
-              agri_year = agri_min + 1984, # Convert to year
+              forest_year = forest_max,
+              agri_year = agri_min
+            ) %>%
+            mutate(
+              forest_year = forest_year + 1 + 1984L, # Deforestation year
+              agri_year = agri_year + 1984 # Convert to year
             ) %>%
             select(
               cell_id, agri_code, forest_year,

### 2022-11-08 R/terra_as_tibble.R
Delete custom function

@@ -1,57 +1,0 @@
-# Convert a SpatRaster to a data frame
-
-# Adapted from the tabularaster package create by Michael Sumner.
-# Instead of creating a tibble from a raster from {raster} package
-# I introduced the raster object from the {terra} package, which the extraction
-# of values is faster.
-
-terra_as_tibble <-
-  function(
-    x,
-    cell = TRUE,
-    dim = nlyr(x) > 1L,
-    value = TRUE,
-    xy = FALSE,
-    ...
-  ) {
-
-    dimindex <- seq(terra::nlyr(x))
-
-    cellvalue <- cellindex <- NULL
-
-    if (value) { cellvalue <- as.vector(terra::values(x)) }
-
-    cellindex <-  rep(seq(terra::ncell(x)), terra::nlyr(x))
-
-    d <- dplyr::bind_cols(cellvalue = cellvalue, cellindex = cellindex)
-
-    if (dim) {
-
-      dimindex <- rep(dimindex, each = terra::ncell(x))
-
-      d[["dimindex"]] <- dimindex
-
-    }
-
-    if (xy) {
-
-      xy <- terra::xyFromCell(x, seq_len(terra::ncell(x)))
-
-      colnames(xy) <- c("x", "y")
-
-      if (terra::nlyr(x) > 1) {
-
-        xy <- xy[rep(seq_len(terra::ncell(x)), terra::nlyr(x)), ]
-
-      }
-
-      d <- dplyr::bind_cols(d, tibble::as_tibble(xy))
-
-    }
-
-    if (!cell) d[["cellindex"]] <- NULL
-
-    return(d)
-
-  }
-

### 2022-11-08 manuscript/manuscript.Rmd
Add small paragraph in the results.
Add text about deforestation measured by MapBiomas.

@@
 The length of the transitions can go from 1 to 36 years, in which transitions closer to 1 year are considered as fast transitions, and transitions closer to 36 years are considered as slow transitions.
+Transitions of 1 year are considered as "direct" transitions, where there were no presence of pasture before the establishment of agriculture, out estimations shows that around 9.2 % of the transitions are considered as "direct".
+
+Although we named as deforestation the last years identified as forest before classification of antropic cover, we acknowledge that it is not a direct measurement of deforestation (such as PRODES), it is however a proxy to deforestation.
 
@@
 
-The deforestation area of primary forests increased largely from 1985 to 2002, which was followed be a steep decrease until 2007, when the deforested areas being converted to agriculture reached a stable rate.
-From 1985 to approximately 1995, most of deforested areas suffered a slow transition, mostly were higher than 10 years, after this period, fast transitions started to become more common, specially from 2001 to 2003.
+The deforestation area of primary forests increased largely from 1986 to 2003, which was followed be a steep decrease until 2008, when the deforested areas being converted to agriculture reached a stable rate.
+From 1986 to approximately 1995, most of deforested areas suffered a slow transition, mostly were higher than 10 years, after this period, fast transitions started to become more common, specially from 2002 to 2004.
 
@@
 Expansion of agriculture areas also expanded over cleared areas of secondary forests in 2012, but with an important amount of fast transitions.
-The causes of the increase of agriculture establishment areas can be numerous, one main driver was the approval of a new Forest Code, in 2012, which is considered to have undermined the environmental protection [@Kroger2017; @Pereira2019].
+The causes of the increase of agriculture establishment areas can be numerous, one main driver was the approval of a new Forest Code, in 2012, which is considered to have undermined the environmental protection of forests [@Kroger2017; @Pereira2019].
 

### 2022-11-08 manuscript/manuscript.tex
Add small paragraph in the results.
Update manuscript

@@
 The length of the transitions can go from 1 to 36 years, in which transitions closer to 1 year are considered as fast transitions, and transitions closer to 36 years are considered as slow transitions.
+Transitions of 1 year are considered as ``direct'' transitions, where there were no presence of pasture before the establishment of agriculture, out estimations shows that around 9.2 \% of the transitions are considered as ``direct''.
+
+Although we named as deforestation the last years identified as forest before classification of antropic cover, we acknowledge that it is not a direct measurement of deforestation (such as PRODES), it is however a proxy to deforestation.
 
@@
 Expansion of agriculture areas also expanded over cleared areas of secondary forests in 2012, but with an important amount of fast transitions.
-The causes of the increase of agriculture establishment areas can be numerous, one main driver was the approval of a new Forest Code, in 2012, which is considered to have undermined the environmental protection \citep{Kroger2017, Pereira2019}.
+The causes of the increase of agriculture establishment areas can be numerous, one main driver was the approval of a new Forest Code, in 2012, which is considered to have undermined the environmental protection of forests \citep{Kroger2017, Pereira2019}.
 

### 2022-11-08 scripts/02_extract_mask.R
Substitute function to transform raster to table.

@@ -30,7 +30,6 @@ library(dplyr)
 # OPTIONS ---------------------------------------------------------------------
 #
 source("conf/config.R")
-source("R/terra_as_tibble.R")
 sf::sf_use_s2(TRUE)
 #
 # LOAD AUXILIARY DATA ---------------------------------------------------------
@@ -119,11 +118,11 @@ mask_table <-
       # Extract values of a single tile ----
       mask_subset <-
         raster %>%
-        terra_as_tibble(xy = TRUE, cell = TRUE) %>%
-        rename(tile_cell_id = cellindex) %>%
-        filter(!(cellvalue == 0)) %>% # Remove empty cells
-        mutate(tile_id = tile) %>%
-        rename(area = cellvalue)
+        as.data.frame(xy = TRUE, cell = TRUE) %>%
+        as_tibble() %>%
+        rename(tile_cell_id = cell) %>%
+        filter(!(area == 0)) %>% # Remove empty cells
+        mutate(tile_id = tile)
 
       # Check if there is any valid pixel ----
       # Avoids possible unexpected errors

### 2022-11-08 scripts/03_extract_transition.R
Fix dates. Change function to transform raster to table.

@@ -88,13 +88,16 @@ walk(
     ## Convert raster stack to tibble ----
     lulc_table <-
       mb_raster %>%
-      terra_as_tibble() %>%
-      filter(cellvalue != 0) %>%
-      rename(
-        class_code = cellvalue,
-        tile_cell_id = cellindex,
-        year = dimindex
+      as.data.frame(cells = TRUE) %>%
+      as_tibble() %>%
+      pivot_longer(
+        cols = starts_with("classification"),
+        names_to = "year",
+        values_to = "class_code"
       ) %>%
+      filter(class_code != 0) %>%
+      mutate(year = as.integer(str_extract(year, "(\\d)+"))) %>%
+      rename(tile_cell_id = cell) %>%
       left_join(mask_subset, by = "tile_cell_id") # Get cell_id from mask
 
     ## Check if raster have any value != 0 ----
@@ -213,7 +216,8 @@ walk(
             filter(class_type == 'forest') %>%
             group_by(cell_id) %>%
             slice(which.max(year)) %>%
-            rename(forest_max = year)
+            rename(forest_max = year) %>%
+            mutate(forest_max = forest_max + 1) # Get deforestation year
 
           #### Get the min year of agriculture in the "before" period ----
           agri_min <-
@@ -263,8 +267,8 @@ walk(
               agri_year = agri_min
             ) %>%
             mutate(
-              forest_year = forest_year + 1 + 1984L, # Deforestation year
-              agri_year = agri_year + 1984 # Convert to year
+              forest_year = forest_year,
+              agri_year = agri_year
             ) %>%
             select(
               cell_id, agri_code, forest_year,
@@ -284,7 +288,7 @@ walk(
           filter(period == 'after') %>%
           select(-period)
 
-        # Check if there is any data in the update version of the table
+        # Check if there is any data in the updated version of the table
         if (nrow(lulc_table) == 0) {
 
           break

### 2022-11-08 scripts/04_rasterize.R
Fix data type of rasters.

@@ -151,7 +151,7 @@ walk(
               'trans_stack_{meta$tile_code}_cycle_{.y}.tif'
             ),
             wopt = list(
-              datatype = 'INT1U',
+              datatype = 'INT2U',
               progress = 0
             ),
             overwrite = TRUE

### 2022-11-08 scripts/05_merge_raster.R

@@ -90,7 +90,7 @@ walk(
       ),
       overwrite = TRUE,
       wopt = list(
-        datatype = "INT1U",
+        datatype = "INT2U",
         progress = 0,
         gdal = c("COMPRESS=LZW")
       )

### 2022-11-08 scripts/06_transition_analysis.R
Delete scripts to create figures.

@@ -1,525 +1,0 @@
-# HEADER ----------------------------------------------------------------------
-#
-# Title:        Analyze the forest-agriculture transition dataset
-# Description:  The objective of this routine is to explore the dataset by
-#               filtering and aggregating data in Spark. Results are supposed
-#               to give us a overall view of the data by plots and tables.
-#
-# Author:       Hugo Tameirao Seixas
-# Contact:      tameirao.hugo@gmail.com
-# Date:         2020-11-10
-#
-# Notes:        In order to run this routine, you will need to successfully
-#               install and connect to Spark using the {sparklyr} package.
-#               Installation and use guide can be found in:
-#               https://github.com/sparklyr/sparklyr
-#
-# LIBRARIES -------------------------------------------------------------------
-#
-library(sf)
-library(geobr)
-library(terra)
-library(magrittr)
-library(fs)
-library(lubridate)
-library(glue)
-library(scales)
-library(sparklyr)
-library(ggridges)
-library(dplyr)
-library(tibble)
-library(readr)
-library(tidyr)
-library(forcats)
-library(ggplot2)
-library(ggtext)
-library(ggfx)
-library(patchwork)
-library(scico)
-#
-# OPTIONS ---------------------------------------------------------------------
-#
-source('conf/spark_config.R')
-#
-# SET COLOR PALETTE AND LULC NAMES --------------------------------------------
-
-## Get the names for each LULC class from MapBiomas ----
-mb_dict <- read_csv("data/mb_class_dictionary.csv") %>%
-  select(class_code, class_name) %>%
-  rename(agri_code = class_code)
-
-## Get the colors for each LULC class from MapBiomas
-palette <- read_csv('data/mb_class_dictionary.csv')$class_color
-names(palette) <- read_csv('data/mb_class_dictionary.csv')$class_code
-
-# LOAD AUXILIARY DATA ---------------------------------------------------------
-
-## Load municipality data ----
-municip <- read_municipality(year = "2019")
-
-amazon <- read_biomes() %>%
-  filter(code_biome == 1) %>%
[…]

### 2022-11-08 scripts/08_validation_results.R

@@ -1,263 +1,0 @@
-# HEADER ----------------------------------------------------------------------
-#
-# Title:        Validate transitions against visual inspection
-# Description:  The objective of this routine is compare the transition length
-#               results, comparing them with visual observations.
-#
-#
-# Author:       Hugo Tameirao Seixas
-# Contact:      tameirao.hugo@gmail.com
-# Date:         2022-08-03
-#
-# Notes:
-#
-#
-#
-#
-# LIBRARIES -------------------------------------------------------------------
-#
-library(sf)
-library(terra)
-library(magrittr)
-library(fs)
-library(lubridate)
-library(glue)
-library(arrow)
-library(dplyr)
-library(units)
-library(tibble)
-library(readr)
-library(tidyr)
-library(forcats)
-library(purrr)
-library(scales)
-library(ggplot2)
-library(ggtext)
-#
-# OPTIONS ---------------------------------------------------------------------
-#
-#
-# LOAD DATA -------------------------------------------------------------------
-
-## Load observed data ----
-observed <- read_csv("data/validation/results.csv") %>%
-  mutate( # Remove points where there were no transition
-    forest_year = if_else(forest_year <= 0, NA_real_, forest_year),
-    agri_year = if_else(agri_year <= 0, NA_real_, agri_year),
-    trans_length = if_else(agri_year <= 0, NA_real_, trans_length),
-    trans_length = if_else(forest_year <= 0, NA_real_, trans_length)
-  ) %>%
-  mutate( # Fix dates to match the estimates
-    forest_year = forest_year - 1984 - 1, # - 1 is for last year of forest
-    agri_year = agri_year - 1984
-  ) %>%
-  pivot_longer(
-    cols =  c(forest_year, agri_year, trans_length),
-    names_to = "var",
-    values_to = "observed"
-  ) %>%
-  arrange(cell_id)
-
-## Load estimates data ----
-estimated <- open_dataset("data/trans_tabular_dataset/trans_length/")
-
-estimated <- estimated %>%
-  filter(
-    cell_id %in% observed$cell_id
-  ) %>%
-  collect() %>%
-  pivot_longer(
[…]

### 2022-11-08 scripts/figures/fig_1.R
Add scripts to create figures separately.

@@ -1,0 +1,208 @@
+# HEADER ----------------------------------------------------------------------
+#
+# Title:        Create figure 1
+# Description:  The objective of this routine is to explore the dataset by
+#               filtering and aggregating data in Spark. Results are supposed
+#               to give us a overall view of the data by plots and tables.
+#
+# Author:       Hugo Tameirao Seixas
+# Contact:      tameirao.hugo@gmail.com
+# Date:         2020-11-10
+#
+# Notes:        In order to run this routine, you will need to successfully
+#               install and connect to Spark using the {sparklyr} package.
+#               Installation and use guide can be found in:
+#               https://github.com/sparklyr/sparklyr
+#
+# LIBRARIES -------------------------------------------------------------------
+#
+library(sf)
+library(geobr)
+library(terra)
+library(magrittr)
+library(sparklyr)
+library(dplyr)
+library(tibble)
+library(readr)
+library(tidyr)
+library(forcats)
+library(ggplot2)
+library(ggfx)
+library(patchwork)
+library(scico)
+#
+# OPTIONS ---------------------------------------------------------------------
+#
+source('conf/spark_config.R')
+#
+# LOAD AUXILIARY DATA ---------------------------------------------------------
+
+amazon <- read_biomes() %>%
+  filter(code_biome == 1) %>%
+  st_transform("EPSG:4326")
+
+# LOAD DATASETS ---------------------------------------------------------------
+
+## Load transition length raster ----
+mosaic <-
+  rast("data/trans_raster_mosaic/mb_mosaic_cycle_1.tif")[["trans_length"]]
+
+mosaic_agg <- terra::aggregate(mosaic, fact = 10, fun = "modal")
+
+## Create hex grid ----
+grid <- vect(st_make_grid(mosaic_agg, cellsize = 0.1, square = FALSE))
+
+## Connect with Spark -----
+sc <- spark_connect(master = "local", config = config, version = "3.3.0")
+
+## Load trans_length table ----
+trans_length <-
+  spark_read_parquet(
+    sc,
+    name = "trans_length",
+    path = "data/trans_tabular_dataset/trans_length/",
+    memory = FALSE
+  )
+
[…]

### 2022-11-08 scripts/figures/fig_2.R

@@ -1,0 +1,230 @@
+# HEADER ----------------------------------------------------------------------
+#
+# Title:        Create figure 2
+# Description:  The objective of this routine is to explore the dataset by
+#               filtering and aggregating data in Spark. Results are supposed
+#               to give us a overall view of the data by plots and tables.
+#
+# Author:       Hugo Tameirao Seixas
+# Contact:      tameirao.hugo@gmail.com
+# Date:         2020-11-10
+#
+# Notes:        In order to run this routine, you will need to successfully
+#               install and connect to Spark using the {sparklyr} package.
+#               Installation and use guide can be found in:
+#               https://github.com/sparklyr/sparklyr
+#
+# LIBRARIES -------------------------------------------------------------------
+#
+library(sparklyr)
+library(magrittr)
+library(lubridate)
+library(scales)
+library(dplyr)
+library(tibble)
+library(readr)
+library(tidyr)
+library(forcats)
+library(ggplot2)
+library(ggtext)
+library(scico)
+#
+# OPTIONS ---------------------------------------------------------------------
+#
+source('conf/spark_config.R')
+#
+# SET LULC NAMES --------------------------------------------------------------
+
+## Get the names for each LULC class from MapBiomas ----
+mb_dict <- read_csv("data/mb_class_dictionary.csv") %>%
+  select(class_code, class_name) %>%
+  rename(agri_code = class_code)
+
+# LOAD DATASETS ---------------------------------------------------------------
+
+## Connect with Spark -----
+sc <- spark_connect(master = "local", config = config, version = "3.3.0")
+
+## Load mask_cells table ----
+mask_cells <-
+  spark_read_parquet(
+    sc,
+    name = "mask_cells",
+    path = "data/trans_tabular_dataset/mask_cells/",
+    memory = FALSE
+  )
+
+## Load trans_length table ----
+trans_length <-
+  spark_read_parquet(
+    sc,
+    name = "trans_length",
+    path = "data/trans_tabular_dataset/trans_length/",
+    memory = FALSE
+  )
+
+# FILTER AND AGGREGATE DATA ---------------------------------------------------
+
[…]

### 2022-11-08 scripts/figures/fig_3.R

@@ -1,0 +1,153 @@
+# HEADER ----------------------------------------------------------------------
+#
+# Title:        Create figure 3
+# Description:  The objective of this routine is to explore the dataset by
+#               filtering and aggregating data in Spark. Results are supposed
+#               to give us a overall view of the data by plots and tables.
+#
+# Author:       Hugo Tameirao Seixas
+# Contact:      tameirao.hugo@gmail.com
+# Date:         2020-11-10
+#
+# Notes:        In order to run this routine, you will need to successfully
+#               install and connect to Spark using the {sparklyr} package.
+#               Installation and use guide can be found in:
+#               https://github.com/sparklyr/sparklyr
+#
+# LIBRARIES -------------------------------------------------------------------
+#
+library(sf)
+library(geobr)
+library(magrittr)
+library(lubridate)
+library(scales)
+library(sparklyr)
+library(ggridges)
+library(dplyr)
+library(tibble)
+library(readr)
+library(tidyr)
+library(forcats)
+library(ggplot2)
+library(scico)
+#
+# OPTIONS ---------------------------------------------------------------------
+#
+source('conf/spark_config.R')
+#
+# LOAD AUXILIARY DATA ---------------------------------------------------------
+
+## Load municipality data ----
+municip <- read_municipality(year = "2019")
+
+# LOAD DATASETS ---------------------------------------------------------------
+
+## Connect with Spark -----
+sc <- spark_connect(master = "local", config = config, version = "3.3.0")
+
+## Load mask_cells table ----
+mask_cells <-
+  spark_read_parquet(
+    sc,
+    name = "mask_cells",
+    path = "data/trans_tabular_dataset/mask_cells/",
+    memory = FALSE
+  )
+
+## Load trans_length table ----
+trans_length <-
+  spark_read_parquet(
+    sc,
+    name = "trans_length",
+    path = "data/trans_tabular_dataset/trans_length/",
+    memory = FALSE
+  )
+
+# FILTER AND AGGREGATE DATA ---------------------------------------------------
+
+## Sample rows to create ridge plot ----
+trans_ridges <-
+  trans_length %>%
+  filter(agri_cycle == 1) %>%
[…]

### 2022-11-08 scripts/figures/fig_4.R

@@ -1,0 +1,188 @@
+# HEADER ----------------------------------------------------------------------
+#
+# Title:        Validate transitions against visual inspection
+# Description:  The objective of this routine is compare the transition length
+#               results, comparing them with visual observations.
+#
+#
+# Author:       Hugo Tameirao Seixas
+# Contact:      tameirao.hugo@gmail.com
+# Date:         2022-08-03
+#
+# Notes:
+#
+#
+#
+#
+# LIBRARIES -------------------------------------------------------------------
+#
+library(arrow)
+library(magrittr)
+library(lubridate)
+library(glue)
+library(dplyr)
+library(units)
+library(tibble)
+library(readr)
+library(tidyr)
+library(forcats)
+library(purrr)
+library(scales)
+library(ggplot2)
+library(ggtext)
+#
+# OPTIONS ---------------------------------------------------------------------
+#
+#
+# LOAD DATA -------------------------------------------------------------------
+
+## Load observed data ----
+observed <- read_csv("data/validation/results.csv") %>%
+  mutate( # Remove points where there were no transition
+    forest_year = if_else(forest_year <= 0, NA_real_, forest_year),
+    agri_year = if_else(agri_year <= 0, NA_real_, agri_year),
+    trans_length = if_else(agri_year <= 0, NA_real_, trans_length),
+    trans_length = if_else(forest_year <= 0, NA_real_, trans_length)
+  ) %>%
+  mutate( # Fix dates to match the estimates
+    forest_year = forest_year,
+    agri_year = agri_year
+  ) %>%
+  pivot_longer(
+    cols =  c(forest_year, agri_year, trans_length),
+    names_to = "var",
+    values_to = "observed"
+  ) %>%
+  arrange(cell_id)
+
+## Load estimates data ----
+estimated <- open_dataset("data/trans_tabular_dataset/trans_length/")
+
+estimated <- estimated %>%
+  filter(
+    cell_id %in% observed$cell_id
+  ) %>%
+  collect() %>%
+  pivot_longer(
+    cols =  c(forest_year, agri_year, trans_length),
+    names_to = "var",
+    values_to = "estimated"
+  ) %>%
+  arrange(cell_id)
+
+# Check if there is any difference between tables
[…]

### 2022-11-08 scripts/figures/fig_5.R

@@ -1,0 +1,207 @@
+# HEADER ----------------------------------------------------------------------
+#
+# Title:        Validate transitions against visual inspection
+# Description:  The objective of this routine is compare the transition length
+#               results, comparing them with visual observations.
+#
+#
+# Author:       Hugo Tameirao Seixas
+# Contact:      tameirao.hugo@gmail.com
+# Date:         2022-08-03
+#
+# Notes:
+#
+#
+#
+#
+# LIBRARIES -------------------------------------------------------------------
+#
+library(arrow)
+library(magrittr)
+library(lubridate)
+library(glue)
+library(dplyr)
+library(units)
+library(tibble)
+library(readr)
+library(tidyr)
+library(forcats)
+library(purrr)
+library(scales)
+library(ggplot2)
+library(ggtext)
+#
+# OPTIONS ---------------------------------------------------------------------
+#
+#
+# LOAD DATA -------------------------------------------------------------------
+
+## Load observed data ----
+observed <- read_csv("data/validation/results.csv") %>%
+  mutate( # Remove points where there were no transition
+    forest_year = if_else(forest_year <= 0, NA_real_, forest_year),
+    agri_year = if_else(agri_year <= 0, NA_real_, agri_year),
+    trans_length = if_else(agri_year <= 0, NA_real_, trans_length),
+    trans_length = if_else(forest_year <= 0, NA_real_, trans_length)
+  ) %>%
+  mutate(across(forest_year:agri_year, ~ .x - 1984)) %>%
+  pivot_longer(
+    cols =  c(forest_year, agri_year, trans_length),
+    names_to = "var",
+    values_to = "observed"
+  ) %>%
+  arrange(cell_id)
+
+## Load estimates data ----
+estimated <- open_dataset("data/trans_tabular_dataset/trans_length/")
+
+estimated %<>%
+  filter(
+    cell_id %in% observed$cell_id
+  ) %>%
+  collect() %>%
+  mutate(across(forest_year:agri_year, ~ .x - 1984)) %>%
+  pivot_longer(
+    cols =  c(forest_year, agri_year, trans_length),
+    names_to = "var",
+    values_to = "estimated"
+  ) %>%
+  arrange(cell_id)
+
+# Check if there is any difference between tables
[…]

### 2022-11-09 manuscript/manuscript.Rmd
Modify methods. Add text to results.

@@
 Transitions were calculated using land use and land cover classification data from MapBiomas [@Souza2020], which ranges from 1985 to 2021, at a spatial resolution of approximately 30 meters.
+The data was filtered to contain only pixels that were occupied by forest and agriculture at some period, and that are not considered as water, according to the Global Surface Water product provided by Copernicus [@Pekel2016].
 
@@
 
-3.  Calculate the first year of "Forest Formation" LULC class after the year calculated in step 2, for each pixel;
+3.  Calculate the first year of "Forest Formation" LULC class after the year calculated in step 2;
 
-4.  Classify rows as "before" or "after" the occurrence of the year calculated in step 3, for each pixel;
+4.  Classify rows as "before" or "after" the occurrence of the year calculated in step 3;
 
-5.  Calculate the last year of "Forest Formation" within the rows classified as "before", for each pixel;
+5.  Calculate the last year of "Forest Formation" within the rows classified as "before", and add 1 year to represent the deforestation year;
 
@@
 
+The steps 2 to 7 are performed recursively to identify multiple transitions, in case they are present.
+Because of the amount of data, the processing and storage was performed in batches of data, structured in raster tiles.
+
 # Description of data collection
 
-We chose two ways to store and share the data.
+After the calculation of transitions, the results are stored in three different types of tables, organized in a folder structure and stored as Apache Parquet files.
 
--   One is to store the data from all three tables in three sets of Parquet files, in a structure of folders and sub folders;
+-   Tables that contains the spatial information (longitude, latitude) of each pixel, its unique id and the code of the municipality which contains the pixel.
+    It is named as "mask_cells";
 
--   The other is to convert the "trans_length.parquet" tables back to raster files.
-    Raster files are useful to efficiently store spatial data, calculate areas, operate in relation to vector objects and perform data summaries, and it is also a very accessible format;
+-   Tables with transition length values, the first and last year of the transition, the resulting agriculture type, and the number of the transition cycle.
+    They also contain the unique id of each pixel (related to the table above);
 
-The new raster files are created based on the tiles downloaded from GEE (which are in the path "./data/raw_raster_tiles/"), using their extent values and its number of rows and columns (located on the "./data/trans_tabular_dataset/tiles_metadata.parquet" file).
-For each tile file from "./data/raw_raster_tiles/", a set of new raster files are created based in the number of agriculture cycle ("agri_cycle") found in the tables, resulting in one raster for each agriculture cycle.
-Each new raster file contains 6 bands:
+-   Table with the LULC classes of all years within the transition, and also the first 5 years after the transition.
 
--   Band 1 = "agri_code" (The agriculture class code that resulted from the transition);
-
--   Band 2 = "forest_year" (The last forest year before the transition, can be interpreted as the deforestation year);
-
--   Band 3 = "agri_year" (The first year of the agriculture establishment, being the end of the transition);
-
--   Band 4 = "trans_length" (The length in years of the transition from forest to agriculture);
-
--   Band 5 = "trans_cycle" (The number of the transition cycle in which the transition occurred);
-
--   Band 6 = "forest_type" (The type of the forest that suffered the transition);
-
-The new raster files are saved in the path "./data/trans_raster_tiles/".
-
-To facilitate the use of these new raster files, we also merged all the raster of each agriculture cycle into a single mosaic, which are saved in the path "./data/trans_raster_mosaic/".
+The three tables are related to each other and can be used altogether, and are separated by tiles.
+Another table containing the metadata of each tile is also created, and holds the spatial characteristics of the tiles.
+With this spatial information, it is possible to convert the tabular data back to spatial raster, with identical spatial properties as the MapBiomas classification data.
 
@@
 
-Although we named as deforestation the last years identified as forest before classification of antropic cover, we acknowledge that it is not a direct measurement of deforestation (such as PRODES), it is however a proxy to deforestation.
+Although we named as deforestation the last years identified as forest before classification of anthropic cover, we acknowledge that it is not a direct measurement of deforestation (such as PRODES), it is however a proxy to deforestation.
 
@@
 
-```{r map-plot, out.width = "17cm", fig.pos = "h", echo = FALSE, fig.cap = "Map of distribution of transitions from forests to agriculture in the Brazillian Amazon biome.
+```{r map-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = "Map of distribution of transitions from forests to agriculture in the Brazillian Amazon biome.
 The hexagonal cells represent the most common transition length, and do not reflect the amount of area of transitions inside a cell.
@@
 
-```{r transbar-plot, out.width = "17cm", fig.pos = "h", echo = FALSE, fig.cap = "Transition area per year and transition length.
+```{r transbar-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = "Transition area per year and transition length.
 The bars represent the total amount of area at some state of the transition for each year.
 The color gradient in each bar represents the transition length related to a deforestation or a agriculture establishment event.
-Blue tones represents fast transitions (transitions closer to one year), pink tones represents slow transitions (transitions closer to 36 years).
-Transition events were separated by deforestation of primary forests (a) and secondary forests (c), and the subsequent  agriculture establishment of primary forests (b) and secondary forests (d).
-TODO: add labels for important years hwere public policies took place."}
+Blue tones represents fast transitions (transitions closer to 0 years), pink tones represents slow transitions (transitions closer to 35 years).
+Transition events were separated by deforestation of primary forests (a) and secondary forests (c), and the subsequent  agriculture establishment of primary forests (b) and secondary forests (d)."}
 
@@
 
-The deforestation area of primary forests increased largely from 1986 to 2003, which was followed be a steep decrease until 2008, when the deforested areas being converted to agriculture reached a stable rate.
+The deforestation area of primary forests increased largely from 1986 to 2003, which was followed be a steep decrease until 2009, when the deforested areas reached a stable rate.
 From 1986 to approximately 1995, most of deforested areas suffered a slow transition, mostly were higher than 10 years, after this period, fast transitions started to become more common, specially from 2002 to 2004.
@@
 Even after the decrease of deforestation after 2002, agriculture areas are expanding over lands where deforestation happened before 2002.
-However, after 2018, a sudden drop of agriculture establishment rate happened.
+However, after 2019, a sudden drop of agriculture establishment rate happened.
+
+The causes of deforestation and agriculture establishment in the Brazilian Amazon are complex and diverse.
+Political context, public policies, market prices and law enforcement can influence how these processes evolve over time.
+In 2004, the Brazilian Government launched the Action Plan for the Prevention and Control of Deforestation in the Legal Amazon (PPCDAm), which was composed by many initiatives to curb deforestation [@West2021].
+The PPCDAm was considered as a successful policy to slow deforestation rates in Brazil, with international recognition.
+Our calculations from MapBiomas data reinforces the corelation of the PPCDAm with the reduction of deforestation after 2004, and reduction of the agriculture establishment after 2005.
 
 In 2006, the Brazilian Association of Vegetable Oil Industries (BIOVE) and the National Association of Cereal Exporters (ANEC) committed to avoid commercialization of soy grains harvested from areas deforested after 2006.
-Our estimates of transitions show a decrease of deforested areas to be converted to agriculture after 2007, where it reached minimal values (Figure \@ref(fig:transbar-plot).a).
+Our estimates of transitions show a decrease of deforested areas to be converted to agriculture after 2008, where it reached minimal values (Figure \@ref(fig:transbar-plot).a).
 After 2006, agriculture establishment over deforested primary forests suffered a decrease, which stayed relatively stable until 2012, where a steep increase occurred, however, the new areas being occupied by agriculture were mainly over areas that were cleared more than a decade before (therefore, before 2006) (Figure \@ref(fig:transbar-plot).b) This shows that agriculture expansion did not halt after the soy moratorium, producers started expanding in old cleared areas.
@@
 
-```{r transridge-plot, out.width = "17cm", fig.pos = "h", echo = FALSE, fig.cap = "Transition length patterns inside states that belongs to the Amazon biome."}
+```{r transridge-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = "Transition length patterns inside states that belongs to the Amazon biome.
+Each year have a density estimate of the transition lengths, represented as colored curves.
+The peak of the curves represents transition length values with more frequency in one year of one state.
+Blue tones represents fast transitions (transitions closer to 0 years), pink tones represents slow transitions (transitions closer to 35 years)"}
 
@@
 The state of Amapá presented fast transitions along all the time series, where slow transitions are not as common.
-In contrast, Acre shows a majority of slow transitions, in which only 2021 showed more fast transitions.There are three states where the pattern of transitions length across time are alike, Mato Grosso, Pará and Rondônia presented more fast transitions from 1995 to 2005, after this period, slow transitions became more common with time, until around 2015, where the occurrence of fast transitions were the majority again.
+In contrast, Acre shows a majority of slow transitions, in which only 2021 showed more fast transitions.
+There are three states where the pattern of transitions length across time are alike, Mato Grosso, Pará and Rondônia presented more fast transitions from 1995 to 2005, after this period, slow transitions became more common with time, until around 2015, where the occurrence of fast transitions were the majority again.
 
@@
 
-```{r errorbar-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = "TODO: Add title."}
+```{r errorbar-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = " "}
 
@@
 
-```{r errorscatter-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = "TODO: Change line showing MAE to shaded areas showing quantiles of the errors. Add title."}
+```{r errorscatter-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = " "}
 

### 2022-11-09 manuscript/manuscript.tex
Modify methods. Add text to results.

@@
 Transitions were calculated using land use and land cover classification data from MapBiomas \citep{Souza2020}, which ranges from 1985 to 2021, at a spatial resolution of approximately 30 meters.
+The data was filtered to contain only pixels that were occupied by forest and agriculture at some period, and that are not considered as water, according to the Global Surface Water product provided by Copernicus \citep{Pekel2016}.
 
@@
 \item
-  Calculate the first year of ``Forest Formation'' LULC class after the year calculated in step 2, for each pixel;
+  Calculate the first year of ``Forest Formation'' LULC class after the year calculated in step 2;
 \item
-  Classify rows as ``before'' or ``after'' the occurrence of the year calculated in step 3, for each pixel;
+  Classify rows as ``before'' or ``after'' the occurrence of the year calculated in step 3;
 \item
-  Calculate the last year of ``Forest Formation'' within the rows classified as ``before'', for each pixel;
+  Calculate the last year of ``Forest Formation'' within the rows classified as ``before'', and add 1 year to represent the deforestation year;
 \item
@@
 
+The steps 2 to 7 are performed recursively to identify multiple transitions, in case they are present.
+Because of the amount of data, the processing and storage was performed in batches of data, structured in raster tiles.
+
 \section{Description of data collection}
 
-We chose two ways to store and share the data.
+After the calculation of transitions, the results are stored in three different types of tables, organized in a folder structure and stored as Apache Parquet files.
 
@@
 \item
-  One is to store the data from all three tables in three sets of Parquet files, in a structure of folders and sub folders;
+  Tables that contains the spatial information (longitude, latitude) of each pixel, its unique id and the code of the municipality which contains the pixel.
+  It is named as ``mask\_cells'';
 \item
-  The other is to convert the ``trans\_length.parquet'' tables back to raster files.
-  Raster files are useful to efficiently store spatial data, calculate areas, operate in relation to vector objects and perform data summaries, and it is also a very accessible format;
+  Tables with transition length values, the first and last year of the transition, the resulting agriculture type, and the number of the transition cycle.
+  They also contain the unique id of each pixel (related to the table above);
+\item
+  Table with the LULC classes of all years within the transition, and also the first 5 years after the transition.
 \end{itemize}
 
-The new raster files are created based on the tiles downloaded from GEE (which are in the path ``./data/raw\_raster\_tiles/''), using their extent values and its number of rows and columns (located on the ``./data/trans\_tabular\_dataset/tiles\_metadata.parquet'' file).
-For each tile file from ``./data/raw\_raster\_tiles/'', a set of new raster files are created based in the number of agriculture cycle (``agri\_cycle'') found in the tables, resulting in one raster for each agriculture cycle.
-Each new raster file contains 6 bands:
-
-\begin{itemize}
-\item
-  Band 1 = ``agri\_code'' (The agriculture class code that resulted from the transition);
-\item
-  Band 2 = ``forest\_year'' (The last forest year before the transition, can be interpreted as the deforestation year);
-\item
-  Band 3 = ``agri\_year'' (The first year of the agriculture establishment, being the end of the transition);
-\item
-  Band 4 = ``trans\_length'' (The length in years of the transition from forest to agriculture);
-\item
-  Band 5 = ``trans\_cycle'' (The number of the transition cycle in which the transition occurred);
-\item
-  Band 6 = ``forest\_type'' (The type of the forest that suffered the transition);
-\end{itemize}
-
-The new raster files are saved in the path ``./data/trans\_raster\_tiles/''.
-
-To facilitate the use of these new raster files, we also merged all the raster of each agriculture cycle into a single mosaic, which are saved in the path ``./data/trans\_raster\_mosaic/''.
+The three tables are related to each other and can be used altogether, and are separated by tiles.
+Another table containing the metadata of each tile is also created, and holds the spatial characteristics of the tiles.
+With this spatial information, it is possible to convert the tabular data back to spatial raster, with identical spatial properties as the MapBiomas classification data.
 
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
 \includegraphics[width=17cm]{figs/map} \caption{Map of distribution of transitions from forests to agriculture in the Brazillian Amazon biome.
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
 \includegraphics[width=17cm]{figs/trans_length_cols} \caption{Transition area per year and transition length.
@@
 The color gradient in each bar represents the transition length related to a deforestation or a agriculture establishment event.
-Blue tones represents fast transitions (transitions closer to one year), pink tones represents slow transitions (transitions closer to 36 years).
-Transition events were separated by deforestation of primary forests (a) and secondary forests (c), and the subsequent  agriculture establishment of primary forests (b) and secondary forests (d).
-TODO: add labels for important years hwere public policies took place.}\label{fig:transbar-plot}
+Blue tones represents fast transitions (transitions closer to 0 years), pink tones represents slow transitions (transitions closer to 35 years).
+Transition events were separated by deforestation of primary forests (a) and secondary forests (c), and the subsequent  agriculture establishment of primary forests (b) and secondary forests (d).}\label{fig:transbar-plot}
 \end{figure}
 
-The deforestation area of primary forests increased largely from 1985 to 2002, which was followed be a steep decrease until 2007, when the deforested areas being converted to agriculture reached a stable rate.
-From 1985 to approximately 1995, most of deforested areas suffered a slow transition, mostly were higher than 10 years, after this period, fast transitions started to become more common, specially from 2001 to 2003.
+The deforestation area of primary forests increased largely from 1986 to 2003, which was followed be a steep decrease until 2009, when the deforested areas reached a stable rate.
+From 1986 to approximately 1995, most of deforested areas suffered a slow transition, mostly were higher than 10 years, after this period, fast transitions started to become more common, specially from 2002 to 2004.
 
@@
 Even after the decrease of deforestation after 2002, agriculture areas are expanding over lands where deforestation happened before 2002.
-However, after 2018, a sudden drop of agriculture establishment rate happened.
+However, after 2019, a sudden drop of agriculture establishment rate happened.
+
+The causes of deforestation and agriculture establishment in the Brazilian Amazon are complex and diverse.
+Political context, public policies, market prices and law enforcement can influence how these processes evolve over time.
+In 2004, the Brazilian Government launched the Action Plan for the Prevention and Control of Deforestation in the Legal Amazon (PPCDAm), which was composed by many initiatives to curb deforestation \citep{West2021}.
+The PPCDAm was considered as a successful policy to slow deforestation rates in Brazil, with international recognition.
+Our calculations from MapBiomas data reinforces the corelation of the PPCDAm with the reduction of deforestation after 2004, and reduction of the agriculture establishment after 2005.
 
 In 2006, the Brazilian Association of Vegetable Oil Industries (BIOVE) and the National Association of Cereal Exporters (ANEC) committed to avoid commercialization of soy grains harvested from areas deforested after 2006.
-Our estimates of transitions show a decrease of deforested areas to be converted to agriculture after 2007, where it reached minimal values (Figure \ref{fig:transbar-plot}.a).
+Our estimates of transitions show a decrease of deforested areas to be converted to agriculture after 2008, where it reached minimal values (Figure \ref{fig:transbar-plot}.a).
 After 2006, agriculture establishment over deforested primary forests suffered a decrease, which stayed relatively stable until 2012, where a steep increase occurred, however, the new areas being occupied by agriculture were mainly over areas that were cleared more than a decade before (therefore, before 2006) (Figure \ref{fig:transbar-plot}.b) This shows that agriculture expansion did not halt after the soy moratorium, producers started expanding in old cleared areas.
@@
 
-\begin{figure}[h]
-\includegraphics[width=17cm]{figs/trans_ridge} \caption{Transition length patterns inside states that belongs to the Amazon biome.}\label{fig:transridge-plot}
+\begin{figure}[ht]
+\includegraphics[width=17cm]{figs/trans_ridge} \caption{Transition length patterns inside states that belongs to the Amazon biome.
+Each year have a density estimate of the transition lengths, represented as colored curves.
+The peak of the curves represents transition length values with more frequency in one year of one state.
+Blue tones represents fast transitions (transitions closer to 0 years), pink tones represents slow transitions (transitions closer to 35 years)}\label{fig:transridge-plot}
 \end{figure}
@@
 \begin{figure}[ht]
-\includegraphics[width=17cm]{figs/error_bars} \caption{TODO: Add title.}\label{fig:errorbar-plot}
+\includegraphics[width=17cm]{figs/error_bars} \caption{ }\label{fig:errorbar-plot}
 \end{figure}
@@
 \begin{figure}[ht]
-\includegraphics[width=17cm]{figs/error_scatter} \caption{TODO: Change line showing MAE to shaded areas showing quantiles of the errors. Add title.}\label{fig:errorscatter-plot}
+\includegraphics[width=17cm]{figs/error_scatter} \caption{ }\label{fig:errorscatter-plot}
 \end{figure}

### 2022-11-09 scripts/figures/fig_1.R
Change line width.

@@ -138,7 +138,7 @@ dist_map <- ggplot() +
   geom_sf(
     data = trans_hex,
     aes(fill = trans_length),
-    lwd = 0.1
+    linewidth = 0.05
   ) +
   ggtitle(
     "Distribution of transitions from forest to agriculture in the Amazon"
@@ -175,7 +175,7 @@ hr_map <- ggplot() +
   theme(
     legend.position = "",
     panel.background = element_rect(fill = "#747171"),
-    panel.border = element_rect(colour = "black", fill = NA, size = 0.5)
+    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.5)
   )
 
 hist_plot <- ggplot() +

### 2022-11-16 scripts/validation_app.R
Add results images.

@@ -50,9 +50,9 @@ if (!file_exists(glue("{proj_path}/data/validation/results.csv"))) {
 ui <- grid_page(
   theme = bslib::bs_theme(bootswatch = "slate"),
   layout = c(
-    "header  header      header       header     ",
-    "sidebar composite1  composite2   composite3 ",
-    "sidebar forest      agri         length     "
+    "header  header      header       header",
+    "sidebar composite1  composite2   composite3",
+    "sidebar forest      agri         length"
   ),
   row_sizes = c(
     "70px",
@@ -128,7 +128,7 @@ ui <- grid_page(
   grid_card_plot(area = "composite3"),
   grid_card_plot(area = "forest"),
   grid_card_plot(area = "agri"),
-  grid_card_plot(area = "lenght")
+  grid_card_plot(area = "length")
 )
 
 # Define server logic
@@ -295,6 +295,141 @@ server <-
         theme(
           plot.title = element_text(size = 14, face = "bold"),
           legend.position = ""
+        )
+
+    })
+
+    output$forest <- renderPlot({
+
+      bbox <-
+        rast(
+          dir_ls(
+            glue(
+              "{proj_path}/data/validation/landsat/{input$sampleCell}/"
+            ),
+            regexp = input$imageYear
+          )
+        )
+
+      img <-
+        crop(
+          rast(
+            glue(
+              "{proj_path}/data/trans_raster_mosaic/mb_mosaic_cycle_1.tif"
+            )
+          )[[2]],
+          bbox
+        )
+
+      img_df <- as.data.frame(img, xy = TRUE)
+
+      point <- sample_points %>%
+        filter(cell_id == input$sampleCell)
+
+      ggplot() +
+        geom_raster(
+          data = img_df,
+          aes(x = x, y = y, fill = forest_year)
+        ) +
+        geom_sf(
+          data = point,
+          alpha = 0.5
+        ) +
+        scale_fill_viridis_c(limits = c(1985, 2021)) +
+        theme_void() +
+        labs(title = "Forest Year") +
+        theme(
+          plot.title = element_text(size = 14, face = "bold")
+        )
+
+    })
+
+    output$agri <- renderPlot({
+
+      bbox <-
+        rast(
+          dir_ls(
+            glue(
+              "{proj_path}/data/validation/landsat/{input$sampleCell}/"
+            ),
+            regexp = input$imageYear
+          )
+        )
+
+      img <-
+        crop(
+          rast(
[…]

### 2022-11-16 manuscript/manuscript.Rmd
Write about accuracy assessment.

@@
 
+## Transition length calculation
+
 The estimations of transition from forest to agriculture were performed for the Amazon biome region, as defined by the Brazilian Institute of Geography and Statistics (IBGE), in 2019.
@@
 Because of the amount of data, the processing and storage was performed in batches of data, structured in raster tiles.
+
+## Accuracy assessment
+
+The accuracy assessment was performed by visual inspection of annual composites of Landsat images from MapBiomas.
+We selected 100 random points to be analysed.
+An area of approximately 4 squared kilometers around the sample point was used in the visual inspection of satellite images.
+
+The visual inspection used several variables derived from the Landsat historical collection.
+The median of Red, Green, Blue, Near Infrared (NIR) and Short Wave Infrared (SWIR1) from dry and wet season were used, and also the annual amplitude of the Normalized Difference Vegetation Index (NDVI).
+The process of accuracy assessment was performed in a Shiny app, and was conducted without any consultation to the transition length results.
+In the validation app, we estimated, by visual inspection, the year of deforestation and the year of agriculture establishment.
+The observed transition length was obtained by subtracting both dates.
+
+To evaluate the accuracy, we calculated the Mean Absolute Error (MAE), the Bias (BIAS), and the Percent Bias (PBIAS).
+We also analysed the results by plotting the errors as frequency bars, and scatter plots between observed and estimated values.
+
+After the completion of the analysis of the 100 sample points, we also conducted a qualitative assessment, where we compared our results with the satellite images.
 
@@
 In contrast, Acre shows a majority of slow transitions, in which only 2021 showed more fast transitions.
-There are three states where the pattern of transitions length across time are alike, Mato Grosso, Pará and Rondônia presented more fast transitions from 1995 to 2005, after this period, slow transitions became more common with time, until around 2015, where the occurrence of fast transitions were the majority again.
+There are three states where the pattern of transitions length across time are alike, Mato Grosso and Pará presented more fast transitions from 1995 to 2005, after this period, slow transitions became more common with time.
+Rondônia and Tocantins also presented a pattern in which transitions became slower with time, however this pattern started to occur earlier than Mato Grosso and Pará.
 
@@
 
-```{r errorbar-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = " "}
+```{r errorbar-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = "Bar with the count of error values (difference between observed and estimated values).
+Positive values indicate underestimation of the variable (estimations were lower than observations), negative values indicates overestimation (estimations higher than observations).
+Error metrics (Mean Absolute Error, Bias and Percent Bias) are displayed in the top right position of each box.
+The white deshed lines represents the MAE values of each variable."}
 
@@
 
-```{r errorscatter-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = " "}
+The dispersion of observed and estimated values shows no clear pattern of errors (Figure \@ref(fig:errorscatter-plot)).
+Transition length values are more concentrated in smaller values, which also shows higher errors (Figure \@ref(fig:errorscatter-plot).c).
+However, this is expected since faster transitions are more common (Figure \@ref(fig:map-plot)).
+
+```{r errorscatter-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = "Scatter plot between paired values of estimated and observed Deforestation and Agriculture Establishment years, and their respective Transition Length.
+Translucid white areas represent quantile ranges (5\\%, 25\\%, 75\\% and 95\\%) of the errors.
+The area between 5\\% and 95\\% quantiles includes 90\\% of points.
+The area between 25\\% and 75\\% quantiles includes 50\\% of points."}
 
@@
 
+According to accuracy assessment from MapBiomas, the collection 7 presents a global accuracy of 96.6% for the Amazon biome, which is the proportion of pixels that were classified correctly.
+For the Forest class, MapBiomas showed small errors of inclusion (proportion of pixels misclassified as other classes, but the real class were Forest), which fluctuated around 1%.
+The omission errors for Forests are also small (proportion of pixels misclassified as Forests, but the real class were not Forest), which fluctuated around 2%.
+The Agriculture class presented more errors, the inclusion errors ranged from 22% to 5%, and were mostly composed by Forests pixels (forest pixels misclassified as agriculture).
+The omission errors of Agriculture ranged from 22% to 8%, and were also mostly composed by Forests pixels (agriculture pixels misclassified as Agriculture).
+
 ## Qualitative assessment
 
+High heterogeneity inside agriculture plots (sample 56), caused by both year of deforestation or year of agriculture establishment.
+Roads being considered as agriculture (sample 2).
+Omissions were detected, specially agriculture plots not being considered as a transition as a whole, parts of a plot that were clearly a transition were not considered as a transition (sample 16, 30, 75).
+Inclusions were detected, areas that clearly showed no agriculture were considered as transitions (sample 10, 98).
+Effect of border on results (sample 25, 56, 84, 92).
+Areas with sparse forest vegetation presented large error of deforestation year (sample 33).
+
+Patterns of agriculture areas are roughly well represented (sample 25, 43, 56, 84).
+Some locations presents a better homogeneity of transitions inside agriculture plots (sample 84).
+
+Interference due to atmospheric conditions and lack of image availability clearly shaped spatial patterns of transitions at some locations.
+Quality of composites trough the time series affected accuracy of results.
+
 \conclusions[Conclusions]

### 2022-11-16 manuscript/manuscript.tex
Write about accuracy assessment.

@@
 
+\subsection{Transition length calculation}
+
 The estimations of transition from forest to agriculture were performed for the Amazon biome region, as defined by the Brazilian Institute of Geography and Statistics (IBGE), in 2019.
@@
 Because of the amount of data, the processing and storage was performed in batches of data, structured in raster tiles.
+
+\subsection{Accuracy assessment}
+
+The accuracy assessment was performed by visual inspection of annual composites of Landsat images from MapBiomas.
+We selected 100 random points to be analysed.
+An area of approximately 4 squared kilometers around the sample point was used in the visual inspection of satellite images.
+
+The visual inspection used several variables derived from the Landsat historical collection.
+The median of Red, Green, Blue, Near Infrared (NIR) and Short Wave Infrared (SWIR1) from dry and wet season were used, and also the annual amplitude of the Normalized Difference Vegetation Index (NDVI).
+The process of accuracy assessment was performed in a Shiny app, and was conducted without any consultation to the transition length results.
+In the validation app, we estimated, by visual inspection, the year of deforestation and the year of agriculture establishment.
+The observed transition length was obtained by subtracting both dates.
+
+To evaluate the accuracy, we calculated the Mean Absolute Error (MAE), the Bias (BIAS), and the Percent Bias (PBIAS).
+We also analysed the results by plotting the errors as frequency bars, and scatter plots between observed and estimated values.
+
+After the completion of the analysis of the 100 sample points, we also conducted a qualitative assessment, where we compared our results with the satellite images.
 
@@
 \begin{figure}[ht]
-\includegraphics[width=17cm]{figs/error_bars} \caption{ }\label{fig:errorbar-plot}
+\includegraphics[width=17cm]{figs/error_bars} \caption{Bar with the count of error values (difference between observed and estimated values).
+Positive values indicate underestimation of the variable (estimations were lower than observations), negative values indicates overestimation (estimations higher than observations).
+Error metrics (Mean Absolute Error, Bias and Percent Bias) are displayed in the top right position of each box.
+The white deshed lines represents the MAE values of each variable.}\label{fig:errorbar-plot}
 \end{figure}
 
+The dispersion of observed and estimated values shows no clear pattern of errors (Figure \ref{fig:errorscatter-plot}).
+Transition length values are more concentrated in smaller values, which also shows higher errors (Figure \ref{fig:errorscatter-plot}.c).
+However, this is expected since faster transitions are more common (Figure \ref{fig:map-plot}).
+
 \begin{figure}[ht]
-\includegraphics[width=17cm]{figs/error_scatter} \caption{ }\label{fig:errorscatter-plot}
+\includegraphics[width=17cm]{figs/error_scatter} \caption{Scatter plot between paired values of estimated and observed Deforestation and Agriculture Establishment years, and their respective Transition Length.
+Translucid white areas represent quantile ranges (5\%, 25\%, 75\% and 95\%) of the errors.
+The area between 5\% and 95\% quantiles includes 90\% of points.
+The area between 25\% and 75\% quantiles includes 50\% of points.}\label{fig:errorscatter-plot}
 \end{figure}
 
+According to accuracy assessment from MapBiomas, the collection 7 presents a global accuracy of 96.6\% for the Amazon biome, which is the proportion of pixels that were classified correctly.
+For the Forest class, MapBiomas showed small errors of inclusion (proportion of pixels misclassified as other classes, but the real class were Forest), which fluctuated around 1\%.
+The omission errors for Forests are also small (proportion of pixels misclassified as Forests, but the real class were not Forest), which fluctuated around 2\%.
+The Agriculture class presented more errors, the inclusion errors ranged from 22\% to 5\%, and were mostly composed by Forests pixels (forest pixels misclassified as agriculture).
+The omission errors of Agriculture ranged from 22\% to 8\%, and were also mostly composed by Forests pixels (agriculture pixels misclassified as Agriculture).
+
 \subsection{Qualitative assessment}
+
+High heterogeneity inside agriculture plots (sample 56), caused by both year of deforestation or year of agriculture establishment.
+Roads being considered as agriculture (sample 2).
+Omissions were detected, specially agriculture plots not being considered as a transition as a whole, parts of a plot that were clearly a transition were not considered as a transition (sample 16, 30, 75).
+Inclusions were detected, areas that clearly showed no agriculture were considered as transitions (sample 10, 98).
+Effect of border on results (sample 25, 56, 84, 92).
+Areas with sparse forest vegetation presented large error of deforestation year (sample 33).
+
+Patterns of agriculture areas are roughly well represented (sample 25, 43, 56, 84).
+Some locations presents a better homogeneity of transitions inside agriculture plots (sample 84).
+
+Interference due to atmospheric conditions and lack of image availability clearly shaped spatial patterns of transitions at some locations.
+Quality of composites trough the time series affected accuracy of results.
 

### 2022-11-16 manuscript/manuscript.Rmd
Add a paragraph in the validation results.

@@
 
+From the 100 random sample points used in the results validation, 1 point (sample 21) was not considered as a transition from our estimations, however, the visual inspection pointed to a likely event of transition from forest to agriculture.
+Also, there were 6 points (sample 6, 7, 55, 56, 78, 97) which the visual inspection did not found a transition from forest to pasture, although our estimations pointed as transitions.
+Therefore, 7% of the sample were completely misclassified by our estimates, and the rest of the accuracy assessment was performed over the remaining 93 sample points.
+
 When analyzing the errors from the transition length estimates, we observe that the year of deforestation shows the least amount of errors (Figure \@ref(fig:errorbar-plot)).

### 2022-11-16 manuscript/manuscript.tex
Add a paragraph in the validation results.

@@
 
+From the 100 random sample points used in the results validation, 1 point (sample 21) was not considered as a transition from our estimations, however, the visual inspection pointed to a likely event of transition from forest to agriculture.
+Also, there were 6 points (sample 6, 7, 55, 56, 78, 97) which the visual inspection did not found a transition from forest to pasture, although our estimations pointed as transitions.
+Therefore, 7\% of the sample were completely misclassified by our estimates, and the rest of the accuracy assessment was performed over the remaining 93 sample points.
+
 When analyzing the errors from the transition length estimates, we observe that the year of deforestation shows the least amount of errors (Figure \ref{fig:errorbar-plot}).

### 2022-11-16 manuscript/manuscript.Rmd
Improve and expand introduction. Add todo paragraphs in the results.

@@
 
-In the last decades, the Amazon biome have been submitted to strong changes, the advance of agricultural areas over native forests is shaping a new landscape in the region.
-The most common pattern of transitions from natural forest formations to agriculture sites, is the initial process of deforestation, followed by the establishment of pasture, where extensive livestock production takes place, until its replacement to agricultural activities.
+Brazil policies since the 1960s were focused on the expansion of the frontier in the Amazon, specially focused in economic income and land security [@Carvalho2002; @Mcdonald2003].
+Such development model led to the construction of thousands of kilometers of roads, and the settlement of large scale livestock farms [@Carvalho2002].
+This period was marked with high rates of deforestation of mature forests [@Fearnside2005].
+At the same time, cattle and soybean productions started to move to the north direction in the Brazilian territory, in which soybeans cultivation areas followed cattle ranches expansion over the Amazon forest [@Simon2005].
+This pattern of extensive livestock occupation after deforestation, followed by the establishment of annual crops is very characteristics in the Amazon biome, and persists nowadays.
 This process can take several decades to be accomplished, or even less than one year, in which it may be considered as a direct transition from forest to agriculture.
@@
 
+TODO: Talk about the Avança Brasil plan [@Carvalho2002].
+
 The deforestation area of primary forests increased largely from 1986 to 2003, which was followed be a steep decrease until 2009, when the deforested areas reached a stable rate.
@@
 Rondônia and Tocantins also presented a pattern in which transitions became slower with time, however this pattern started to occur earlier than Mato Grosso and Pará.
+
+TODO: Talk about the percentage of each agriculture type in the total transitions.
+And show numbers about what happens 5 years after transitions for each crop type.
 

### 2022-11-16 manuscript/manuscript.tex
Improve and expand introduction. Add todo paragraphs in the results.

@@
 
-In the last decades, the Amazon biome have been submitted to strong changes, the advance of agricultural areas over native forests is shaping a new landscape in the region.
-The most common pattern of transitions from natural forest formations to agriculture sites, is the initial process of deforestation, followed by the establishment of pasture, where extensive livestock production takes place, until its replacement to agricultural activities.
+Brazil policies since the 1960s were focused on the expansion of the frontier in the Amazon, specially focused in economic income and land security \citep{Carvalho2002, Mcdonald2003}.
+Such development model led to the construction of thousands of kilometers of roads, and the settlement of large scale livestock farms \citep{Carvalho2002}.
+This period was marked with high rates of deforestation of mature forests \citep{Fearnside2005}.
+At the same time, cattle and soybean productions started to move to the north direction in the Brazilian territory, in which soybeans cultivation areas followed cattle ranches expansion over the Amazon forest \citep{Simon2005}.
+This pattern of extensive livestock occupation after deforestation, followed by the establishment of annual crops is very characteristics in the Amazon biome, and persists nowadays.
 This process can take several decades to be accomplished, or even less than one year, in which it may be considered as a direct transition from forest to agriculture.
@@
 
+TODO: Talk about the Avança Brasil plan \citep{Carvalho2002}.
+
 The deforestation area of primary forests increased largely from 1986 to 2003, which was followed be a steep decrease until 2009, when the deforested areas reached a stable rate.

### 2022-11-16 manuscript/manuscript.Rmd
Fix dates in the results.

@@
 The transition calculations shows that between 1985 and 2021, 64874 squared kilometers of forests were converted to agriculture, in the Brazilian Amazon biome.
-The length of the transitions can go from 1 to 36 years, in which transitions closer to 1 year are considered as fast transitions, and transitions closer to 36 years are considered as slow transitions.
-Transitions of 1 year are considered as "direct" transitions, where there were no presence of pasture before the establishment of agriculture, out estimations shows that around 9.2 % of the transitions are considered as "direct".
+The length of the transitions can go from 0 to 35 years, in which transitions closer to 0 year are considered as fast transitions, and transitions closer to 35 years are considered as slow transitions.
+Transitions of 0 year are considered as "direct" transitions, where there were no presence of pasture before the establishment of agriculture, our estimations shows that around 9.2 % of the transitions are considered as "direct".
 
-Although we named as deforestation the last years identified as forest before classification of anthropic cover, we acknowledge that it is not a direct measurement of deforestation (such as PRODES), it is however a proxy to deforestation.
+Although we named as deforestation the last years identified as forest before classification of antropic cover, we acknowledge that it is not a direct measurement of deforestation (such as PRODES), it is however a proxy to deforestation.
 

### 2022-11-16 manuscript/manuscript.tex
Fix dates in the results.

@@
 The transition calculations shows that between 1985 and 2021, 64874 squared kilometers of forests were converted to agriculture, in the Brazilian Amazon biome.
-The length of the transitions can go from 1 to 36 years, in which transitions closer to 1 year are considered as fast transitions, and transitions closer to 36 years are considered as slow transitions.
-Transitions of 1 year are considered as ``direct'' transitions, where there were no presence of pasture before the establishment of agriculture, out estimations shows that around 9.2 \% of the transitions are considered as ``direct''.
+The length of the transitions can go from 0 to 35 years, in which transitions closer to 0 year are considered as fast transitions, and transitions closer to 35 years are considered as slow transitions.
+Transitions of 0 year are considered as ``direct'' transitions, where there were no presence of pasture before the establishment of agriculture, our estimations shows that around 9.2 \% of the transitions are considered as ``direct''.
 
-Although we named as deforestation the last years identified as forest before classification of anthropic cover, we acknowledge that it is not a direct measurement of deforestation (such as PRODES), it is however a proxy to deforestation.
+Although we named as deforestation the last years identified as forest before classification of antropic cover, we acknowledge that it is not a direct measurement of deforestation (such as PRODES), it is however a proxy to deforestation.
 

### 2022-11-17 manuscript/manuscript.Rmd
Add reference. Add TODO. Lining format was changed.

@@
 
-Brazil policies since the 1960s were focused on the expansion of the frontier in the Amazon, specially focused in economic income and land security [@Carvalho2002; @Mcdonald2003].
-Such development model led to the construction of thousands of kilometers of roads, and the settlement of large scale livestock farms [@Carvalho2002].
+Brazil policies since the 1960s were focused on the expansion of the frontier in the Amazon, specially focused in economic income and land security [@Carvalho2002; @Mcdonald2003; @Banerjee2009].
+Such development model led to the construction of thousands of kilometers of roads, and the settlement of large scale livestock farms [@Carvalho2002; @Banerjee2009].
 This period was marked with high rates of deforestation of mature forests [@Fearnside2005].
@@
 
--   Tables that contains the spatial information (longitude, latitude) of each pixel, its unique id and the code of the municipality which contains the pixel.
-    It is named as "mask_cells";
+-   Tables that contains the spatial information (longitude, latitude) of each pixel, its unique id and the code of the municipality which contains the pixel. It is named as "mask_cells";
 
@@
 
-TODO: Talk about the Avança Brasil plan [@Carvalho2002].
+TODO: Talk about public policies from 1985 to 2000 (greater carajas program, national environmental policy) [@Banerjee2009].
+
+TODO: Talk about the Avança Brasil plan [@Carvalho2002; @Banerjee2009].
+
+TODO: confront our results with Morton 2006 (see the year of 2003).
 
@@
 
-TODO: Talk about the percentage of each agriculture type in the total transitions.
-And show numbers about what happens 5 years after transitions for each crop type.
+TODO: Talk about the percentage of each agriculture type in the total transitions. And show numbers about what happens 5 years after transitions for each crop type.
 
@@
 
-Patterns of agriculture areas are roughly well represented (sample 25, 43, 56, 84).
-Some locations presents a better homogeneity of transitions inside agriculture plots (sample 84).
+Patterns of agriculture areas are roughly well represented (sample 25, 43, 56, 84). Some locations presents a better homogeneity of transitions inside agriculture plots (sample 84).
 

### 2022-11-21 manuscript/manuscript.Rmd
Add some references and TODOs in the results.

@@
 This period was marked with high rates of deforestation of mature forests [@Fearnside2005].
-At the same time, cattle and soybean productions started to move to the north direction in the Brazilian territory, in which soybeans cultivation areas followed cattle ranches expansion over the Amazon forest [@Simon2005].
-This pattern of extensive livestock occupation after deforestation, followed by the establishment of annual crops is very characteristics in the Amazon biome, and persists nowadays.
+At the same time, cattle and soybean productions started to move to the north direction in the Brazilian territory, in which soybeans cultivation areas followed cattle ranches expansion over the Amazon forest [@Simon2005; @Barona2010; @Arima2011].
+This pattern of extensive livestock occupation after deforestation, followed by the establishment of annual crops is very characteristic in the Amazon biome, and persists nowadays.
 This process can take several decades to be accomplished, or even less than one year, in which it may be considered as a direct transition from forest to agriculture.
@@
 TODO: confront our results with Morton 2006 (see the year of 2003).
+Confront results from [@Macedo2012], talking about how the expansion of soybean areas are still responsible for production growth, because of very slow transitions.
+
+TODO: talk that even with soy moratorium, and that the deforestation related to transitions were smaller, the deforestation related to pasture occupation kept going, with indirect effects of soybean expansion (Arima 2011).
 
@@
 Quality of composites trough the time series affected accuracy of results.
+Regions of the North part of the Amazon presents lower quality than regions in the South of the Amazon BIome.
 

### 2022-11-22 scripts/figures/table_1.R
Add script to generate table.

@@ -1,0 +1,135 @@
+# HEADER ----------------------------------------------------------------------
+#
+# Title:        Create figure 2
+# Description:  The objective of this routine is to explore the dataset by
+#               filtering and aggregating data in Spark. Results are supposed
+#               to give us a overall view of the data by plots and tables.
+#
+# Author:       Hugo Tameirao Seixas
+# Contact:      tameirao.hugo@gmail.com
+# Date:         2020-11-10
+#
+# Notes:        In order to run this routine, you will need to successfully
+#               install and connect to Spark using the {sparklyr} package.
+#               Installation and use guide can be found in:
+#               https://github.com/sparklyr/sparklyr
+#
+# LIBRARIES -------------------------------------------------------------------
+#
+library(sparklyr)
+library(magrittr)
+library(lubridate)
+library(scales)
+library(dplyr)
+library(tibble)
+library(readr)
+library(tidyr)
+library(forcats)
+library(ggplot2)
+library(ggtext)
+library(scico)
+library(gt)
+#
+# OPTIONS ---------------------------------------------------------------------
+#
+source('conf/spark_config.R')
+#
+# SET LULC NAMES --------------------------------------------------------------
+
+## Get the names for each LULC class from MapBiomas ----
+mb_dict <- read_csv("data/mb_class_dictionary.csv") %>%
+  select(class_code, class_name) %>%
+  rename(agri_code = class_code)
+
+# LOAD DATASETS ---------------------------------------------------------------
+
+## Connect with Spark -----
+sc <- spark_connect(master = "local", config = config, version = "3.3.0")
+
+## Load trans_length table ----
+trans_length <-
+  spark_read_parquet(
+    sc,
+    name = "trans_length",
+    path = "data/trans_tabular_dataset/trans_length/",
+    memory = FALSE
+  )
+
+## Load transition time series table ----
+trans_classes <-
+  spark_read_parquet(
+    sc,
+    name = "trans_classes",
+    path = "data/trans_tabular_dataset/trans_classes/",
+    memory = FALSE
+  )
+
+# FILTER AND AGGREGATE DATA ---------------------------------------------------
+
[…]

### 2022-11-22 manuscript/manuscript.Rmd
Add text to the results.

@@
 
--   Tables that contains the spatial information (longitude, latitude) of each pixel, its unique id and the code of the municipality which contains the pixel. It is named as "mask_cells";
+-   Tables that contains the spatial information (longitude, latitude) of each pixel, its unique id and the code of the municipality which contains the pixel.
+    It is named as "mask_cells";
 
@@
 
-TODO: Talk about the percentage of each agriculture type in the total transitions. And show numbers about what happens 5 years after transitions for each crop type.
+The great majority of transitions are from forests to Soybeans and "Other Temporary Crops", which represents 95% of the transitions in the Amazon biome (14% of Soybean and 81% of "Other Temporary Crops").
+When analyzing what happened after 5 years since the transition, Soybean farms are persistent, 80% of these areas remained as Soybean, 14% is converted to "Other Temporary Crops" and 7% in occupied by Pasture.
+"Other Temporary Crops" are less persistent, 27% of these areas remains as the same use, 56% is converted to Soybean, and 15% to Pasture.
+Sugar Cane, Cotton and Perennial Crops also appears, but in a negligible proportions.
+
+TODO: Talk about the percentage of each agriculture type in the total transitions.
+And show numbers about what happens 5 years after transitions for each crop type.
 
@@
 
-Patterns of agriculture areas are roughly well represented (sample 25, 43, 56, 84). Some locations presents a better homogeneity of transitions inside agriculture plots (sample 84).
+Patterns of agriculture areas are roughly well represented (sample 25, 43, 56, 84).
+Some locations presents a better homogeneity of transitions inside agriculture plots (sample 84).
 
@@
 Quality of composites trough the time series affected accuracy of results.
-Regions of the North part of the Amazon presents lower quality than regions in the South of the Amazon BIome.
+Regions of the North part of the Amazon presents lower quality than regions in the South of the Amazon biome.
 

### 2022-11-22 manuscript/manuscript.tex
Add text to the results.

@@
 
-Brazil policies since the 1960s were focused on the expansion of the frontier in the Amazon, specially focused in economic income and land security \citep{Carvalho2002, Mcdonald2003}.
-Such development model led to the construction of thousands of kilometers of roads, and the settlement of large scale livestock farms \citep{Carvalho2002}.
+Brazil policies since the 1960s were focused on the expansion of the frontier in the Amazon, specially focused in economic income and land security \citep{Carvalho2002, Mcdonald2003, Banerjee2009}.
+Such development model led to the construction of thousands of kilometers of roads, and the settlement of large scale livestock farms \citep{Carvalho2002, Banerjee2009}.
 This period was marked with high rates of deforestation of mature forests \citep{Fearnside2005}.
-At the same time, cattle and soybean productions started to move to the north direction in the Brazilian territory, in which soybeans cultivation areas followed cattle ranches expansion over the Amazon forest \citep{Simon2005}.
-This pattern of extensive livestock occupation after deforestation, followed by the establishment of annual crops is very characteristics in the Amazon biome, and persists nowadays.
+At the same time, cattle and soybean productions started to move to the north direction in the Brazilian territory, in which soybeans cultivation areas followed cattle ranches expansion over the Amazon forest \citep{Simon2005, Barona2010, Arima2011}.
+This pattern of extensive livestock occupation after deforestation, followed by the establishment of annual crops is very characteristic in the Amazon biome, and persists nowadays.
 This process can take several decades to be accomplished, or even less than one year, in which it may be considered as a direct transition from forest to agriculture.
@@
 
-TODO: Talk about the Avança Brasil plan \citep{Carvalho2002}.
+TODO: Talk about public policies from 1985 to 2000 (greater carajas program, national environmental policy) \citep{Banerjee2009}.
+
+TODO: Talk about the Avança Brasil plan \citep{Carvalho2002, Banerjee2009}.
+
+TODO: confront our results with Morton 2006 (see the year of 2003).
+Confront results from \citep{Macedo2012}, talking about how the expansion of soybean areas are still responsible for production growth, because of very slow transitions.
+
+TODO: talk that even with soy moratorium, and that the deforestation related to transitions were smaller, the deforestation related to pasture occupation kept going, with indirect effects of soybean expansion (Arima 2011).
 
@@
 Rondônia and Tocantins also presented a pattern in which transitions became slower with time, however this pattern started to occur earlier than Mato Grosso and Pará.
+
+The great majority of transitions are from forests to Soybeans and ``Other Temporary Crops'', which represents 95\% of the transitions in the Amazon biome (14\% of Soybean and 81\% of ``Other Temporary Crops'').
+When analyzing what happened after 5 years since the transition, Soybean farms are persistent, 80\% of these areas remained as Soybean, 14\% is converted to ``Other Temporary Crops'' and 7\% in occupied by Pasture.
+``Other Temporary Crops'' are less persistent, 27\% of these areas remains as the same use, 56\% is converted to Soybean, and 15\% to Pasture.
+Sugar Cane, Cotton and Perennial Crops also appears, but in a negligible proportions.
 
@@
 Quality of composites trough the time series affected accuracy of results.
+Regions of the North part of the Amazon presents lower quality than regions in the South of the Amazon biome.
 

### 2022-11-22 manuscript/manuscript.Rmd
Remove TODO.
Add text to the introduction.

@@
 
-Land cover transitions can cause important impacts on ecosystem properties [@Nunes2022].
+Land cover transitions are known to cause impacts on different scales.
+Deforestation and land use changes can affect hidrological processes in large basins [@Arias2018]; reduce the forest resilience to extreme events and other sources of perturbation [@Boulton2022]; deforestation and farming practices can reduce convective rainfall and increase surface temperature [@Maeda2021]; increase carbon emissions [@Gatti2021]; cause important impacts on fauna and flora diversity, soil properties and carbon stocks [@Nunes2022]; and even negatively impact public health [@Ellwanger2020].
 
-This project aims to characterize and quantify the length of the transitions from forest formations to agriculture in the Amazon.
+Data on land use and land cover (LULC) classification are evolving rapidly in the last decade.
+Initiatives such as MapBiomas [@Souza2020], launched in 2015, which provides high resolution, annual classifications for the Brazilian territory.
+Such type of data are essential to analyse LULC change processes, make causal inferences, and understand what drives and what are the impacts of the transformation of the landscape in Brazil.
+In this context, this project aims to characterize and quantify the length of the transitions from forest formations to agriculture in the Amazon.
+The availability of this data can be useful to the development of interdisciplinary research involved with LULC changes.
 
@@
 The estimations of transition from forest to agriculture were performed for the Amazon biome region, as defined by the Brazilian Institute of Geography and Statistics (IBGE), in 2019.
-Transitions were calculated using land use and land cover classification data from MapBiomas [@Souza2020], which ranges from 1985 to 2021, at a spatial resolution of approximately 30 meters.
+Transitions were calculated using land use and land cover classification data from MapBiomas, which ranges from 1985 to 2021, at a spatial resolution of approximately 30 meters.
 The data was filtered to contain only pixels that were occupied by forest and agriculture at some period, and that are not considered as water, according to the Global Surface Water product provided by Copernicus [@Pekel2016].
@@
 The great majority of transitions are from forests to Soybeans and "Other Temporary Crops", which represents 95% of the transitions in the Amazon biome (14% of Soybean and 81% of "Other Temporary Crops").
-When analyzing what happened after 5 years since the transition, Soybean farms are persistent, 80% of these areas remained as Soybean, 14% is converted to "Other Temporary Crops" and 7% in occupied by Pasture.
+When analyzing what happened after 5 years since the transition, Soybean farms are persistent, 79% of these areas remained as Soybean, 14% is converted to "Other Temporary Crops" and 7% in occupied by Pasture.
 "Other Temporary Crops" are less persistent, 27% of these areas remains as the same use, 56% is converted to Soybean, and 15% to Pasture.
+Conversions from Soybean and "Other Temporary Uses" to other covers and uses (apart from the already cited) are negligible.
 Sugar Cane, Cotton and Perennial Crops also appears, but in a negligible proportions.
-
-TODO: Talk about the percentage of each agriculture type in the total transitions.
-And show numbers about what happens 5 years after transitions for each crop type.
 

### 2022-11-22 manuscript/manuscript.tex
Add text to the introduction.

@@
 
-Land cover transitions can cause important impacts on ecosystem properties \citep{Nunes2022}.
+Land cover transitions are known to cause impacts on different scales.
+Deforestation and land use changes can affect hidrological processes in large basins \citep{Arias2018}; reduce the forest resilience to extreme events and other sources of perturbation \citep{Boulton2022}; deforestation and farming practices can reduce convective rainfall and increase surface temperature \citep{Maeda2021}; increase carbon emissions \citep{Gatti2021}; cause important impacts on fauna and flora diversity, soil properties and carbon stocks \citep{Nunes2022}; and even negatively impact public health \citep{Ellwanger2020}.
 
-This project aims to characterize and quantify the length of the transitions from forest formations to agriculture in the Amazon.
+Data on land use and land cover (LULC) classification are evolving rapidly in the last decade.
+Initiatives such as MapBiomas \citep{Souza2020}, launched in 2015, which provides high resolution, annual classifications for the Brazilian territory.
+Such type of data are essential to analyse LULC change processes, make causal inferences, and understand what drives and what are the impacts of the transformation of the landscape in Brazil.
+In this context, this project aims to characterize and quantify the length of the transitions from forest formations to agriculture in the Amazon.
+The availability of this data can be useful to the development of interdisciplinary research involved with LULC changes.
 
@@
 The estimations of transition from forest to agriculture were performed for the Amazon biome region, as defined by the Brazilian Institute of Geography and Statistics (IBGE), in 2019.
-Transitions were calculated using land use and land cover classification data from MapBiomas \citep{Souza2020}, which ranges from 1985 to 2021, at a spatial resolution of approximately 30 meters.
+Transitions were calculated using land use and land cover classification data from MapBiomas, which ranges from 1985 to 2021, at a spatial resolution of approximately 30 meters.
 The data was filtered to contain only pixels that were occupied by forest and agriculture at some period, and that are not considered as water, according to the Global Surface Water product provided by Copernicus \citep{Pekel2016}.
@@
 The great majority of transitions are from forests to Soybeans and ``Other Temporary Crops'', which represents 95\% of the transitions in the Amazon biome (14\% of Soybean and 81\% of ``Other Temporary Crops'').
-When analyzing what happened after 5 years since the transition, Soybean farms are persistent, 80\% of these areas remained as Soybean, 14\% is converted to ``Other Temporary Crops'' and 7\% in occupied by Pasture.
+When analyzing what happened after 5 years since the transition, Soybean farms are persistent, 79\% of these areas remained as Soybean, 14\% is converted to ``Other Temporary Crops'' and 7\% in occupied by Pasture.
 ``Other Temporary Crops'' are less persistent, 27\% of these areas remains as the same use, 56\% is converted to Soybean, and 15\% to Pasture.
+Conversions from Soybean and ``Other Temporary Uses'' to other covers and uses (apart from the already cited) are negligible.
 Sugar Cane, Cotton and Perennial Crops also appears, but in a negligible proportions.
-
-TODO: Talk about the percentage of each agriculture type in the total transitions.
-And show numbers about what happens 5 years after transitions for each crop type.
 

### 2022-11-22 manuscript/manuscript.Rmd
Add text to the abstract.

@@
 abstract: |
+  Land use and land cover changes in the Amazon biome are a main process that influences many aspects at local, national, and global levels.
+A large quantity of studies already reliad on land use classification to analyse this process, drivers and impacts.
 A new dataset was created by calculating the time necessary for deforested forests to transition to agriculture in the Brazillian Amazon biome.

### 2022-11-22 manuscript/manuscript.tex
Add text to the abstract.

@@
 \begin{abstract}
+Land use and land cover changes in the Amazon biome are a main process that influences many aspects at local, national, and global levels.
+A large quantity of studies already reliad on land use classification to analyse this process, drivers and impacts.
 A new dataset was created by calculating the time necessary for deforested forests to transition to agriculture in the Brazillian Amazon biome.

### 2022-11-22 manuscript/manuscript.Rmd
Add text to methods.

@@
 The estimations of transition from forest to agriculture were performed for the Amazon biome region, as defined by the Brazilian Institute of Geography and Statistics (IBGE), in 2019.
-Transitions were calculated using land use and land cover classification data from MapBiomas, which ranges from 1985 to 2021, at a spatial resolution of approximately 30 meters.
-The data was filtered to contain only pixels that were occupied by forest and agriculture at some period, and that are not considered as water, according to the Global Surface Water product provided by Copernicus [@Pekel2016].
+Transitions were calculated using land use and land cover classification data from MapBiomas, collection 7, which ranges from 1985 to 2021, at a spatial resolution of approximately 30 meters.
+
+The MapBiomas data was downloaded from Google Earth Engine platform [@Gorelick2017].
+The pixels of the Amazon biome were filtered to contain only those that were occupied by forest and agriculture at some period, and that are not considered as water, according to the Global Surface Water product provided by Copernicus [@Pekel2016].
+For the download, the data was divided in tiles, to allow local processing of a large quantity of pixels.Two sets of data were downloaded with the same spatial dimensions, one with the LULC classifications for each year, and another with binary values, representing whether that pixels was valid for processing or not.
+
+The first part of the processing was to create a table from the binary values data, and create unique identification for each pixel, their coordinates, calculate the area, and also address the municipality in which each pixel was inside.
+In this part, a table was also generated with the spatial information of each tile, so that it is possible to convert the generated tables back to raster format.
 

### 2022-11-22 manuscript/manuscript.tex
Add text to methods.

@@
 The estimations of transition from forest to agriculture were performed for the Amazon biome region, as defined by the Brazilian Institute of Geography and Statistics (IBGE), in 2019.
-Transitions were calculated using land use and land cover classification data from MapBiomas, which ranges from 1985 to 2021, at a spatial resolution of approximately 30 meters.
-The data was filtered to contain only pixels that were occupied by forest and agriculture at some period, and that are not considered as water, according to the Global Surface Water product provided by Copernicus \citep{Pekel2016}.
+Transitions were calculated using land use and land cover classification data from MapBiomas, collection 7, which ranges from 1985 to 2021, at a spatial resolution of approximately 30 meters.
+
+The MapBiomas data was downloaded from Google Earth Engine platform \citep{Gorelick2017}.
+The pixels of the Amazon biome were filtered to contain only those that were occupied by forest and agriculture at some period, and that are not considered as water, according to the Global Surface Water product provided by Copernicus \citep{Pekel2016}.
+For the download, the data was divided in tiles, to allow local processing of a large quantity of pixels.Two sets of data were downloaded with the same spatial dimensions, one with the LULC classifications for each year, and another with binary values, representing whether that pixels was valid for processing or not.
+
+The first part of the processing was to create a table from the binary values data, and create unique identification for each pixel, their coordinates, calculate the area, and also address the municipality in which each pixel was inside.
+In this part, a table was also generated with the spatial information of each tile, so that it is possible to convert the generated tables back to raster format.
 

### 2022-11-23 manuscript/manuscript.Rmd
Add text to the methods and the results.

@@
 
-Transition length was calculated pixel by pixel, by performing the following steps:
+The second part of the processing was to load each tile separately, convert it to a table and calculate every possible transition from forest to agriculture.
+The transitions were calculated pixel by pixel, by performing the following steps:
 
@@
 
-2.  Calculate the year of first occurrence of a non-natural LULC class for each pixel;
+2.  Calculate the year of first occurrence of any agriculture class for each pixel;
 
-3.  Calculate the first year of "Forest Formation" LULC class after the year calculated in step 2;
+3.  Calculate the first year of "Forest Formation" LULC class after the year calculated in step 2.
+    This step identifies if here are more than one transition in a pixel;
 
-4.  Classify rows as "before" or "after" the occurrence of the year calculated in step 3;
+4.  Classify rows as "before" or "after" the occurrence of the year calculated in step 3.
+    This is the identification of periods before and after the recurrence of forest after a first transition to agriculture;
 
@@
 The steps 2 to 7 are performed recursively to identify multiple transitions, in case they are present.
-Because of the amount of data, the processing and storage was performed in batches of data, structured in raster tiles.
+In addition to the transition length values and the years of the transition, transitions are also qualified by the type of agriculture that were established after deforestation, if the transition occurred from primary or secondary forests, and the number of recurrence of transitions in a single pixel.
+The LULC classes within the transition and until 5 years after the transitions were also stored.
+
+The transition results are stored in a data set of tabular files, where data can be queried for further analysis.
+After the calculations, data was also converted back to raster format, with the aim to expand the accessibility of the data set.
 
@@
 
-The accuracy assessment was performed by visual inspection of annual composites of Landsat images from MapBiomas.
+An accuracy assessment was performed by visual inspection of annual composites of Landsat images from MapBiomas.
 We selected 100 random points to be analysed.
@@
 
+Our transitions calculations found pixels that presented up to six transitions from forest to agriculture since 1985, as these numbers are unlikely to happen (they are distributed as sparse pixels, and do not show patterns of real transitions, being common at borders), we proceeded the results analysis only in the first transitions found in each pixel.
+
 Although we named as deforestation the last years identified as forest before classification of antropic cover, we acknowledge that it is not a direct measurement of deforestation (such as PRODES), it is however a proxy to deforestation.
@@
 
-TODO: Talk about public policies from 1985 to 2000 (greater carajas program, national environmental policy) [@Banerjee2009].
+The deforestation area of primary forests increased largely from 1986 to 2003, which was followed be a steep decrease until 2009, when the deforested areas reached a stable rate (Figure \@ref(fig:transbar-plot).a).
+From 1986 to approximately 1995, most of deforested areas suffered a slow transition, mostly were higher than 10 years, after this period, fast transitions started to become more common, specially from 2002 to 2004 (Figure \@ref(fig:transbar-plot).a).
+The peak of deforestation in 2003, according to our results, shows that 12.6% of the transitions were a direct conversion from forest to agriculture, estimates by @Morton2006 point to a proportion of 23%.
+If we include fast transitions (from 0 to 2 years) from our results, the proportion jumps to 48.5 % of all the transitions.
 
-TODO: Talk about the Avança Brasil plan [@Carvalho2002; @Banerjee2009].
+Agriculture establishment over areas of primary forests peaked in 2005 and 2013 (Figure \@ref(fig:transbar-plot).b).
+Despite similar rates between both years, their transition lengths differ greatly, in 2005 most of the transitions were faster than 10 years, while in 2013 the great majority of transitions were slower than 10 years (Figure \@ref(fig:transbar-plot).b).
+The year of 2003 marked a change in the transition length of establishment of agriculture areas, after this year, most of the transitions happened in areas deforested at least 10 years before (Figure \@ref(fig:transbar-plot).b).
+Even after the decrease of deforestation after 2002, agriculture areas are expanding over lands where deforestation happened before 2002.
+However, after 2019, a sudden drop of agriculture establishment rate happened (Figure \@ref(fig:transbar-plot).b).
 
-TODO: confront our results with Morton 2006 (see the year of 2003).
-Confront results from [@Macedo2012], talking about how the expansion of soybean areas are still responsible for production growth, because of very slow transitions.
-
-TODO: talk that even with soy moratorium, and that the deforestation related to transitions were smaller, the deforestation related to pasture occupation kept going, with indirect effects of soybean expansion (Arima 2011).
-
-The deforestation area of primary forests increased largely from 1986 to 2003, which was followed be a steep decrease until 2009, when the deforested areas reached a stable rate.
-From 1986 to approximately 1995, most of deforested areas suffered a slow transition, mostly were higher than 10 years, after this period, fast transitions started to become more common, specially from 2002 to 2004.
-
-Agriculture establishment over areas of primary forests peaked in 2005 and 2013.
-Despite similar rates between both years, their transition lengths differ greatly, in 2005 most of the transitions were faster than 10 years, while in 2013 the great majority of transitions were slower than 10 years.
-The year of 2003 marked a change in the transition length of establishment of agriculture areas, after this year, most of the transitions happened in areas deforested at least 10 years before.
-Even after the decrease of deforestation after 2002, agriculture areas are expanding over lands where deforestation happened before 2002.
-However, after 2019, a sudden drop of agriculture establishment rate happened.
+The period between 2001 and 2005 showed a peak of deforestation destined to agriculture (Figure \@ref(fig:transbar-plot).a), and of croplands expansion (Figure \@ref(fig:transbar-plot).b).
+From 2006 to 2010, we observe a decrease of both variables, a process that were already described by an analysis from 2001 and 2010 by @Macedo2012, which pointed for the decoupling of soybean expansion and deforestation.
+However, our results shows that after 2011, agriculture areas rapidly expanded in the amazon biome, specially at areas deforested many years before.
 
@@
 Political context, public policies, market prices and law enforcement can influence how these processes evolve over time.
+The end of the 1980s and the 1990s were marked by the development of policies for protection of the environment, with the creation of the National Environmental Policy, and the establishment of the Brazilian Institute of Environment and Natural Resources (IBAMA) and the Ministry of Environment [@Banerjee2009].
+However, this was not immediately translated as a reduction of deforestation in the Amazon, which remained at high levels.
+Our results shows that deforestation in this period were mainly occupied by pasture for a long time before being converted to agriculture (Figure \@ref(fig:transbar-plot).a).
+
+Apart from protectionist policies created in the 1990s, development programs kept pressuring forests in the Amazon.
+In the late 1990s, the development policies "Brasil em Ação" and "Avança Brasil" (1995 - 2003) accelerated the infrastructure expansion, including in the Amazon [@Carvalho2002].
+This is a period when the deforestation areas to be converted to agriculture reached the highest values in the time series (Figure \@ref(fig:transbar-plot).a), and also a peak in the establishment of agriculture in the Amazon biome, predominantly with fast transitions (Figure \@ref(fig:transbar-plot).b).
+
 In 2004, the Brazilian Government launched the Action Plan for the Prevention and Control of Deforestation in the Legal Amazon (PPCDAm), which was composed by many initiatives to curb deforestation [@West2021].
 The PPCDAm was considered as a successful policy to slow deforestation rates in Brazil, with international recognition.
-Our calculations from MapBiomas data reinforces the corelation of the PPCDAm with the reduction of deforestation after 2004, and reduction of the agriculture establishment after 2005.
+Our calculations from MapBiomas data reinforces the corelation of the PPCDAm with the reduction of deforestation after 2004, and reduction of the agriculture establishment after 2005 (Figure \@ref(fig:transbar-plot).a).
 
@@
 Our estimates of transitions show a decrease of deforested areas to be converted to agriculture after 2008, where it reached minimal values (Figure \@ref(fig:transbar-plot).a).
-After 2006, agriculture establishment over deforested primary forests suffered a decrease, which stayed relatively stable until 2012, where a steep increase occurred, however, the new areas being occupied by agriculture were mainly over areas that were cleared more than a decade before (therefore, before 2006) (Figure \@ref(fig:transbar-plot).b) This shows that agriculture expansion did not halt after the soy moratorium, producers started expanding in old cleared areas.
-Expansion of agriculture areas also expanded over cleared areas of secondary forests in 2012, but with an important amount of fast transitions.
+After 2006, agriculture establishment over deforested primary forests suffered a decrease, which stayed relatively stable until 2012, where a steep increase occurred, however, the new areas being occupied by agriculture were mainly over areas that were cleared more than a decade before (therefore, before 2006) (Figure \@ref(fig:transbar-plot).b) This shows that agriculture expansion did not halt after the soy moratorium, producers started expanding in old cleared areas, which may have indirect impacts on deforestation increase over distant areas [@Arima2011].
+However, this indirect effect is not always clear when analyzing the Amazon biome as a whole, our data shows that periods of fast agriculture expansion occurred at the same time with minimal deforestation values, according to PRODES.
+Expansion of agriculture areas also expanded over cleared areas of secondary forests in 2012, but with an important amount of fast transitions (Figure @ref(fig:transbar-plot).d).
 The causes of the increase of agriculture establishment areas can be numerous, one main driver was the approval of a new Forest Code, in 2012, which is considered to have undermined the environmental protection of forests [@Kroger2017; @Pereira2019].

### 2022-11-23 manuscript/manuscript.tex
Add text to the methods and the results.

@@
 \item
-  Calculate the year of first occurrence of a non-natural LULC class for each pixel;
+  Calculate the year of first occurrence of any agriculture class for each pixel;
 \item
-  Calculate the first year of ``Forest Formation'' LULC class after the year calculated in step 2;
+  Calculate the first year of ``Forest Formation'' LULC class after the year calculated in step 2.
+  This step identifies if here are more than one transition in a pixel;
 \item
-  Classify rows as ``before'' or ``after'' the occurrence of the year calculated in step 3;
+  Classify rows as ``before'' or ``after'' the occurrence of the year calculated in step 3.
+  This is the identification of periods before and after the recurrence of forest after a first transition to agriculture;
 \item
@@
 
-TODO: Talk about public policies from 1985 to 2000 (greater carajas program, national environmental policy) \citep{Banerjee2009}.
+The deforestation area of primary forests increased largely from 1986 to 2003, which was followed be a steep decrease until 2009, when the deforested areas reached a stable rate (Figure \ref{fig:transbar-plot}.a).
+From 1986 to approximately 1995, most of deforested areas suffered a slow transition, mostly were higher than 10 years, after this period, fast transitions started to become more common, specially from 2002 to 2004 (Figure \ref{fig:transbar-plot}.a).
+The peak of deforestation in 2003, according to our results, shows that 12.6\% of the transitions were a direct conversion from forest to agriculture, estimates by \citet{Morton2006} point to a proportion of 23\%.
+If we include fast transitions (from 0 to 2 years) from our results, the proportion jumps to 48.5 \% of all the transitions.
 
-TODO: Talk about the Avança Brasil plan \citep{Carvalho2002, Banerjee2009}.
+Agriculture establishment over areas of primary forests peaked in 2005 and 2013 (Figure \ref{fig:transbar-plot}.b).
+Despite similar rates between both years, their transition lengths differ greatly, in 2005 most of the transitions were faster than 10 years, while in 2013 the great majority of transitions were slower than 10 years (Figure \ref{fig:transbar-plot}.b).
+The year of 2003 marked a change in the transition length of establishment of agriculture areas, after this year, most of the transitions happened in areas deforested at least 10 years before (Figure \ref{fig:transbar-plot}.b).
+Even after the decrease of deforestation after 2002, agriculture areas are expanding over lands where deforestation happened before 2002.
+However, after 2019, a sudden drop of agriculture establishment rate happened (Figure \ref{fig:transbar-plot}.b).
 
-TODO: confront our results with Morton 2006 (see the year of 2003).
-Confront results from \citep{Macedo2012}, talking about how the expansion of soybean areas are still responsible for production growth, because of very slow transitions.
-
-TODO: talk that even with soy moratorium, and that the deforestation related to transitions were smaller, the deforestation related to pasture occupation kept going, with indirect effects of soybean expansion (Arima 2011).
-
-The deforestation area of primary forests increased largely from 1986 to 2003, which was followed be a steep decrease until 2009, when the deforested areas reached a stable rate.
-From 1986 to approximately 1995, most of deforested areas suffered a slow transition, mostly were higher than 10 years, after this period, fast transitions started to become more common, specially from 2002 to 2004.
-
-Agriculture establishment over areas of primary forests peaked in 2005 and 2013.
-Despite similar rates between both years, their transition lengths differ greatly, in 2005 most of the transitions were faster than 10 years, while in 2013 the great majority of transitions were slower than 10 years.
-The year of 2003 marked a change in the transition length of establishment of agriculture areas, after this year, most of the transitions happened in areas deforested at least 10 years before.
-Even after the decrease of deforestation after 2002, agriculture areas are expanding over lands where deforestation happened before 2002.
-However, after 2019, a sudden drop of agriculture establishment rate happened.
+The period between 2001 and 2005 showed a peak of deforestation destined to agriculture (Figure \ref{fig:transbar-plot}.a), and of croplands expansion (Figure \ref{fig:transbar-plot}.b).
+From 2006 to 2010, we observe a decrease of both variables, a process that were already described by an analysis from 2001 and 2010 by \citet{Macedo2012}, which pointed for the decoupling of soybean expansion and deforestation.
+However, our results shows that after 2011, agriculture areas rapidly expanded in the amazon biome, specially at areas deforested many years before.
 
@@
 Political context, public policies, market prices and law enforcement can influence how these processes evolve over time.
+The end of the 1980s and the 1990s were marked by the development of policies for protection of the environment, with the creation of the National Environmental Policy, and the establishment of the Brazilian Institute of Environment and Natural Resources (IBAMA) and the Ministry of Environment \citep{Banerjee2009}.
+However, this was not immediately translated as a reduction of deforestation in the Amazon, which remained at high levels.
+Our results shows that deforestation in this period were mainly occupied by pasture for a long time before being converted to agriculture (Figure \ref{fig:transbar-plot}.a).
+
+Apart from protectionist policies created in the 1990s, development programs kept pressuring forests in the Amazon.
+In the late 1990s, the development policies ``Brasil em Ação'' and ``Avança Brasil'' (1995 - 2003) accelerated the infrastructure expansion, including in the Amazon \citep{Carvalho2002}.
+This is a period when the deforestation areas to be converted to agriculture reached the highest values in the time series (Figure \ref{fig:transbar-plot}.a), and also a peak in the establishment of agriculture in the Amazon biome, predominantly with fast transitions (Figure \ref{fig:transbar-plot}.b).
+
 In 2004, the Brazilian Government launched the Action Plan for the Prevention and Control of Deforestation in the Legal Amazon (PPCDAm), which was composed by many initiatives to curb deforestation \citep{West2021}.
 The PPCDAm was considered as a successful policy to slow deforestation rates in Brazil, with international recognition.
-Our calculations from MapBiomas data reinforces the corelation of the PPCDAm with the reduction of deforestation after 2004, and reduction of the agriculture establishment after 2005.
+Our calculations from MapBiomas data reinforces the corelation of the PPCDAm with the reduction of deforestation after 2004, and reduction of the agriculture establishment after 2005 (Figure \ref{fig:transbar-plot}.a).
 
@@
 Our estimates of transitions show a decrease of deforested areas to be converted to agriculture after 2008, where it reached minimal values (Figure \ref{fig:transbar-plot}.a).
-After 2006, agriculture establishment over deforested primary forests suffered a decrease, which stayed relatively stable until 2012, where a steep increase occurred, however, the new areas being occupied by agriculture were mainly over areas that were cleared more than a decade before (therefore, before 2006) (Figure \ref{fig:transbar-plot}.b) This shows that agriculture expansion did not halt after the soy moratorium, producers started expanding in old cleared areas.
-Expansion of agriculture areas also expanded over cleared areas of secondary forests in 2012, but with an important amount of fast transitions.
+After 2006, agriculture establishment over deforested primary forests suffered a decrease, which stayed relatively stable until 2012, where a steep increase occurred, however, the new areas being occupied by agriculture were mainly over areas that were cleared more than a decade before (therefore, before 2006) (Figure \ref{fig:transbar-plot}.b) This shows that agriculture expansion did not halt after the soy moratorium, producers started expanding in old cleared areas, which may have indirect impacts on deforestation increase over distant areas \citep{Arima2011}.
+However, this indirect effect is not always clear when analyzing the Amazon biome as a whole, our data shows that periods of fast agriculture expansion occurred at the same time with minimal deforestation values, according to PRODES.
+Expansion of agriculture areas also expanded over cleared areas of secondary forests in 2012, but with an important amount of fast transitions (Figure \citet{ref}(fig:transbar-plot).d).
 The causes of the increase of agriculture establishment areas can be numerous, one main driver was the approval of a new Forest Code, in 2012, which is considered to have undermined the environmental protection of forests \citep{Kroger2017, Pereira2019}.

### 2022-11-29 manuscript/manuscript.Rmd
Update author information. Update abstract.

@@
     affiliation: 2
-  - given_name: Ramon
-    surname: Bicudo
-    affiliation: 1
+  - given_name: Ramon Felipe
+    surname: Bicudo da Silva
+    affiliation: "1, 3"
 
@@
     address: Embrapa Territorial, Av. Soldado Passarinho 303, Brazil
+  - code: 3
+    address: Center for Systems Integration and Sustainability, Department of Fisheries and Wildlife, Michigan State University, East Lansing, MI, 48823, USA
 
 abstract: |
-  Land use and land cover changes in the Amazon biome are a main process that influences many aspects at local, national, and global levels.
-A large quantity of studies already reliad on land use classification to analyse this process, drivers and impacts.
-A new dataset was created by calculating the time necessary for deforested forests to transition to agriculture in the Brazillian Amazon biome.
-The new data can be useful in interdiciplinary studies about land cover and land use change in brazil, its drivers and implications.
-A main inovation is that the dataset links the deforestation year with the year of agriculture establishment, which can provide new information about this process.
+  Land-use and land-cover changes in the Amazon biome are a main process that influences the environment and societies at local, national, and global levels.
+A large quantity of studies already relied on land-use classification to analyze this process.
+This study brings a new dataset created by calculating the time necessary for deforested areas to change to agriculture in the Brazilian Amazon biome.
+The new data can be useful in interdisciplinary studies about land-use and land-cover change analysis in Brazil, such as exploring its drivers and evaluating the effects on natural vegetation cover and ecosystem services.
+The major innovation brought by this study are links between the deforestation year of a given pixel with its respective ywar of agriculture establishment, which can provide new insights to understand long-term land transition processes in tropical ecosystems.
 

### 2022-11-29 manuscript/manuscript.tex
Update author information. Update abstract.

@@
 \Author[2]{Fabiana}{Da Silva Soares}
-\Author[1]{Ramon}{Bicudo}
+\Author[1, 3]{Ramon Felipe}{Bicudo da Silva}
 
@@
 \affil[2]{Embrapa Territorial, Av. Soldado Passarinho 303, Brazil}
+\affil[3]{Center for Systems Integration and Sustainability, Department of Fisheries and Wildlife, Michigan State University, East Lansing, MI, 48823, USA}
 
@@
 \begin{abstract}
-Land use and land cover changes in the Amazon biome are a main process that influences many aspects at local, national, and global levels.
-A large quantity of studies already reliad on land use classification to analyse this process, drivers and impacts.
-A new dataset was created by calculating the time necessary for deforested forests to transition to agriculture in the Brazillian Amazon biome.
-The new data can be useful in interdiciplinary studies about land cover and land use change in brazil, its drivers and implications.
-A main inovation is that the dataset links the deforestation year with the year of agriculture establishment, which can provide new information about this process.
+Land-use and land-cover changes in the Amazon biome are a main process that influences the environment and societies at local, national, and global levels.
+A large quantity of studies already relied on land-use classification to analyze this process.
+This study brings a new dataset created by calculating the time necessary for deforested areas to change to agriculture in the Brazilian Amazon biome.
+The new data can be useful in interdisciplinary studies about land-use and land-cover change analysis in Brazil, such as exploring its drivers and evaluating the effects on natural vegetation cover and ecosystem services.
+The major innovation brought by this study are links between the deforestation year of a given pixel with its respective ywar of agriculture establishment, which can provide new insights to understand long-term land transition processes in tropical ecosystems.
 \end{abstract}

### 2022-11-29 manuscript/manuscript.Rmd
Add todo for the abstract.
Update introduction.

@@
 
-Brazil policies since the 1960s were focused on the expansion of the frontier in the Amazon, specially focused in economic income and land security [@Carvalho2002; @Mcdonald2003; @Banerjee2009].
-Such development model led to the construction of thousands of kilometers of roads, and the settlement of large scale livestock farms [@Carvalho2002; @Banerjee2009].
+Brazilian policies since the 1960s were focused on the expansion of the agricultural frontier in the Amazon, specially focused in in fostering economic growth and national security of the territory [@Carvalho2002; @Mcdonald2003; @Banerjee2009].
+Such development model led to the construction of thousands of kilometers of roads, and the settlement of large scale crop and livestock farms [@Carvalho2002; @Banerjee2009].
 This period was marked with high rates of deforestation of mature forests [@Fearnside2005].
-At the same time, cattle and soybean productions started to move to the north direction in the Brazilian territory, in which soybeans cultivation areas followed cattle ranches expansion over the Amazon forest [@Simon2005; @Barona2010; @Arima2011].
-This pattern of extensive livestock occupation after deforestation, followed by the establishment of annual crops is very characteristic in the Amazon biome, and persists nowadays.
-This process can take several decades to be accomplished, or even less than one year, in which it may be considered as a direct transition from forest to agriculture.
+At the same time, cattle and soybean production started to move from Southern Brazil into the Midwest and Northern parts of the country, in which soybean cultivation areas followed cattle expansion (i.e., pastureland expansion) over the Amazon forest [@Simon2005; @Barona2010; @Arima2011].
+The pattern of extensive livestock production after deforestation, followed by the establishment of annual crops after a given period is typical in the Amazon biome, and persists nowadays (needs reference).
+This process, also known as \"indirect land-use change\" (needs reference) can take several decades to be accomplished, but sometimes less than one year.
+In this last case, the land change process can be considered as a direct transition (i.e., from forest to agriculture).
 
-Land cover transitions are known to cause impacts on different scales.
-Deforestation and land use changes can affect hidrological processes in large basins [@Arias2018]; reduce the forest resilience to extreme events and other sources of perturbation [@Boulton2022]; deforestation and farming practices can reduce convective rainfall and increase surface temperature [@Maeda2021]; increase carbon emissions [@Gatti2021]; cause important impacts on fauna and flora diversity, soil properties and carbon stocks [@Nunes2022]; and even negatively impact public health [@Ellwanger2020].
+Land-use and land-cover changes (LULC) are known to cause impacts on different scales.
+Deforestation and land use changes can affect hydrological processes in large basins [@Arias2018]; reduce the forest resilience to extreme events (e.g., drought) and other sources of perturbation [@Boulton2022]; deforestation and farming practices can reduce convective rainfall and increase surface temperature [@Maeda2021]; increase carbon emissions [@Gatti2021]; cause important impacts on fauna and flora diversity, soil properties and carbon stocks [@Nunes2022]; and even negatively impact public health [@Ellwanger2020].
 
-Data on land use and land cover (LULC) classification are evolving rapidly in the last decade.
-Initiatives such as MapBiomas [@Souza2020], launched in 2015, which provides high resolution, annual classifications for the Brazilian territory.
+Data on LULC classification are evolving rapidly in the last decade.
+Initiatives such as MapBiomas [@Souza2020], launched in 2015, provides high resolution, annual classifications for the Brazilian territory.
 Such type of data are essential to analyse LULC change processes, make causal inferences, and understand what drives and what are the impacts of the transformation of the landscape in Brazil.
-In this context, this project aims to characterize and quantify the length of the transitions from forest formations to agriculture in the Amazon.
+In this context, this study aims to characterize and quantify the length of the transitions from forest formations to agriculture in the Amazon over the last 36 years.
 The availability of this data can be useful to the development of interdisciplinary research involved with LULC changes.
@@
 \conclusions[Conclusions]
+
+TODO: resumo ainda precisa de mais informações, por exemplo, a série temporal da análise, a escala (ex. todo o bioma Amazônia no Brasil, na América do Sul, apenas alguns estados da Amazônia).
+Falar do método, que tipo de álgebra de mapa foi usada, algum tipo de regressão...enfim.

### 2022-11-29 manuscript/manuscript.Rmd
Minor fix of afiliation.

@@
   - code: 1
-    address: Center for Environmental Studies and Research, University of Campinas, Rua dos Flamboyants 155, Brazil
+    address: Center for Environmental Studies and Research, State University of Campinas, Rua dos Flamboyants 155, Brazil
   - code: 2

### 2022-11-30 manuscript/manuscript.Rmd
Update introducation and methods.

@@
 The pattern of extensive livestock production after deforestation, followed by the establishment of annual crops after a given period is typical in the Amazon biome, and persists nowadays (needs reference).
-This process, also known as \"indirect land-use change\" (needs reference) can take several decades to be accomplished, but sometimes less than one year.
+This process, also known as "indirect land-use change" (needs reference) can take several decades to be accomplished, but sometimes less than one year.
 In this last case, the land change process can be considered as a direct transition (i.e., from forest to agriculture).
 
-Land-use and land-cover changes (LULC) are known to cause impacts on different scales.
+Land-use and land-cover (LULC) changes are known to cause impacts on different scales.
 Deforestation and land use changes can affect hydrological processes in large basins [@Arias2018]; reduce the forest resilience to extreme events (e.g., drought) and other sources of perturbation [@Boulton2022]; deforestation and farming practices can reduce convective rainfall and increase surface temperature [@Maeda2021]; increase carbon emissions [@Gatti2021]; cause important impacts on fauna and flora diversity, soil properties and carbon stocks [@Nunes2022]; and even negatively impact public health [@Ellwanger2020].
@@
 The estimations of transition from forest to agriculture were performed for the Amazon biome region, as defined by the Brazilian Institute of Geography and Statistics (IBGE), in 2019.
-Transitions were calculated using land use and land cover classification data from MapBiomas, collection 7, which ranges from 1985 to 2021, at a spatial resolution of approximately 30 meters.
+Transitions were calculated using LULC classification data from MapBiomas, collection 7, which ranges from 1985 to 2021, at a spatial resolution of approximately 30 meters, and global accuracy at 91.3%.
 
@@
 The pixels of the Amazon biome were filtered to contain only those that were occupied by forest and agriculture at some period, and that are not considered as water, according to the Global Surface Water product provided by Copernicus [@Pekel2016].
-For the download, the data was divided in tiles, to allow local processing of a large quantity of pixels.Two sets of data were downloaded with the same spatial dimensions, one with the LULC classifications for each year, and another with binary values, representing whether that pixels was valid for processing or not.
+For the download, the data was divided in 2732 tiles, to allow local processing of a large quantity of pixels.Two sets of data were downloaded with the same spatial dimensions, one with the LULC classifications for each year, and another with binary values, representing whether that pixels was valid for processing or not.
 
@@
 
-3.  Calculate the first year of "Forest Formation" LULC class after the year calculated in step 2.
-    This step identifies if here are more than one transition in a pixel;
+3.  Calculate the first year of \"Forest Formation\" LULC class after the year calculated in step 2.
+    This step identifies the occurrence of more than one transition in a given pixel;
 
-4.  Classify rows as "before" or "after" the occurrence of the year calculated in step 3.
+4.  Classify rows as \"before\" or \"after\" the occurrence of the year calculated in step 3.
     This is the identification of periods before and after the recurrence of forest after a first transition to agriculture;
 
-5.  Calculate the last year of "Forest Formation" within the rows classified as "before", and add 1 year to represent the deforestation year;
+5.  Calculate the last year of \"Forest Formation\" within the rows classified as \"before\", and add 1 year to represent the deforestation year;
 
-6.  Calculate the first year of any agriculture type class within the rows classified as "before", for each pixel;
+6.  Calculate the first year of any agriculture type class within the rows classified as \"before\", for each pixel;
 
@@
 The steps 2 to 7 are performed recursively to identify multiple transitions, in case they are present.
-In addition to the transition length values and the years of the transition, transitions are also qualified by the type of agriculture that were established after deforestation, if the transition occurred from primary or secondary forests, and the number of recurrence of transitions in a single pixel.
+In addition to the transition length values and the years of the transition, transitions are also qualified by the type of agriculture that were established after deforestation, if the transition occurred from primary or secondary forests, and the number of recurrences of transitions at the pixel.
 The LULC classes within the transition and until 5 years after the transitions were also stored.
@@
 
-An accuracy assessment was performed by visual inspection of annual composites of Landsat images from MapBiomas.
-We selected 100 random points to be analysed.
-An area of approximately 4 squared kilometers around the sample point was used in the visual inspection of satellite images.
+Accuracy assessment was performed by visual inspection of annual composites of Landsat images from MapBiomas.
+We selected 100 random points to be analyzed.
+An area of approximately 4 square kilometers around the sample point was used in the visual inspection of satellite images.
 
@@
 The process of accuracy assessment was performed in a Shiny app, and was conducted without any consultation to the transition length results.
-In the validation app, we estimated, by visual inspection, the year of deforestation and the year of agriculture establishment.
+In the validation app, we estimated, by visual inspection, the year of deforestation and the year of agricultural establishment.
 The observed transition length was obtained by subtracting both dates.
@@
 To evaluate the accuracy, we calculated the Mean Absolute Error (MAE), the Bias (BIAS), and the Percent Bias (PBIAS).
-We also analysed the results by plotting the errors as frequency bars, and scatter plots between observed and estimated values.
-
+We also analyzed the results by plotting the errors as frequency bars, and scatter plots between observed and estimated values.
 After the completion of the analysis of the 100 sample points, we also conducted a qualitative assessment, where we compared our results with the satellite images.
@@
 
--   Tables that contains the spatial information (longitude, latitude) of each pixel, its unique id and the code of the municipality which contains the pixel.
-    It is named as "mask_cells";
+-   Tables that contain the spatial information (longitude, latitude) of each pixel, its unique id and the code of the municipality which contains the pixel.
+    It is named as \"mask_cells\";
 
@@
 Another table containing the metadata of each tile is also created, and holds the spatial characteristics of the tiles.
-With this spatial information, it is possible to convert the tabular data back to spatial raster, with identical spatial properties as the MapBiomas classification data.
+With this spatial information, it is possible to convert the tabular data back to spatial raster, with identical spatial properties as the original MapBiomas classification data.
 

### 2022-11-30 manuscript/manuscript.Rmd
Update results and discussion.

@@
 
-The transition calculations shows that between 1985 and 2021, 64874 squared kilometers of forests were converted to agriculture, in the Brazilian Amazon biome.
-The length of the transitions can go from 0 to 35 years, in which transitions closer to 0 year are considered as fast transitions, and transitions closer to 35 years are considered as slow transitions.
-Transitions of 0 year are considered as "direct" transitions, where there were no presence of pasture before the establishment of agriculture, our estimations shows that around 9.2 % of the transitions are considered as "direct".
+The transition calculations show that between 1985 and 2021, 64,874 square kilometers of forests were converted to agriculture in the Brazilian Amazon biome.
+The length of the transitions can go from 0 to 35 years, in which transitions closer to 0 years are considered as fast transitions, and transitions closer to 35 years are considered as slow transitions.
+Transitions of 0 years are considered \"direct\" transitions, where there was no presence of pasture before the establishment of agriculture.
+Our estimations show that around 9.2% of the transitions were \"direct\".
 
-Our transitions calculations found pixels that presented up to six transitions from forest to agriculture since 1985, as these numbers are unlikely to happen (they are distributed as sparse pixels, and do not show patterns of real transitions, being common at borders), we proceeded the results analysis only in the first transitions found in each pixel.
+Our transitions calculations found pixels that presented up to six transitions from forest to agriculture since 1985, as these numbers are unlikely to happen (they are distributed as sparse pixels, and do not show patterns of real transitions, being common at borders), we proceeded the results analysis only in the first transitions found in each of these pixels.
 
-Although we named as deforestation the last years identified as forest before classification of antropic cover, we acknowledge that it is not a direct measurement of deforestation (such as PRODES), it is however a proxy to deforestation.
+Although we named the year of change from forest to anthropic land uses as deforestation, we acknowledge that it is not a direct measurement of deforestation (such as PRODES), however it represents a proxy to deforestation.
 
@@
 
-```{r map-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = "Map of distribution of transitions from forests to agriculture in the Brazillian Amazon biome.
+```{r map-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = "Map of distribution of transitions from forest to agriculture in the Brazilian Amazon biome.
 The hexagonal cells represent the most common transition length, and do not reflect the amount of area of transitions inside a cell.
-Transitions are concentraded in the south (Mato Grosso state), and in the east (Pará and Maranhão states).
-The transition length ranges from 1 (blue tones) to 36 years (pink tones), and clusters of fast transitions (transitions closer to 1 year) can be discerned from clusters of slow transitions (transitions closer to 36 years).
+Transitions are concentrated in the south (Mato Grosso state), and in the east (Pará and Maranhão states).
+The transition length ranges from 0 (blue tones) to 35 years (pink tones), and clusters of fast transitions (transitions closer to 1 year) can be discerned from clusters of slow transitions (transitions closer to 36 years).
 The histogram located in the bottom left shows that fast transitions are more common than slower transitions.
@@
 
-Well defined clusters can be observed in the map created with aggregated transition length data (Figure \@ref(fig:map-plot)), slow transition areas tend to concentrate in specific regions in the Amazon biome, while fast transitions seem to have a wider distributions, but also tend to form clusters.
+Well defined clusters can be observed in the map created with aggregated transition length data (Figure \@ref(fig:map-plot)).
+Slow transition areas tend to concentrate in specific regions of the biome, while fast transitions seem to have a wider distribution, but also tend to form spatial clusters.
 However, this pattern does not hold completely when observing the data at its original scale (Figure \@ref(fig:map-plot)), where areas with different transition lengths are mixed between each other.
@@
 Other studies investigating patterns of transitions in the Amazon also found the formation of clusters of patterns, although there is a big heterogeneity at larger scales [@MullerHansen2017].
+The data can be analyzed year by year, and also be separated by primary and secondary forests being converted to agriculture (Figure \@ref(fig:transbar-plot)).
 
-The data can be analysed year by year, and also be separated by primary and secondary forests being converted to agriculture (Figure \@ref(fig:transbar-plot)).
-
-```{r transbar-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = "Transition area per year and transition length.
+```{r transbar-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = "Transition area from forest to agriculture, per year and transition length.
 The bars represent the total amount of area at some state of the transition for each year.
-The color gradient in each bar represents the transition length related to a deforestation or a agriculture establishment event.
-Blue tones represents fast transitions (transitions closer to 0 years), pink tones represents slow transitions (transitions closer to 35 years).
+The color gradient in each bar represents the transition length related to a deforestation or an agriculture establishment event.
+Blue tones represents fast transitions (transitions closer to 0 years), pink tones represent slow transitions (transitions closer to 35 years).
 Transition events were separated by deforestation of primary forests (a) and secondary forests (c), and the subsequent agriculture establishment of primary forests (b) and secondary forests (d)."}
@@
 
-The deforestation area of primary forests increased largely from 1986 to 2003, which was followed be a steep decrease until 2009, when the deforested areas reached a stable rate (Figure \@ref(fig:transbar-plot).a).
-From 1986 to approximately 1995, most of deforested areas suffered a slow transition, mostly were higher than 10 years, after this period, fast transitions started to become more common, specially from 2002 to 2004 (Figure \@ref(fig:transbar-plot).a).
-The peak of deforestation in 2003, according to our results, shows that 12.6% of the transitions were a direct conversion from forest to agriculture, estimates by @Morton2006 point to a proportion of 23%.
-If we include fast transitions (from 0 to 2 years) from our results, the proportion jumps to 48.5 % of all the transitions.
+The deforestation area of primary forests increased largely from 1986 to 2003, which was followed by a steep decrease until 2009, when the deforested areas reached a stable rate (Figure \@ref(fig:transbar-plot).a).
+From 1986 to approximately 1995, most of the deforested areas suffered a slow transition, mostly were higher than 10 years.
+After this period, fast transitions started to become more common, specially from 2002 to 2004 (Figure \@ref(fig:transbar-plot).a).
+The peak of deforestation in 2003, according to our results, shows that 12.6% of the transitions were a direct conversion from forest to agriculture.
+Previous studies estimated a proportion of 23% [@Morton2006].
+If we include fast transitions (from 0 to 2 years) from our results, the proportion jumps to 48.5 % of all transitions.
 
-Agriculture establishment over areas of primary forests peaked in 2005 and 2013 (Figure \@ref(fig:transbar-plot).b).
-Despite similar rates between both years, their transition lengths differ greatly, in 2005 most of the transitions were faster than 10 years, while in 2013 the great majority of transitions were slower than 10 years (Figure \@ref(fig:transbar-plot).b).
-The year of 2003 marked a change in the transition length of establishment of agriculture areas, after this year, most of the transitions happened in areas deforested at least 10 years before (Figure \@ref(fig:transbar-plot).b).
-Even after the decrease of deforestation after 2002, agriculture areas are expanding over lands where deforestation happened before 2002.
+Agricultural establishment over areas of primary forests peaked in 2005 and 2013 (Figure \@ref(fig:transbar-plot).b).
+Despite similar agricultural establishment rates between both years, their transition lengths differ greatly, in 2005 most of the transitions were faster than 10 years, while in 2013 the great majority of transitions were slower than 10 years (Figure \@ref(fig:transbar-plot).b).
+In 2003 it was observed a shift in the transition length from forest until the establishment of agricultural areas.
+From this year onwards, most transitions happened in areas deforested at least 10 years before (Figure \@ref(fig:transbar-plot).b).
+Even after the decrease of deforestation after 2002, areas under agriculture expanded over lands where deforestation happened before 2002 (mostly over pastureland).
 However, after 2019, a sudden drop of agriculture establishment rate happened (Figure \@ref(fig:transbar-plot).b).
@@
 The period between 2001 and 2005 showed a peak of deforestation destined to agriculture (Figure \@ref(fig:transbar-plot).a), and of croplands expansion (Figure \@ref(fig:transbar-plot).b).
-From 2006 to 2010, we observe a decrease of both variables, a process that were already described by an analysis from 2001 and 2010 by @Macedo2012, which pointed for the decoupling of soybean expansion and deforestation.
-However, our results shows that after 2011, agriculture areas rapidly expanded in the amazon biome, specially at areas deforested many years before.
+From 2006 to 2010, we observed a decrease in both variables, a process that was already observed between 2001 and 2010 [@Macedo2012], which points out a decoupling of soybean expansion and deforestation.
+However, our result shows that after 2011, agricultural areas rapidly expanded in the amazon biome, especially at areas deforested many years before.
 
-The causes of deforestation and agriculture establishment in the Brazilian Amazon are complex and diverse.
+The causes of deforestation and agricultural establishment in the Brazilian Amazon are complex and diverse.
 Political context, public policies, market prices and law enforcement can influence how these processes evolve over time.
-The end of the 1980s and the 1990s were marked by the development of policies for protection of the environment, with the creation of the National Environmental Policy, and the establishment of the Brazilian Institute of Environment and Natural Resources (IBAMA) and the Ministry of Environment [@Banerjee2009].
-However, this was not immediately translated as a reduction of deforestation in the Amazon, which remained at high levels.
-Our results shows that deforestation in this period were mainly occupied by pasture for a long time before being converted to agriculture (Figure \@ref(fig:transbar-plot).a).
+The end of the 1980s and the 1990s were marked by the development of policies for protection of the environment, with the creation of the National Environmental Policy, the establishment of the Brazilian Institute of Environment and Natural Resources (IBAMA), and the Ministry of Environment [@Banerjee2009].
+However, this new policy environment was not immediately translated into a significant deforestation decrease in the Amazon, which remained at high levels until the beginning of 2000s.
+Our results show that deforestation in this period (1986 - 2000) was mainly followed by pasture, which remained for a long time before being converted to agriculture (Figure \@ref(fig:transbar-plot).a).
 
 Apart from protectionist policies created in the 1990s, development programs kept pressuring forests in the Amazon.
-In the late 1990s, the development policies "Brasil em Ação" and "Avança Brasil" (1995 - 2003) accelerated the infrastructure expansion, including in the Amazon [@Carvalho2002].
+In the late 1990s, the development policies "Brasil em Ação" and "Avança Brasil" (1995 - 2003) accelerated the national infrastructure expansion, including in the Amazon [@Carvalho2002].
 This is a period when the deforestation areas to be converted to agriculture reached the highest values in the time series (Figure \@ref(fig:transbar-plot).a), and also a peak in the establishment of agriculture in the Amazon biome, predominantly with fast transitions (Figure \@ref(fig:transbar-plot).b).
 
-In 2004, the Brazilian Government launched the Action Plan for the Prevention and Control of Deforestation in the Legal Amazon (PPCDAm), which was composed by many initiatives to curb deforestation [@West2021].
+In 2004, the Brazilian Government launched the Action Plan for the Prevention and Control of Deforestation in the Legal Amazon (PPCDAm), which was composed of many initiatives to curb deforestation [@West2021].
 The PPCDAm was considered as a successful policy to slow deforestation rates in Brazil, with international recognition.
-Our calculations from MapBiomas data reinforces the corelation of the PPCDAm with the reduction of deforestation after 2004, and reduction of the agriculture establishment after 2005 (Figure \@ref(fig:transbar-plot).a).
+Our calculations from MapBiomas data reinforces the correlation of the PPCDAm with the reduction of deforestation after 2004 (Figure \@ref(fig:transbar-plot).a), and reduction of the agriculture establishment after 2005 (Figure \@ref(fig:transbar-plot).b).
 
-In 2006, the Brazilian Association of Vegetable Oil Industries (BIOVE) and the National Association of Cereal Exporters (ANEC) committed to avoid commercialization of soy grains harvested from areas deforested after 2006.
+In 2006, the Brazilian Association of Vegetable Oil Industries (BIOVE) and the National Association of Cereal Exporters (ANEC) committed to avoid commercialization of soybean grains harvested from areas deforested after 2008.
 Our estimates of transitions show a decrease of deforested areas to be converted to agriculture after 2008, where it reached minimal values (Figure \@ref(fig:transbar-plot).a).
-After 2006, agriculture establishment over deforested primary forests suffered a decrease, which stayed relatively stable until 2012, where a steep increase occurred, however, the new areas being occupied by agriculture were mainly over areas that were cleared more than a decade before (therefore, before 2006) (Figure \@ref(fig:transbar-plot).b) This shows that agriculture expansion did not halt after the soy moratorium, producers started expanding in old cleared areas, which may have indirect impacts on deforestation increase over distant areas [@Arima2011].
-However, this indirect effect is not always clear when analyzing the Amazon biome as a whole, our data shows that periods of fast agriculture expansion occurred at the same time with minimal deforestation values, according to PRODES.
-Expansion of agriculture areas also expanded over cleared areas of secondary forests in 2012, but with an important amount of fast transitions (Figure @ref(fig:transbar-plot).d).
-The causes of the increase of agriculture establishment areas can be numerous, one main driver was the approval of a new Forest Code, in 2012, which is considered to have undermined the environmental protection of forests [@Kroger2017; @Pereira2019].
+After 2006, agriculture establishment over deforested primary forests suffered a decrease, which stayed relatively stable until 2012, where a steep increase occurred, however, the new areas being occupied by agriculture were mainly over areas that were cleared more than a decade before (therefore, before 2006) (Figure \@ref(fig:transbar-plot).b) This shows that agriculture expansion did not halt after the soy moratorium, but producers expended over areas cleared many years ago, which may have indirect impacts on deforestation increase over distant areas [@Arima2011].
+However, this indirect effect is not always clearly observed when analyzing the Amazon biome as a whole, but our data shows that periods of fast agriculture expansion occurred at the same time with minimal deforestation values, according to PRODES estimates (add reference).
+Expansion of agricultural areas also occurred over cleared areas of secondary forests in 2012, but with an important amount of fast transitions (Figure @ref(fig:transbar-plot).d).
+The causes of the increase of agricultural establishment areas can be numerous, and one main driver was the approval of a new Forest Code, in 2012, which is considered to have undermined the environmental protection of forests [@Kroger2017; @Pereira2019].
 
@@
 
-```{r transridge-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = "Transition length patterns inside states that belongs to the Amazon biome.
-Each year have a density estimate of the transition lengths, represented as colored curves.
+```{r transridge-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = "Transition length patterns inside states in the Amazon biome.
+Each year has a density estimate of the transition lengths, represented as colored curves.
 The peak of the curves represents transition length values with more frequency in one year of one state.
@@
 
-The state of Amapá presented fast transitions along all the time series, where slow transitions are not as common.
-In contrast, Acre shows a majority of slow transitions, in which only 2021 showed more fast transitions.
-There are three states where the pattern of transitions length across time are alike, Mato Grosso and Pará presented more fast transitions from 1995 to 2005, after this period, slow transitions became more common with time.
-Rondônia and Tocantins also presented a pattern in which transitions became slower with time, however this pattern started to occur earlier than Mato Grosso and Pará.
+The state of Amapá presented fast transitions along all the time series, without any significant period of slow transition.
+In contrast, transitions in Acre were majorly slow, in which only 2021 showed faster transitions.
+There are two states where the pattern of transition length are alike, Mato Grosso (MT) and Pará (PA).
+Both states underwent fast transitions from 1995 to 2005, and after this period slow transitions became predominant until the end of the time series.
+Rondônia (RO) and Tocantins (TO) presented a similar pattern of MT and PA, in which transitions decelerated , but for RO and TO since the beginning of 2000s, earlier than MA and PA.
 
-The great majority of transitions are from forests to Soybeans and "Other Temporary Crops", which represents 95% of the transitions in the Amazon biome (14% of Soybean and 81% of "Other Temporary Crops").
-When analyzing what happened after 5 years since the transition, Soybean farms are persistent, 79% of these areas remained as Soybean, 14% is converted to "Other Temporary Crops" and 7% in occupied by Pasture.
-"Other Temporary Crops" are less persistent, 27% of these areas remains as the same use, 56% is converted to Soybean, and 15% to Pasture.
-Conversions from Soybean and "Other Temporary Uses" to other covers and uses (apart from the already cited) are negligible.
-Sugar Cane, Cotton and Perennial Crops also appears, but in a negligible proportions.
+The great majority of transitions are from forests to soybean and \"Other Temporary Crops\", which represents 95% of the transitions in the Amazon biome (14% of Soybean and 81% of \"Other Temporary Crops\").
+When analyzing what happened after 5 years since the transition, soybean areas persistency rate was at 79% (i.e., areas that remained as soybean), 14% was converted to \"Other Temporary Crops\" and 7% was converted to pasture.
+\"Other Temporary Crops\" are less persistent.
+Hence, 27% of these areas remain in the same class, 56% were converted to soybean, and 15% to pasture.
+Conversions from soybean and \"Other Temporary Uses\" to other LULC classes (apart from the already cited) are negligible.
+Sugar cane, cotton and perennial crops also appear, but in a negligible proportion.
 
@@
 From the 100 random sample points used in the results validation, 1 point (sample 21) was not considered as a transition from our estimations, however, the visual inspection pointed to a likely event of transition from forest to agriculture.
-Also, there were 6 points (sample 6, 7, 55, 56, 78, 97) which the visual inspection did not found a transition from forest to pasture, although our estimations pointed as transitions.
+Also, there were 6 points (sample 6, 7, 55, 56, 78, 97) which the visual inspection did not find a transition from forest to pasture, although our estimations pointed as transitions.
 Therefore, 7% of the sample were completely misclassified by our estimates, and the rest of the accuracy assessment was performed over the remaining 93 sample points.
 
-When analyzing the errors from the transition length estimates, we observe that the year of deforestation shows the least amount of errors (Figure \@ref(fig:errorbar-plot)).
-The MAE of the deforestation year is of 1.42 years, and the error shows a bias towards underestimation.
-The year of the agriculture establishment and the transition length estimates showed larger errors when comparing with visual inspection.
+When analyzing the errors from the transition length estimates, we observed that the year of deforestation shows the least amount of errors (Figure \@ref(fig:errorbar-plot)).
+The MAE of the deforestation year is 1.42 years, and the error shows a bias towards underestimation.
+The year of the agriculture establishment and the transition length estimates showed larger errors when compared with visual inspection.
 
 ```{r errorbar-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = "Bar with the count of error values (difference between observed and estimated values).
-Positive values indicate underestimation of the variable (estimations were lower than observations), negative values indicates overestimation (estimations higher than observations).
+Positive values indicate underestimation of the variable (estimations were lower than observations), negative values indicate overestimation (estimations higher than observations).
 Error metrics (Mean Absolute Error, Bias and Percent Bias) are displayed in the top right position of each box.
-The white deshed lines represents the MAE values of each variable."}
+The white dashed lines represent the MAE values of each variable."}
 
@@
 The Agriculture class presented more errors, the inclusion errors ranged from 22% to 5%, and were mostly composed by Forests pixels (forest pixels misclassified as agriculture).
-The omission errors of Agriculture ranged from 22% to 8%, and were also mostly composed by Forests pixels (agriculture pixels misclassified as Agriculture).
+The omission errors of Agriculture ranged from 22% to 8%, and were also mostly composed of Forest pixels (agriculture pixels misclassified as Agriculture).
 
@@
 Effect of border on results (sample 25, 56, 84, 92).
-Areas with sparse forest vegetation presented large error of deforestation year (sample 33).
+Areas with sparse forest vegetation presented a large error of deforestation year (sample 33).
 
-Patterns of agriculture areas are roughly well represented (sample 25, 43, 56, 84).
-Some locations presents a better homogeneity of transitions inside agriculture plots (sample 84).
+Patterns of agricultural areas are roughly well represented (sample 25, 43, 56, 84).
+Some locations present a better homogeneity of transitions inside agriculture plots (sample 84).
 
 Interference due to atmospheric conditions and lack of image availability clearly shaped spatial patterns of transitions at some locations.
-Quality of composites trough the time series affected accuracy of results.
-Regions of the North part of the Amazon presents lower quality than regions in the South of the Amazon biome.
+Quality of composites through the time series affected accuracy of results.
+Regions of the North part of the Amazon present lower quality than regions in the South of the Amazon biome.
 

### 2022-11-30 manuscript/manuscript.Rmd
Fix figure reference.

@@
 However, this indirect effect is not always clearly observed when analyzing the Amazon biome as a whole, but our data shows that periods of fast agriculture expansion occurred at the same time with minimal deforestation values, according to PRODES estimates (add reference).
-Expansion of agricultural areas also occurred over cleared areas of secondary forests in 2012, but with an important amount of fast transitions (Figure @ref(fig:transbar-plot).d).
+Expansion of agricultural areas also occurred over cleared areas of secondary forests in 2012, but with an important amount of fast transitions (Figure \@ref(fig:transbar-plot).d).
 The causes of the increase of agricultural establishment areas can be numerous, and one main driver was the approval of a new Forest Code, in 2012, which is considered to have undermined the environmental protection of forests [@Kroger2017; @Pereira2019].

### 2022-11-30 manuscript/manuscript.tex
Fix figure reference.

@@
 
-\affil[1]{Center for Environmental Studies and Research, University of Campinas, Rua dos Flamboyants 155, Brazil}
+\affil[1]{Center for Environmental Studies and Research, State University of Campinas, Rua dos Flamboyants 155, Brazil}
 \affil[2]{Embrapa Territorial, Av. Soldado Passarinho 303, Brazil}
@@
 
-Brazil policies since the 1960s were focused on the expansion of the frontier in the Amazon, specially focused in economic income and land security \citep{Carvalho2002, Mcdonald2003, Banerjee2009}.
-Such development model led to the construction of thousands of kilometers of roads, and the settlement of large scale livestock farms \citep{Carvalho2002, Banerjee2009}.
+Brazilian policies since the 1960s were focused on the expansion of the agricultural frontier in the Amazon, specially focused in in fostering economic growth and national security of the territory \citep{Carvalho2002, Mcdonald2003, Banerjee2009}.
+Such development model led to the construction of thousands of kilometers of roads, and the settlement of large scale crop and livestock farms \citep{Carvalho2002, Banerjee2009}.
 This period was marked with high rates of deforestation of mature forests \citep{Fearnside2005}.
-At the same time, cattle and soybean productions started to move to the north direction in the Brazilian territory, in which soybeans cultivation areas followed cattle ranches expansion over the Amazon forest \citep{Simon2005, Barona2010, Arima2011}.
-This pattern of extensive livestock occupation after deforestation, followed by the establishment of annual crops is very characteristic in the Amazon biome, and persists nowadays.
-This process can take several decades to be accomplished, or even less than one year, in which it may be considered as a direct transition from forest to agriculture.
+At the same time, cattle and soybean production started to move from Southern Brazil into the Midwest and Northern parts of the country, in which soybean cultivation areas followed cattle expansion (i.e., pastureland expansion) over the Amazon forest \citep{Simon2005, Barona2010, Arima2011}.
+The pattern of extensive livestock production after deforestation, followed by the establishment of annual crops after a given period is typical in the Amazon biome, and persists nowadays (needs reference).
+This process, also known as ``indirect land-use change'' (needs reference) can take several decades to be accomplished, but sometimes less than one year.
+In this last case, the land change process can be considered as a direct transition (i.e., from forest to agriculture).
 
-Land cover transitions are known to cause impacts on different scales.
-Deforestation and land use changes can affect hidrological processes in large basins \citep{Arias2018}; reduce the forest resilience to extreme events and other sources of perturbation \citep{Boulton2022}; deforestation and farming practices can reduce convective rainfall and increase surface temperature \citep{Maeda2021}; increase carbon emissions \citep{Gatti2021}; cause important impacts on fauna and flora diversity, soil properties and carbon stocks \citep{Nunes2022}; and even negatively impact public health \citep{Ellwanger2020}.
+Land-use and land-cover (LULC) changes are known to cause impacts on different scales.
+Deforestation and land use changes can affect hydrological processes in large basins \citep{Arias2018}; reduce the forest resilience to extreme events (e.g., drought) and other sources of perturbation \citep{Boulton2022}; deforestation and farming practices can reduce convective rainfall and increase surface temperature \citep{Maeda2021}; increase carbon emissions \citep{Gatti2021}; cause important impacts on fauna and flora diversity, soil properties and carbon stocks \citep{Nunes2022}; and even negatively impact public health \citep{Ellwanger2020}.
 
-Data on land use and land cover (LULC) classification are evolving rapidly in the last decade.
-Initiatives such as MapBiomas \citep{Souza2020}, launched in 2015, which provides high resolution, annual classifications for the Brazilian territory.
+Data on LULC classification are evolving rapidly in the last decade.
+Initiatives such as MapBiomas \citep{Souza2020}, launched in 2015, provides high resolution, annual classifications for the Brazilian territory.
 Such type of data are essential to analyse LULC change processes, make causal inferences, and understand what drives and what are the impacts of the transformation of the landscape in Brazil.
-In this context, this project aims to characterize and quantify the length of the transitions from forest formations to agriculture in the Amazon.
+In this context, this study aims to characterize and quantify the length of the transitions from forest formations to agriculture in the Amazon over the last 36 years.
 The availability of this data can be useful to the development of interdisciplinary research involved with LULC changes.
@@
 The estimations of transition from forest to agriculture were performed for the Amazon biome region, as defined by the Brazilian Institute of Geography and Statistics (IBGE), in 2019.
-Transitions were calculated using land use and land cover classification data from MapBiomas, collection 7, which ranges from 1985 to 2021, at a spatial resolution of approximately 30 meters.
+Transitions were calculated using LULC classification data from MapBiomas, collection 7, which ranges from 1985 to 2021, at a spatial resolution of approximately 30 meters, and global accuracy at 91.3\%.
 
@@
 The pixels of the Amazon biome were filtered to contain only those that were occupied by forest and agriculture at some period, and that are not considered as water, according to the Global Surface Water product provided by Copernicus \citep{Pekel2016}.
-For the download, the data was divided in tiles, to allow local processing of a large quantity of pixels.Two sets of data were downloaded with the same spatial dimensions, one with the LULC classifications for each year, and another with binary values, representing whether that pixels was valid for processing or not.
+For the download, the data was divided in 2732 tiles, to allow local processing of a large quantity of pixels.Two sets of data were downloaded with the same spatial dimensions, one with the LULC classifications for each year, and another with binary values, representing whether that pixels was valid for processing or not.
 
@@
 \item
-  Calculate the first year of ``Forest Formation'' LULC class after the year calculated in step 2.
-  This step identifies if here are more than one transition in a pixel;
+  Calculate the first year of "Forest Formation" LULC class after the year calculated in step 2.
+  This step identifies the occurrence of more than one transition in a given pixel;
 \item
-  Classify rows as ``before'' or ``after'' the occurrence of the year calculated in step 3.
+  Classify rows as "before" or "after" the occurrence of the year calculated in step 3.
   This is the identification of periods before and after the recurrence of forest after a first transition to agriculture;
 \item
-  Calculate the last year of ``Forest Formation'' within the rows classified as ``before'', and add 1 year to represent the deforestation year;
+  Calculate the last year of "Forest Formation" within the rows classified as "before", and add 1 year to represent the deforestation year;
 \item
-  Calculate the first year of any agriculture type class within the rows classified as ``before'', for each pixel;
+  Calculate the first year of any agriculture type class within the rows classified as "before", for each pixel;
 \item
@@
 \item
-  Tables that contains the spatial information (longitude, latitude) of each pixel, its unique id and the code of the municipality which contains the pixel.
-  It is named as ``mask\_cells'';
+  Tables that contain the spatial information (longitude, latitude) of each pixel, its unique id and the code of the municipality which contains the pixel.
+  It is named as "mask\_cells";
 \item
@@
 
-The transition calculations shows that between 1985 and 2021, 64874 squared kilometers of forests were converted to agriculture, in the Brazilian Amazon biome.
-The length of the transitions can go from 0 to 35 years, in which transitions closer to 0 year are considered as fast transitions, and transitions closer to 35 years are considered as slow transitions.
-Transitions of 0 year are considered as ``direct'' transitions, where there were no presence of pasture before the establishment of agriculture, our estimations shows that around 9.2 \% of the transitions are considered as ``direct''.
+The transition calculations show that between 1985 and 2021, 64,874 square kilometers of forests were converted to agriculture in the Brazilian Amazon biome.
+The length of the transitions can go from 0 to 35 years, in which transitions closer to 0 years are considered as fast transitions, and transitions closer to 35 years are considered as slow transitions.
+Transitions of 0 years are considered "direct" transitions, where there was no presence of pasture before the establishment of agriculture.
+Our estimations show that around 9.2\% of the transitions were "direct".
 
-Our transitions calculations found pixels that presented up to six transitions from forest to agriculture since 1985, as these numbers are unlikely to happen (they are distributed as sparse pixels, and do not show patterns of real transitions, being common at borders), we proceeded the results analysis only in the first transitions found in each pixel.
+Our transitions calculations found pixels that presented up to six transitions from forest to agriculture since 1985, as these numbers are unlikely to happen (they are distributed as sparse pixels, and do not show patterns of real transitions, being common at borders), we proceeded the results analysis only in the first transitions found in each of these pixels.
 
-Although we named as deforestation the last years identified as forest before classification of antropic cover, we acknowledge that it is not a direct measurement of deforestation (such as PRODES), it is however a proxy to deforestation.
+Although we named the year of change from forest to anthropic land uses as deforestation, we acknowledge that it is not a direct measurement of deforestation (such as PRODES), however it represents a proxy to deforestation.
 
@@
 \begin{figure}[ht]
-\includegraphics[width=17cm]{figs/map} \caption{Map of distribution of transitions from forests to agriculture in the Brazillian Amazon biome.
+\includegraphics[width=17cm]{figs/map} \caption{Map of distribution of transitions from forest to agriculture in the Brazilian Amazon biome.
 The hexagonal cells represent the most common transition length, and do not reflect the amount of area of transitions inside a cell.
-Transitions are concentraded in the south (Mato Grosso state), and in the east (Pará and Maranhão states).
-The transition length ranges from 1 (blue tones) to 36 years (pink tones), and clusters of fast transitions (transitions closer to 1 year) can be discerned from clusters of slow transitions (transitions closer to 36 years).
+Transitions are concentrated in the south (Mato Grosso state), and in the east (Pará and Maranhão states).
+The transition length ranges from 0 (blue tones) to 35 years (pink tones), and clusters of fast transitions (transitions closer to 1 year) can be discerned from clusters of slow transitions (transitions closer to 36 years).
 The histogram located in the bottom left shows that fast transitions are more common than slower transitions.
@@
 
-Well defined clusters can be observed in the map created with aggregated transition length data (Figure \ref{fig:map-plot}), slow transition areas tend to concentrate in specific regions in the Amazon biome, while fast transitions seem to have a wider distributions, but also tend to form clusters.
+Well defined clusters can be observed in the map created with aggregated transition length data (Figure \ref{fig:map-plot}).
+Slow transition areas tend to concentrate in specific regions of the biome, while fast transitions seem to have a wider distribution, but also tend to form spatial clusters.
 However, this pattern does not hold completely when observing the data at its original scale (Figure \ref{fig:map-plot}), where areas with different transition lengths are mixed between each other.
@@
 Other studies investigating patterns of transitions in the Amazon also found the formation of clusters of patterns, although there is a big heterogeneity at larger scales \citep{MullerHansen2017}.
-
-The data can be analysed year by year, and also be separated by primary and secondary forests being converted to agriculture (Figure \ref{fig:transbar-plot}).
+The data can be analyzed year by year, and also be separated by primary and secondary forests being converted to agriculture (Figure \ref{fig:transbar-plot}).
 
 \begin{figure}[ht]
-\includegraphics[width=17cm]{figs/trans_length_cols} \caption{Transition area per year and transition length.
+\includegraphics[width=17cm]{figs/trans_length_cols} \caption{Transition area from forest to agriculture, per year and transition length.
 The bars represent the total amount of area at some state of the transition for each year.
-The color gradient in each bar represents the transition length related to a deforestation or a agriculture establishment event.
-Blue tones represents fast transitions (transitions closer to 0 years), pink tones represents slow transitions (transitions closer to 35 years).
+The color gradient in each bar represents the transition length related to a deforestation or an agriculture establishment event.
+Blue tones represents fast transitions (transitions closer to 0 years), pink tones represent slow transitions (transitions closer to 35 years).
 Transition events were separated by deforestation of primary forests (a) and secondary forests (c), and the subsequent agriculture establishment of primary forests (b) and secondary forests (d).}\label{fig:transbar-plot}
@@
 
-The deforestation area of primary forests increased largely from 1986 to 2003, which was followed be a steep decrease until 2009, when the deforested areas reached a stable rate (Figure \ref{fig:transbar-plot}.a).
-From 1986 to approximately 1995, most of deforested areas suffered a slow transition, mostly were higher than 10 years, after this period, fast transitions started to become more common, specially from 2002 to 2004 (Figure \ref{fig:transbar-plot}.a).
-The peak of deforestation in 2003, according to our results, shows that 12.6\% of the transitions were a direct conversion from forest to agriculture, estimates by \citet{Morton2006} point to a proportion of 23\%.
-If we include fast transitions (from 0 to 2 years) from our results, the proportion jumps to 48.5 \% of all the transitions.
+The deforestation area of primary forests increased largely from 1986 to 2003, which was followed by a steep decrease until 2009, when the deforested areas reached a stable rate (Figure \ref{fig:transbar-plot}.a).
+From 1986 to approximately 1995, most of the deforested areas suffered a slow transition, mostly were higher than 10 years.
+After this period, fast transitions started to become more common, specially from 2002 to 2004 (Figure \ref{fig:transbar-plot}.a).
+The peak of deforestation in 2003, according to our results, shows that 12.6\% of the transitions were a direct conversion from forest to agriculture.
+Previous studies estimated a proportion of 23\% \citep{Morton2006}.
+If we include fast transitions (from 0 to 2 years) from our results, the proportion jumps to 48.5 \% of all transitions.
 
-Agriculture establishment over areas of primary forests peaked in 2005 and 2013 (Figure \ref{fig:transbar-plot}.b).
-Despite similar rates between both years, their transition lengths differ greatly, in 2005 most of the transitions were faster than 10 years, while in 2013 the great majority of transitions were slower than 10 years (Figure \ref{fig:transbar-plot}.b).
-The year of 2003 marked a change in the transition length of establishment of agriculture areas, after this year, most of the transitions happened in areas deforested at least 10 years before (Figure \ref{fig:transbar-plot}.b).
-Even after the decrease of deforestation after 2002, agriculture areas are expanding over lands where deforestation happened before 2002.
+Agricultural establishment over areas of primary forests peaked in 2005 and 2013 (Figure \ref{fig:transbar-plot}.b).
+Despite similar agricultural establishment rates between both years, their transition lengths differ greatly, in 2005 most of the transitions were faster than 10 years, while in 2013 the great majority of transitions were slower than 10 years (Figure \ref{fig:transbar-plot}.b).
+In 2003 it was observed a shift in the transition length from forest until the establishment of agricultural areas.
+From this year onwards, most transitions happened in areas deforested at least 10 years before (Figure \ref{fig:transbar-plot}.b).
+Even after the decrease of deforestation after 2002, areas under agriculture expanded over lands where deforestation happened before 2002 (mostly over pastureland).
 However, after 2019, a sudden drop of agriculture establishment rate happened (Figure \ref{fig:transbar-plot}.b).
@@
 The period between 2001 and 2005 showed a peak of deforestation destined to agriculture (Figure \ref{fig:transbar-plot}.a), and of croplands expansion (Figure \ref{fig:transbar-plot}.b).
-From 2006 to 2010, we observe a decrease of both variables, a process that were already described by an analysis from 2001 and 2010 by \citet{Macedo2012}, which pointed for the decoupling of soybean expansion and deforestation.
-However, our results shows that after 2011, agriculture areas rapidly expanded in the amazon biome, specially at areas deforested many years before.
+From 2006 to 2010, we observed a decrease in both variables, a process that was already observed between 2001 and 2010 \citep{Macedo2012}, which points out a decoupling of soybean expansion and deforestation.
+However, our result shows that after 2011, agricultural areas rapidly expanded in the amazon biome, especially at areas deforested many years before.
 
-The causes of deforestation and agriculture establishment in the Brazilian Amazon are complex and diverse.
+The causes of deforestation and agricultural establishment in the Brazilian Amazon are complex and diverse.
 Political context, public policies, market prices and law enforcement can influence how these processes evolve over time.
-The end of the 1980s and the 1990s were marked by the development of policies for protection of the environment, with the creation of the National Environmental Policy, and the establishment of the Brazilian Institute of Environment and Natural Resources (IBAMA) and the Ministry of Environment \citep{Banerjee2009}.
-However, this was not immediately translated as a reduction of deforestation in the Amazon, which remained at high levels.
-Our results shows that deforestation in this period were mainly occupied by pasture for a long time before being converted to agriculture (Figure \ref{fig:transbar-plot}.a).
+The end of the 1980s and the 1990s were marked by the development of policies for protection of the environment, with the creation of the National Environmental Policy, the establishment of the Brazilian Institute of Environment and Natural Resources (IBAMA), and the Ministry of Environment \citep{Banerjee2009}.
+However, this new policy environment was not immediately translated into a significant deforestation decrease in the Amazon, which remained at high levels until the beginning of 2000s.
+Our results show that deforestation in this period (1986 - 2000) was mainly followed by pasture, which remained for a long time before being converted to agriculture (Figure \ref{fig:transbar-plot}.a).
 
 Apart from protectionist policies created in the 1990s, development programs kept pressuring forests in the Amazon.
-In the late 1990s, the development policies ``Brasil em Ação'' and ``Avança Brasil'' (1995 - 2003) accelerated the infrastructure expansion, including in the Amazon \citep{Carvalho2002}.
+In the late 1990s, the development policies ``Brasil em Ação'' and ``Avança Brasil'' (1995 - 2003) accelerated the national infrastructure expansion, including in the Amazon \citep{Carvalho2002}.
 This is a period when the deforestation areas to be converted to agriculture reached the highest values in the time series (Figure \ref{fig:transbar-plot}.a), and also a peak in the establishment of agriculture in the Amazon biome, predominantly with fast transitions (Figure \ref{fig:transbar-plot}.b).
 
-In 2004, the Brazilian Government launched the Action Plan for the Prevention and Control of Deforestation in the Legal Amazon (PPCDAm), which was composed by many initiatives to curb deforestation \citep{West2021}.
+In 2004, the Brazilian Government launched the Action Plan for the Prevention and Control of Deforestation in the Legal Amazon (PPCDAm), which was composed of many initiatives to curb deforestation \citep{West2021}.
 The PPCDAm was considered as a successful policy to slow deforestation rates in Brazil, with international recognition.
-Our calculations from MapBiomas data reinforces the corelation of the PPCDAm with the reduction of deforestation after 2004, and reduction of the agriculture establishment after 2005 (Figure \ref{fig:transbar-plot}.a).
+Our calculations from MapBiomas data reinforces the correlation of the PPCDAm with the reduction of deforestation after 2004 (Figure \ref{fig:transbar-plot}.a), and reduction of the agriculture establishment after 2005 (Figure \ref{fig:transbar-plot}.b).
 
-In 2006, the Brazilian Association of Vegetable Oil Industries (BIOVE) and the National Association of Cereal Exporters (ANEC) committed to avoid commercialization of soy grains harvested from areas deforested after 2006.
+In 2006, the Brazilian Association of Vegetable Oil Industries (BIOVE) and the National Association of Cereal Exporters (ANEC) committed to avoid commercialization of soybean grains harvested from areas deforested after 2008.
 Our estimates of transitions show a decrease of deforested areas to be converted to agriculture after 2008, where it reached minimal values (Figure \ref{fig:transbar-plot}.a).
-After 2006, agriculture establishment over deforested primary forests suffered a decrease, which stayed relatively stable until 2012, where a steep increase occurred, however, the new areas being occupied by agriculture were mainly over areas that were cleared more than a decade before (therefore, before 2006) (Figure \ref{fig:transbar-plot}.b) This shows that agriculture expansion did not halt after the soy moratorium, producers started expanding in old cleared areas, which may have indirect impacts on deforestation increase over distant areas \citep{Arima2011}.
-However, this indirect effect is not always clear when analyzing the Amazon biome as a whole, our data shows that periods of fast agriculture expansion occurred at the same time with minimal deforestation values, according to PRODES.
-Expansion of agriculture areas also expanded over cleared areas of secondary forests in 2012, but with an important amount of fast transitions (Figure \citet{ref}(fig:transbar-plot).d).
-The causes of the increase of agriculture establishment areas can be numerous, one main driver was the approval of a new Forest Code, in 2012, which is considered to have undermined the environmental protection of forests \citep{Kroger2017, Pereira2019}.
+After 2006, agriculture establishment over deforested primary forests suffered a decrease, which stayed relatively stable until 2012, where a steep increase occurred, however, the new areas being occupied by agriculture were mainly over areas that were cleared more than a decade before (therefore, before 2006) (Figure \ref{fig:transbar-plot}.b) This shows that agriculture expansion did not halt after the soy moratorium, but producers expended over areas cleared many years ago, which may have indirect impacts on deforestation increase over distant areas \citep{Arima2011}.
+However, this indirect effect is not always clearly observed when analyzing the Amazon biome as a whole, but our data shows that periods of fast agriculture expansion occurred at the same time with minimal deforestation values, according to PRODES estimates (add reference).
+Expansion of agricultural areas also occurred over cleared areas of secondary forests in 2012, but with an important amount of fast transitions (Figure \ref{fig:transbar-plot}.d).
+The causes of the increase of agricultural establishment areas can be numerous, and one main driver was the approval of a new Forest Code, in 2012, which is considered to have undermined the environmental protection of forests \citep{Kroger2017, Pereira2019}.
 
@@
 \begin{figure}[ht]
-\includegraphics[width=17cm]{figs/trans_ridge} \caption{Transition length patterns inside states that belongs to the Amazon biome.
-Each year have a density estimate of the transition lengths, represented as colored curves.
+\includegraphics[width=17cm]{figs/trans_ridge} \caption{Transition length patterns inside states in the Amazon biome.
+Each year has a density estimate of the transition lengths, represented as colored curves.
 The peak of the curves represents transition length values with more frequency in one year of one state.
@@
 
-The state of Amapá presented fast transitions along all the time series, where slow transitions are not as common.
-In contrast, Acre shows a majority of slow transitions, in which only 2021 showed more fast transitions.
-There are three states where the pattern of transitions length across time are alike, Mato Grosso and Pará presented more fast transitions from 1995 to 2005, after this period, slow transitions became more common with time.
-Rondônia and Tocantins also presented a pattern in which transitions became slower with time, however this pattern started to occur earlier than Mato Grosso and Pará.
+The state of Amapá presented fast transitions along all the time series, without any significant period of slow transition.
+In contrast, transitions in Acre were majorly slow, in which only 2021 showed faster transitions.
+There are two states where the pattern of transition length are alike, Mato Grosso (MT) and Pará (PA).
+Both states underwent fast transitions from 1995 to 2005, and after this period slow transitions became predominant until the end of the time series.
+Rondônia (RO) and Tocantins (TO) presented a similar pattern of MT and PA, in which transitions decelerated , but for RO and TO since the beginning of 2000s, earlier than MA and PA.
 
-The great majority of transitions are from forests to Soybeans and ``Other Temporary Crops'', which represents 95\% of the transitions in the Amazon biome (14\% of Soybean and 81\% of ``Other Temporary Crops'').
-When analyzing what happened after 5 years since the transition, Soybean farms are persistent, 79\% of these areas remained as Soybean, 14\% is converted to ``Other Temporary Crops'' and 7\% in occupied by Pasture.
-``Other Temporary Crops'' are less persistent, 27\% of these areas remains as the same use, 56\% is converted to Soybean, and 15\% to Pasture.
-Conversions from Soybean and ``Other Temporary Uses'' to other covers and uses (apart from the already cited) are negligible.
-Sugar Cane, Cotton and Perennial Crops also appears, but in a negligible proportions.
+The great majority of transitions are from forests to soybean and "Other Temporary Crops", which represents 95\% of the transitions in the Amazon biome (14\% of Soybean and 81\% of "Other Temporary Crops").
+When analyzing what happened after 5 years since the transition, soybean areas persistency rate was at 79\% (i.e., areas that remained as soybean), 14\% was converted to "Other Temporary Crops" and 7\% was converted to pasture.
+"Other Temporary Crops" are less persistent.
+Hence, 27\% of these areas remain in the same class, 56\% were converted to soybean, and 15\% to pasture.
+Conversions from soybean and "Other Temporary Uses" to other LULC classes (apart from the already cited) are negligible.
+Sugar cane, cotton and perennial crops also appear, but in a negligible proportion.
 
@@
 From the 100 random sample points used in the results validation, 1 point (sample 21) was not considered as a transition from our estimations, however, the visual inspection pointed to a likely event of transition from forest to agriculture.
-Also, there were 6 points (sample 6, 7, 55, 56, 78, 97) which the visual inspection did not found a transition from forest to pasture, although our estimations pointed as transitions.
+Also, there were 6 points (sample 6, 7, 55, 56, 78, 97) which the visual inspection did not find a transition from forest to pasture, although our estimations pointed as transitions.
 Therefore, 7\% of the sample were completely misclassified by our estimates, and the rest of the accuracy assessment was performed over the remaining 93 sample points.
 
-When analyzing the errors from the transition length estimates, we observe that the year of deforestation shows the least amount of errors (Figure \ref{fig:errorbar-plot}).
-The MAE of the deforestation year is of 1.42 years, and the error shows a bias towards underestimation.
-The year of the agriculture establishment and the transition length estimates showed larger errors when comparing with visual inspection.
+When analyzing the errors from the transition length estimates, we observed that the year of deforestation shows the least amount of errors (Figure \ref{fig:errorbar-plot}).
+The MAE of the deforestation year is 1.42 years, and the error shows a bias towards underestimation.
+The year of the agriculture establishment and the transition length estimates showed larger errors when compared with visual inspection.
 
@@
 \includegraphics[width=17cm]{figs/error_bars} \caption{Bar with the count of error values (difference between observed and estimated values).
-Positive values indicate underestimation of the variable (estimations were lower than observations), negative values indicates overestimation (estimations higher than observations).
+Positive values indicate underestimation of the variable (estimations were lower than observations), negative values indicate overestimation (estimations higher than observations).
 Error metrics (Mean Absolute Error, Bias and Percent Bias) are displayed in the top right position of each box.
-The white deshed lines represents the MAE values of each variable.}\label{fig:errorbar-plot}
+The white dashed lines represent the MAE values of each variable.}\label{fig:errorbar-plot}
 \end{figure}
@@
 The Agriculture class presented more errors, the inclusion errors ranged from 22\% to 5\%, and were mostly composed by Forests pixels (forest pixels misclassified as agriculture).
-The omission errors of Agriculture ranged from 22\% to 8\%, and were also mostly composed by Forests pixels (agriculture pixels misclassified as Agriculture).
+The omission errors of Agriculture ranged from 22\% to 8\%, and were also mostly composed of Forest pixels (agriculture pixels misclassified as Agriculture).
 
@@
 Effect of border on results (sample 25, 56, 84, 92).
-Areas with sparse forest vegetation presented large error of deforestation year (sample 33).
+Areas with sparse forest vegetation presented a large error of deforestation year (sample 33).
 
-Patterns of agriculture areas are roughly well represented (sample 25, 43, 56, 84).
-Some locations presents a better homogeneity of transitions inside agriculture plots (sample 84).
+Patterns of agricultural areas are roughly well represented (sample 25, 43, 56, 84).
+Some locations present a better homogeneity of transitions inside agriculture plots (sample 84).
 
 Interference due to atmospheric conditions and lack of image availability clearly shaped spatial patterns of transitions at some locations.
-Quality of composites trough the time series affected accuracy of results.
-Regions of the North part of the Amazon presents lower quality than regions in the South of the Amazon biome.
+Quality of composites through the time series affected accuracy of results.
+Regions of the North part of the Amazon present lower quality than regions in the South of the Amazon biome.
 
 \conclusions[Conclusions]
+
+TODO: resumo ainda precisa de mais informações, por exemplo, a série temporal da análise, a escala (ex. todo o bioma Amazônia no Brasil, na América do Sul, apenas alguns estados da Amazônia).
+Falar do método, que tipo de álgebra de mapa foi usada, algum tipo de regressão\ldots enfim.
 

### 2022-11-30 scripts/figures/fig_1.R
Add new elements to the figure (state limits, inset box, subplots tags).

@@ -41,6 +41,10 @@ amazon <- read_biomes() %>%
   filter(code_biome == 1) %>%
   st_transform("EPSG:4326")
 
+states <- read_state() %>%
+  st_transform("EPSG:4326") %>%
+  st_intersection(amazon)
+
 # LOAD DATASETS ---------------------------------------------------------------
 
 ## Load transition length raster ----
@@ -83,21 +87,24 @@ trans_hex <- grid %>%
   st_intersection(amazon)
 
 ## Crop raster to extract full resolution values ----
+hr_bbox <-
+  st_as_sfc(
+    st_bbox(
+      c(
+        xmin = -52.3262,
+        xmax = -52.0786,
+        ymax =  -12.3993,
+        ymin =  -12.6592
+      ),
+      crs = st_crs(4326)
+    )
+  )
+
 hr_raster <-
   as.data.frame(
     crop(
       mosaic,
-      st_as_sfc(
-        st_bbox(
-          c(
-            xmin = -52.3262,
-            xmax = -52.0786,
-            ymax =  -12.3993,
-            ymin =  -12.6592
-          ),
-          crs = st_crs(4326)
-        )
-      )
+      hr_bbox
     ),
     xy = TRUE
   ) %>%
@@ -136,9 +143,19 @@ dist_map <- ggplot() +
     y_offset = 5
   ) +
   geom_sf(
+    data = states,
+    fill = "transparent",
+    color = "black"
+  ) +
+  geom_sf(
     data = trans_hex,
     aes(fill = trans_length),
     linewidth = 0.05
+  ) +
+  geom_sf(
+    data = hr_bbox,
+    fill = "transparent",
+    color = "#c50505"
   ) +
   ggtitle(
     "Distribution of transitions from forest to agriculture in the Amazon"
@@ -193,9 +210,12 @@ hist_plot <- ggplot() +
     legend.position = ""
   )
 
-full_map <- dist_map +
+full_map <- dist_map + theme(plot.tag.position = c(0.2, 0.8)) +
   inset_element(hr_map, 0.76, 0, 1, 0.34, align_to = "panel") +
-  inset_element(hist_plot, 0, 0, 0.5, 0.4, align_to = "panel")
+  theme(plot.tag.position = c(0.95, 1.08)) +
+  inset_element(hist_plot, 0, 0, 0.5, 0.4, align_to = "panel") +
+  theme(plot.tag.position = c(0.4, 0.4)) +
[…]

### 2022-12-01 manuscript/manuscript.Rmd
Change "transition" to "conversion". Improve abstract.Small changes to the introduction.

--- 2022-11-30 manuscript/manuscript.Rmd
@@
 ---
-title: Transition from forest to agriculture in the Brazilian Amazon from 1985 to 2021
+title: Conversion from forest to agriculture in the Brazilian Amazon from 1985 to 2021
 journal: essd
@@
 A large quantity of studies already relied on land-use classification to analyze this process.
-This study brings a new dataset created by calculating the time necessary for deforested areas to change to agriculture in the Brazilian Amazon biome.
+This study brings a new dataset created by calculating the time necessary for deforested areas to change to agriculture (annual and permanent crops) in the Brazilian Amazon biome.
+The calculations were performed over high resolution land-use classification by using simple algebra and recursion to identify every conversion from forest to agriculture, and ranges from 1985 to 2021.
 The new data can be useful in interdisciplinary studies about land-use and land-cover change analysis in Brazil, such as exploring its drivers and evaluating the effects on natural vegetation cover and ecosystem services.
-The major innovation brought by this study are links between the deforestation year of a given pixel with its respective ywar of agriculture establishment, which can provide new insights to understand long-term land transition processes in tropical ecosystems.
+The major innovation brought by this study are links between the deforestation year of a given pixel with its respective year of agriculture establishment, which can provide new insights to understand long-term land-use conversion processes in tropical ecosystems.
 
@@
 running:
-  title: Transition from forest to agriculture in the Brazilian Amazon from 1985 to 2021
+  title: Conversion from forest to agriculture in the Brazilian Amazon from 1985 to 2021
   author: Seixas et al.
@@
 
-Brazilian policies since the 1960s were focused on the expansion of the agricultural frontier in the Amazon, specially focused in in fostering economic growth and national security of the territory [@Carvalho2002; @Mcdonald2003; @Banerjee2009].
-Such development model led to the construction of thousands of kilometers of roads, and the settlement of large scale crop and livestock farms [@Carvalho2002; @Banerjee2009].
+Brazilian policies since the 1960s were focused on the expansion of the agricultural frontier in the Amazon, specially focused in fostering economic growth and national security of the territory [@Carvalho2002; @Mcdonald2003; @Banerjee2009].
+Such a development model led to the construction of thousands of kilometers of roads, and the settlement of large scale crop and livestock farms [@Carvalho2002; @Banerjee2009].
 This period was marked with high rates of deforestation of mature forests [@Fearnside2005].
 At the same time, cattle and soybean production started to move from Southern Brazil into the Midwest and Northern parts of the country, in which soybean cultivation areas followed cattle expansion (i.e., pastureland expansion) over the Amazon forest [@Simon2005; @Barona2010; @Arima2011].
-The pattern of extensive livestock production after deforestation, followed by the establishment of annual crops after a given period is typical in the Amazon biome, and persists nowadays (needs reference).
-This process, also known as "indirect land-use change" (needs reference) can take several decades to be accomplished, but sometimes less than one year.
-In this last case, the land change process can be considered as a direct transition (i.e., from forest to agriculture).
+The pattern of extensive livestock production after deforestation, followed by the establishment of annual crops after a given period is typical in the Amazon biome [@Barona2010], and persists nowadays.
+This process can take several decades to be accomplished, but sometimes less than one year.
+In this last case, the land change process can be considered as a direct conversion [@Morton2006] (i.e., from forest to agriculture).
 
 Land-use and land-cover (LULC) changes are known to cause impacts on different scales.
-Deforestation and land use changes can affect hydrological processes in large basins [@Arias2018]; reduce the forest resilience to extreme events (e.g., drought) and other sources of perturbation [@Boulton2022]; deforestation and farming practices can reduce convective rainfall and increase surface temperature [@Maeda2021]; increase carbon emissions [@Gatti2021]; cause important impacts on fauna and flora diversity, soil properties and carbon stocks [@Nunes2022]; and even negatively impact public health [@Ellwanger2020].
+Deforestation and land use changes can affect hydrological processes in large basins [@Arias2018]; reduce the forest resilience to extreme events (e.g., drought) and other sources of perturbation [@Boulton2022]; deforestation and farming practices can reduce convective rainfall and increase surface temperature [@Maeda2021]; increase carbon emissions [@Gatti2021]; cause important impacts on fauna and flora diversity, soil properties and carbon stocks [@Nunes2022; @Rittl2017]; and even negatively impact public health [@Ellwanger2020].
 
@@
 Initiatives such as MapBiomas [@Souza2020], launched in 2015, provides high resolution, annual classifications for the Brazilian territory.
-Such type of data are essential to analyse LULC change processes, make causal inferences, and understand what drives and what are the impacts of the transformation of the landscape in Brazil.
-In this context, this study aims to characterize and quantify the length of the transitions from forest formations to agriculture in the Amazon over the last 36 years.
+Such types of data are essential to analyze LULC change processes, make causal inferences, and understand what drives and what are the impacts of the transformation of the landscape in Brazil.
+In this context, this study aims to characterize and quantify the length of the conversions from forest formations to agriculture in the Amazon over the last 36 years.
 The availability of this data can be useful to the development of interdisciplinary research involved with LULC changes.
@@
 
-## Transition length calculation
+## Conversion length calculation
 
-The estimations of transition from forest to agriculture were performed for the Amazon biome region, as defined by the Brazilian Institute of Geography and Statistics (IBGE), in 2019.
+The estimations of conversion from forest to agriculture were performed for the Amazon biome region (biome limits defined by the Brazilian Institute of Geography and Statistics (IBGE), in 2019).
 Transitions were calculated using LULC classification data from MapBiomas, collection 7, which ranges from 1985 to 2021, at a spatial resolution of approximately 30 meters, and global accuracy at 91.3%.
@@
 
-The second part of the processing was to load each tile separately, convert it to a table and calculate every possible transition from forest to agriculture.
-The transitions were calculated pixel by pixel, by performing the following steps:
+The second part of the processing was to load each tile separately, convert it to a table and calculate every possible conversion from forest to agriculture.
+The conversions were calculated pixel by pixel, by performing the following steps:
 
@@
 
-3.  Calculate the first year of \"Forest Formation\" LULC class after the year calculated in step 2.
-    This step identifies the occurrence of more than one transition in a given pixel;
+3.  Calculate the first year of "Forest Formation" LULC class after the year calculated in step 2.
+    This step identifies the occurrence of more than one conversion in a given pixel;
 
-4.  Classify rows as \"before\" or \"after\" the occurrence of the year calculated in step 3.
-    This is the identification of periods before and after the recurrence of forest after a first transition to agriculture;
+4.  Classify rows as "before" or "after" the occurrence of the year calculated in step 3.
+    This is the identification of periods before and after the recurrence of forest after a first conversion to agriculture;
 
-5.  Calculate the last year of \"Forest Formation\" within the rows classified as \"before\", and add 1 year to represent the deforestation year;
+5.  Calculate the last year of "Forest Formation" within the rows classified as "before", and add 1 year to represent the deforestation year;
 
-6.  Calculate the first year of any agriculture type class within the rows classified as \"before\", for each pixel;
+6.  Calculate the first year of any agriculture type class within the rows classified as "before", for each pixel;
 
-7.  Calculate the difference between years from items 5 and 6 to get the LULC transition length in years, for each pixel;
+7.  Calculate the difference between years from items 5 and 6 to get the LULC conversion length in years, for each pixel;
 
-The steps 2 to 7 are performed recursively to identify multiple transitions, in case they are present.
-In addition to the transition length values and the years of the transition, transitions are also qualified by the type of agriculture that were established after deforestation, if the transition occurred from primary or secondary forests, and the number of recurrences of transitions at the pixel.
-The LULC classes within the transition and until 5 years after the transitions were also stored.
+The steps 2 to 7 are performed recursively to identify multiple conversions, in case they are present.
+In addition to the conversion length values and the years of the conversion, they are also qualified by the type of agriculture that were established after deforestation, if the conversion occurred from primary or secondary forests, and the number of recurrences of conversions at the pixel.
+The LULC classes within the conversion and until 5 years after were also stored.
 
-The transition results are stored in a data set of tabular files, where data can be queried for further analysis.
+The conversion results are stored in a data set of tabular files, where data can be queried for further analysis.
 After the calculations, data was also converted back to raster format, with the aim to expand the accessibility of the data set.
@@
 The median of Red, Green, Blue, Near Infrared (NIR) and Short Wave Infrared (SWIR1) from dry and wet season were used, and also the annual amplitude of the Normalized Difference Vegetation Index (NDVI).
-The process of accuracy assessment was performed in a Shiny app, and was conducted without any consultation to the transition length results.
+The process of accuracy assessment was performed in a Shiny app, and was conducted without any consultation to the conversion length results.
 In the validation app, we estimated, by visual inspection, the year of deforestation and the year of agricultural establishment.
-The observed transition length was obtained by subtracting both dates.
+The observed conversion length was obtained by subtracting both dates.
 
@@
 
-After the calculation of transitions, the results are stored in three different types of tables, organized in a folder structure and stored as Apache Parquet files.
+After the calculation of conversions, the results are stored in three different types of tables, organized in a folder structure and stored as Apache Parquet files.
 
 -   Tables that contain the spatial information (longitude, latitude) of each pixel, its unique id and the code of the municipality which contains the pixel.
-    It is named as \"mask_cells\";
+    It is named as "mask_cells";
 
--   Tables with transition length values, the first and last year of the transition, the resulting agriculture type, and the number of the transition cycle.
+-   Tables with conversion length values, the first and last year of the conversion, the resulting agriculture type, and the number of the conversion cycle.
     They also contain the unique id of each pixel (related to the table above);
 
--   Table with the LULC classes of all years within the transition, and also the first 5 years after the transition.
+-   Table with the LULC classes of all years within the conversion, and also the first 5 years after the conversion.
 
@@
 
-The transition calculations show that between 1985 and 2021, 64,874 square kilometers of forests were converted to agriculture in the Brazilian Amazon biome.
-The length of the transitions can go from 0 to 35 years, in which transitions closer to 0 years are considered as fast transitions, and transitions closer to 35 years are considered as slow transitions.
-Transitions of 0 years are considered \"direct\" transitions, where there was no presence of pasture before the establishment of agriculture.
-Our estimations show that around 9.2% of the transitions were \"direct\".
+The conversion calculations show that between 1985 and 2021, 64,874 square kilometers of forests were converted to agriculture in the Brazilian Amazon biome.
+The length of the conversions can go from 0 to 35 years, in which conversions closer to 0 years are considered as fast conversions, and conversions closer to 35 years are considered as slow conversions.
+Transitions of 0 years are considered "direct" conversions, where there was no presence of pasture before the establishment of agriculture.
+Our estimations show that around 9.2% of the conversions were "direct".
 
-Our transitions calculations found pixels that presented up to six transitions from forest to agriculture since 1985, as these numbers are unlikely to happen (they are distributed as sparse pixels, and do not show patterns of real transitions, being common at borders), we proceeded the results analysis only in the first transitions found in each of these pixels.
+Our conversions calculations found pixels that presented up to six conversions from forest to agriculture since 1985, as these numbers are unlikely to happen (they are distributed as sparse pixels, and do not show patterns of real conversions, being common at borders), we proceeded the results analysis only in the first conversions found in each of these pixels.
 
@@
 
-```{r map-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = "Map of distribution of transitions from forest to agriculture in the Brazilian Amazon biome.
-The hexagonal cells represent the most common transition length, and do not reflect the amount of area of transitions inside a cell.
+```{r map-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = "Map of distribution of conversions from forest to agriculture in the Brazilian Amazon biome.
+The hexagonal cells represent the most common conversion length, and do not reflect the amount of area of conversions inside a cell.
 Transitions are concentrated in the south (Mato Grosso state), and in the east (Pará and Maranhão states).
-The transition length ranges from 0 (blue tones) to 35 years (pink tones), and clusters of fast transitions (transitions closer to 1 year) can be discerned from clusters of slow transitions (transitions closer to 36 years).
-The histogram located in the bottom left shows that fast transitions are more common than slower transitions.
-The zoomed map in the bottom right shows the results in finer resolution, where it is possible to observe different transition lengths between properties."}
+The conversion length ranges from 0 (dark blue) to 35 years (light pink), and clusters of fast conversions (conversions closer to 1 year) can be discerned from clusters of slow conversions (conversions closer to 36 years).
+The histogram located in the bottom left shows that fast conversions are more common than slower conversions.
+The zoomed map in the bottom right shows the results in finer resolution, where it is possible to observe different conversion lengths between properties.
+The red square shows the extent of the zoomed map."}
 
@@
 
-Well defined clusters can be observed in the map created with aggregated transition length data (Figure \@ref(fig:map-plot)).
-Slow transition areas tend to concentrate in specific regions of the biome, while fast transitions seem to have a wider distribution, but also tend to form spatial clusters.
-However, this pattern does not hold completely when observing the data at its original scale (Figure \@ref(fig:map-plot)), where areas with different transition lengths are mixed between each other.
-When observing at the original scale, we could not spot any well defined pattern or direction of the occurrence of faster to slower transitions (Figure \@ref(fig:map-plot)).
+Well defined clusters can be observed in the map created with aggregated conversion length data (Figure \@ref(fig:map-plot)).
+Slow conversion areas tend to concentrate in specific regions of the biome, while fast conversions seem to have a wider distribution, but also tend to form spatial clusters.
+However, this pattern does not hold completely when observing the data at its original scale (Figure \@ref(fig:map-plot)), where areas with different conversion lengths are mixed between each other.
+When observing at the original scale, we could not spot any well defined pattern or direction of the occurrence of faster to slower conversions (Figure \@ref(fig:map-plot)).
 
-Other studies investigating patterns of transitions in the Amazon also found the formation of clusters of patterns, although there is a big heterogeneity at larger scales [@MullerHansen2017].
+Other studies investigating patterns of conversions in the Amazon also found the formation of clusters of patterns, although there is a big heterogeneity at larger scales [@MullerHansen2017].
 The data can be analyzed year by year, and also be separated by primary and secondary forests being converted to agriculture (Figure \@ref(fig:transbar-plot)).
 
-```{r transbar-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = "Transition area from forest to agriculture, per year and transition length.
-The bars represent the total amount of area at some state of the transition for each year.
-The color gradient in each bar represents the transition length related to a deforestation or an agriculture establishment event.
-Blue tones represents fast transitions (transitions closer to 0 years), pink tones represent slow transitions (transitions closer to 35 years).
+```{r transbar-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = "Transition area from forest to agriculture, per year and conversion length.
+The bars represent the total amount of area at some state of the conversion for each year.
+The color gradient in each bar represents the conversion length related to a deforestation or an agriculture establishment event.
+Blue tones represents fast conversions (closer to 0 years), pink tones represent slow conversions (closer to 35 years).
 Transition events were separated by deforestation of primary forests (a) and secondary forests (c), and the subsequent agriculture establishment of primary forests (b) and secondary forests (d)."}
@@
 The deforestation area of primary forests increased largely from 1986 to 2003, which was followed by a steep decrease until 2009, when the deforested areas reached a stable rate (Figure \@ref(fig:transbar-plot).a).
-From 1986 to approximately 1995, most of the deforested areas suffered a slow transition, mostly were higher than 10 years.
-After this period, fast transitions started to become more common, specially from 2002 to 2004 (Figure \@ref(fig:transbar-plot).a).
-The peak of deforestation in 2003, according to our results, shows that 12.6% of the transitions were a direct conversion from forest to agriculture.
+From 1986 to approximately 1995, most of the deforested areas suffered a slow conversion, mostly were higher than 10 years.
+After this period, fast conversions started to become more common, specially from 2002 to 2004 (Figure \@ref(fig:transbar-plot).a).
+The peak of deforestation in 2003, according to our results, shows that 12.6% of the conversions were a direct conversion from forest to agriculture.
 Previous studies estimated a proportion of 23% [@Morton2006].
-If we include fast transitions (from 0 to 2 years) from our results, the proportion jumps to 48.5 % of all transitions.
+If we include fast conversions (from 0 to 2 years) from our results, the proportion jumps to 48.5 % of all conversions.
 
 Agricultural establishment over areas of primary forests peaked in 2005 and 2013 (Figure \@ref(fig:transbar-plot).b).
-Despite similar agricultural establishment rates between both years, their transition lengths differ greatly, in 2005 most of the transitions were faster than 10 years, while in 2013 the great majority of transitions were slower than 10 years (Figure \@ref(fig:transbar-plot).b).
-In 2003 it was observed a shift in the transition length from forest until the establishment of agricultural areas.
-From this year onwards, most transitions happened in areas deforested at least 10 years before (Figure \@ref(fig:transbar-plot).b).
+Despite similar agricultural establishment rates between both years, their conversion lengths differ greatly, in 2005 most of the conversions were faster than 10 years, while in 2013 the great majority of conversions were slower than 10 years (Figure \@ref(fig:transbar-plot).b).
+In 2003 it was observed a shift in the conversion length from forest until the establishment of agricultural areas.
+From this year onward, most conversions happened in areas deforested at least 10 years before (Figure \@ref(fig:transbar-plot).b).
 Even after the decrease of deforestation after 2002, areas under agriculture expanded over lands where deforestation happened before 2002 (mostly over pastureland).
@@
 In the late 1990s, the development policies "Brasil em Ação" and "Avança Brasil" (1995 - 2003) accelerated the national infrastructure expansion, including in the Amazon [@Carvalho2002].
-This is a period when the deforestation areas to be converted to agriculture reached the highest values in the time series (Figure \@ref(fig:transbar-plot).a), and also a peak in the establishment of agriculture in the Amazon biome, predominantly with fast transitions (Figure \@ref(fig:transbar-plot).b).
+This is a period when the deforestation areas to be converted to agriculture reached the highest values in the time series (Figure \@ref(fig:transbar-plot).a), and also a peak in the establishment of agriculture in the Amazon biome, predominantly with fast conversions (Figure \@ref(fig:transbar-plot).b).
 
@@
 In 2006, the Brazilian Association of Vegetable Oil Industries (BIOVE) and the National Association of Cereal Exporters (ANEC) committed to avoid commercialization of soybean grains harvested from areas deforested after 2008.
-Our estimates of transitions show a decrease of deforested areas to be converted to agriculture after 2008, where it reached minimal values (Figure \@ref(fig:transbar-plot).a).
+Our estimates of conversions show a decrease of deforested areas to be converted to agriculture after 2008, where it reached minimal values (Figure \@ref(fig:transbar-plot).a).
 After 2006, agriculture establishment over deforested primary forests suffered a decrease, which stayed relatively stable until 2012, where a steep increase occurred, however, the new areas being occupied by agriculture were mainly over areas that were cleared more than a decade before (therefore, before 2006) (Figure \@ref(fig:transbar-plot).b) This shows that agriculture expansion did not halt after the soy moratorium, but producers expended over areas cleared many years ago, which may have indirect impacts on deforestation increase over distant areas [@Arima2011].
 However, this indirect effect is not always clearly observed when analyzing the Amazon biome as a whole, but our data shows that periods of fast agriculture expansion occurred at the same time with minimal deforestation values, according to PRODES estimates (add reference).
-Expansion of agricultural areas also occurred over cleared areas of secondary forests in 2012, but with an important amount of fast transitions (Figure @ref(fig:transbar-plot).d).
+Expansion of agricultural areas also occurred over cleared areas of secondary forests in 2012, but with an important amount of fast conversions (Figure \@ref(fig:transbar-plot).d).
 The causes of the increase of agricultural establishment areas can be numerous, and one main driver was the approval of a new Forest Code, in 2012, which is considered to have undermined the environmental protection of forests [@Kroger2017; @Pereira2019].
 
-The transition length patterns across years can change significantly between different states (Figure \@ref(fig:transridge-plot)).
+The conversion length patterns across years can change significantly between different states (Figure \@ref(fig:transridge-plot)).
 
 ```{r transridge-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = "Transition length patterns inside states in the Amazon biome.
-Each year has a density estimate of the transition lengths, represented as colored curves.
-The peak of the curves represents transition length values with more frequency in one year of one state.
-Blue tones represents fast transitions (transitions closer to 0 years), pink tones represents slow transitions (transitions closer to 35 years)"}
+Each year has a density estimate of the conversion lengths, represented as colored curves.
+The peak of the curves represents conversion length values with more frequency in one year of one state.
+Blue tones represents fast conversions (conversions closer to 0 years), pink tones represents slow conversions (conversions closer to 35 years)"}
 
@@
 
-The state of Amapá presented fast transitions along all the time series, without any significant period of slow transition.
-In contrast, transitions in Acre were majorly slow, in which only 2021 showed faster transitions.
-There are two states where the pattern of transition length are alike, Mato Grosso (MT) and Pará (PA).
-Both states underwent fast transitions from 1995 to 2005, and after this period slow transitions became predominant until the end of the time series.
-Rondônia (RO) and Tocantins (TO) presented a similar pattern of MT and PA, in which transitions decelerated , but for RO and TO since the beginning of 2000s, earlier than MA and PA.
+The state of Amapá presented fast conversions along all the time series, without any significant period of slow conversion.
+In contrast, conversions in Acre were majorly slow, in which only 2021 showed faster conversions.
+There are two states where the pattern of conversion length are alike, Mato Grosso (MT) and Pará (PA).
+Both states underwent fast conversions from 1995 to 2005, and after this period slow conversions became predominant until the end of the time series.
+Rondônia (RO) and Tocantins (TO) presented a similar pattern of MT and PA, in which conversions decelerated , but for RO and TO since the beginning of 2000s, earlier than MA and PA.
 
-The great majority of transitions are from forests to soybean and \"Other Temporary Crops\", which represents 95% of the transitions in the Amazon biome (14% of Soybean and 81% of \"Other Temporary Crops\").
-When analyzing what happened after 5 years since the transition, soybean areas persistency rate was at 79% (i.e., areas that remained as soybean), 14% was converted to \"Other Temporary Crops\" and 7% was converted to pasture.
-\"Other Temporary Crops\" are less persistent.
+The great majority of conversions are from forests to soybean and "Other Temporary Crops", which represents 95% of the conversions in the Amazon biome (14% of Soybean and 81% of "Other Temporary Crops").
+When analyzing what happened after 5 years since the conversion, soybean areas persistence rate was at 79% (i.e., areas that remained as soybean), 14% was converted to "Other Temporary Crops" and 7% was converted to pasture.
+"Other Temporary Crops" are less persistent.
 Hence, 27% of these areas remain in the same class, 56% were converted to soybean, and 15% to pasture.
-Conversions from soybean and \"Other Temporary Uses\" to other LULC classes (apart from the already cited) are negligible.
+Conversions from soybean and "Other Temporary Uses" to other LULC classes (apart from the already cited) are negligible.
 Sugar cane, cotton and perennial crops also appear, but in a negligible proportion.
@@
 
-From the 100 random sample points used in the results validation, 1 point (sample 21) was not considered as a transition from our estimations, however, the visual inspection pointed to a likely event of transition from forest to agriculture.
-Also, there were 6 points (sample 6, 7, 55, 56, 78, 97) which the visual inspection did not find a transition from forest to pasture, although our estimations pointed as transitions.
+From the 100 random sample points used in the results validation, 1 point (sample 21) was not considered as a conversion from our estimations, however, the visual inspection pointed to a likely event of conversion from forest to agriculture.
+Also, there were 6 points (sample 6, 7, 55, 56, 78, 97) which the visual inspection did not find a conversion from forest to pasture, although our estimations pointed as conversions.
 Therefore, 7% of the sample were completely misclassified by our estimates, and the rest of the accuracy assessment was performed over the remaining 93 sample points.
 
-When analyzing the errors from the transition length estimates, we observed that the year of deforestation shows the least amount of errors (Figure \@ref(fig:errorbar-plot)).
+When analyzing the errors from the conversion length estimates, we observed that the year of deforestation shows the least amount of errors (Figure \@ref(fig:errorbar-plot)).
 The MAE of the deforestation year is 1.42 years, and the error shows a bias towards underestimation.
-The year of the agriculture establishment and the transition length estimates showed larger errors when compared with visual inspection.
+The year of the agriculture establishment and the conversion length estimates showed larger errors when compared with visual inspection.
 
@@
 Transition length values are more concentrated in smaller values, which also shows higher errors (Figure \@ref(fig:errorscatter-plot).c).
-However, this is expected since faster transitions are more common (Figure \@ref(fig:map-plot)).
+However, this is expected since faster conversions are more common (Figure \@ref(fig:map-plot)).
 
@@
 Roads being considered as agriculture (sample 2).
-Omissions were detected, specially agriculture plots not being considered as a transition as a whole, parts of a plot that were clearly a transition were not considered as a transition (sample 16, 30, 75).
-Inclusions were detected, areas that clearly showed no agriculture were considered as transitions (sample 10, 98).
+Omissions were detected, specially agriculture plots not being considered as a conversion as a whole, parts of a plot that were clearly a conversion were not considered as a conversion (sample 16, 30, 75).
+Inclusions were detected, areas that clearly showed no agriculture were considered as conversions (sample 10, 98).
 Effect of border on results (sample 25, 56, 84, 92).
@@
 Patterns of agricultural areas are roughly well represented (sample 25, 43, 56, 84).
-Some locations present a better homogeneity of transitions inside agriculture plots (sample 84).
+Some locations present a better homogeneity of conversions inside agriculture plots (sample 84).
 
-Interference due to atmospheric conditions and lack of image availability clearly shaped spatial patterns of transitions at some locations.
+Interference due to atmospheric conditions and lack of image availability clearly shaped spatial patterns of conversions at some locations.
 Quality of composites through the time series affected accuracy of results.

### 2022-12-01 manuscript/manuscript.tex
Change "transition" to "conversion". Improve abstract.Small changes to the introduction.

@@
 
-\title{Transition from forest to agriculture in the Brazilian Amazon from 1985 to 2021}
+\title{Conversion from forest to agriculture in the Brazilian Amazon from 1985 to 2021}
 
@@
 
-\runningtitle{Transition from forest to agriculture in the Brazilian Amazon from 1985 to 2021}
+\runningtitle{Conversion from forest to agriculture in the Brazilian Amazon from 1985 to 2021}
 
@@
 A large quantity of studies already relied on land-use classification to analyze this process.
-This study brings a new dataset created by calculating the time necessary for deforested areas to change to agriculture in the Brazilian Amazon biome.
+This study brings a new dataset created by calculating the time necessary for deforested areas to change to agriculture (annual and permanent crops) in the Brazilian Amazon biome.
+The calculations were performed over high resolution land-use classification by using simple algebra and recursion to identify every conversion from forest to agriculture, and ranges from 1985 to 2021.
 The new data can be useful in interdisciplinary studies about land-use and land-cover change analysis in Brazil, such as exploring its drivers and evaluating the effects on natural vegetation cover and ecosystem services.
-The major innovation brought by this study are links between the deforestation year of a given pixel with its respective ywar of agriculture establishment, which can provide new insights to understand long-term land transition processes in tropical ecosystems.
+The major innovation brought by this study are links between the deforestation year of a given pixel with its respective year of agriculture establishment, which can provide new insights to understand long-term land-use conversion processes in tropical ecosystems.
 \end{abstract}
@@
 
-Brazilian policies since the 1960s were focused on the expansion of the agricultural frontier in the Amazon, specially focused in in fostering economic growth and national security of the territory \citep{Carvalho2002, Mcdonald2003, Banerjee2009}.
-Such development model led to the construction of thousands of kilometers of roads, and the settlement of large scale crop and livestock farms \citep{Carvalho2002, Banerjee2009}.
+Brazilian policies since the 1960s were focused on the expansion of the agricultural frontier in the Amazon, specially focused in fostering economic growth and national security of the territory \citep{Carvalho2002, Mcdonald2003, Banerjee2009}.
+Such a development model led to the construction of thousands of kilometers of roads, and the settlement of large scale crop and livestock farms \citep{Carvalho2002, Banerjee2009}.
 This period was marked with high rates of deforestation of mature forests \citep{Fearnside2005}.
 At the same time, cattle and soybean production started to move from Southern Brazil into the Midwest and Northern parts of the country, in which soybean cultivation areas followed cattle expansion (i.e., pastureland expansion) over the Amazon forest \citep{Simon2005, Barona2010, Arima2011}.
-The pattern of extensive livestock production after deforestation, followed by the establishment of annual crops after a given period is typical in the Amazon biome, and persists nowadays (needs reference).
-This process, also known as ``indirect land-use change'' (needs reference) can take several decades to be accomplished, but sometimes less than one year.
-In this last case, the land change process can be considered as a direct transition (i.e., from forest to agriculture).
+The pattern of extensive livestock production after deforestation, followed by the establishment of annual crops after a given period is typical in the Amazon biome \citep{Barona2010}, and persists nowadays.
+This process can take several decades to be accomplished, but sometimes less than one year.
+In this last case, the land change process can be considered as a direct conversion \citep{Morton2006} (i.e., from forest to agriculture).
 
 Land-use and land-cover (LULC) changes are known to cause impacts on different scales.
-Deforestation and land use changes can affect hydrological processes in large basins \citep{Arias2018}; reduce the forest resilience to extreme events (e.g., drought) and other sources of perturbation \citep{Boulton2022}; deforestation and farming practices can reduce convective rainfall and increase surface temperature \citep{Maeda2021}; increase carbon emissions \citep{Gatti2021}; cause important impacts on fauna and flora diversity, soil properties and carbon stocks \citep{Nunes2022}; and even negatively impact public health \citep{Ellwanger2020}.
+Deforestation and land use changes can affect hydrological processes in large basins \citep{Arias2018}; reduce the forest resilience to extreme events (e.g., drought) and other sources of perturbation \citep{Boulton2022}; deforestation and farming practices can reduce convective rainfall and increase surface temperature \citep{Maeda2021}; increase carbon emissions \citep{Gatti2021}; cause important impacts on fauna and flora diversity, soil properties and carbon stocks \citep{Nunes2022, Rittl2017}; and even negatively impact public health \citep{Ellwanger2020}.
 
@@
 Initiatives such as MapBiomas \citep{Souza2020}, launched in 2015, provides high resolution, annual classifications for the Brazilian territory.
-Such type of data are essential to analyse LULC change processes, make causal inferences, and understand what drives and what are the impacts of the transformation of the landscape in Brazil.
-In this context, this study aims to characterize and quantify the length of the transitions from forest formations to agriculture in the Amazon over the last 36 years.
+Such types of data are essential to analyze LULC change processes, make causal inferences, and understand what drives and what are the impacts of the transformation of the landscape in Brazil.
+In this context, this study aims to characterize and quantify the length of the conversions from forest formations to agriculture in the Amazon over the last 36 years.
 The availability of this data can be useful to the development of interdisciplinary research involved with LULC changes.
@@
 
-\subsection{Transition length calculation}
+\subsection{Conversion length calculation}
 
-The estimations of transition from forest to agriculture were performed for the Amazon biome region, as defined by the Brazilian Institute of Geography and Statistics (IBGE), in 2019.
+The estimations of conversion from forest to agriculture were performed for the Amazon biome region (biome limits defined by the Brazilian Institute of Geography and Statistics (IBGE), in 2019).
 Transitions were calculated using LULC classification data from MapBiomas, collection 7, which ranges from 1985 to 2021, at a spatial resolution of approximately 30 meters, and global accuracy at 91.3\%.
@@
 \item
-  Calculate the first year of "Forest Formation" LULC class after the year calculated in step 2.
-  This step identifies the occurrence of more than one transition in a given pixel;
+  Calculate the first year of ``Forest Formation'' LULC class after the year calculated in step 2.
+  This step identifies the occurrence of more than one conversion in a given pixel;
 \item
-  Classify rows as "before" or "after" the occurrence of the year calculated in step 3.
-  This is the identification of periods before and after the recurrence of forest after a first transition to agriculture;
+  Classify rows as ``before'' or ``after'' the occurrence of the year calculated in step 3.
+  This is the identification of periods before and after the recurrence of forest after a first conversion to agriculture;
 \item
-  Calculate the last year of "Forest Formation" within the rows classified as "before", and add 1 year to represent the deforestation year;
+  Calculate the last year of ``Forest Formation'' within the rows classified as ``before'', and add 1 year to represent the deforestation year;
 \item
-  Calculate the first year of any agriculture type class within the rows classified as "before", for each pixel;
+  Calculate the first year of any agriculture type class within the rows classified as ``before'', for each pixel;
 \item
-  Calculate the difference between years from items 5 and 6 to get the LULC transition length in years, for each pixel;
+  Calculate the difference between years from items 5 and 6 to get the LULC conversion length in years, for each pixel;
 \end{enumerate}
 
-The steps 2 to 7 are performed recursively to identify multiple transitions, in case they are present.
-In addition to the transition length values and the years of the transition, transitions are also qualified by the type of agriculture that were established after deforestation, if the transition occurred from primary or secondary forests, and the number of recurrences of transitions at the pixel.
-The LULC classes within the transition and until 5 years after the transitions were also stored.
+The steps 2 to 7 are performed recursively to identify multiple conversions, in case they are present.
+In addition to the conversion length values and the years of the conversion, they are also qualified by the type of agriculture that were established after deforestation, if the conversion occurred from primary or secondary forests, and the number of recurrences of conversions at the pixel.
+The LULC classes within the conversion and until 5 years after were also stored.
 
-The transition results are stored in a data set of tabular files, where data can be queried for further analysis.
+The conversion results are stored in a data set of tabular files, where data can be queried for further analysis.
 After the calculations, data was also converted back to raster format, with the aim to expand the accessibility of the data set.
@@
 
-After the calculation of transitions, the results are stored in three different types of tables, organized in a folder structure and stored as Apache Parquet files.
+After the calculation of conversions, the results are stored in three different types of tables, organized in a folder structure and stored as Apache Parquet files.
 
@@
   Tables that contain the spatial information (longitude, latitude) of each pixel, its unique id and the code of the municipality which contains the pixel.
-  It is named as "mask\_cells";
+  It is named as ``mask\_cells'';
 \item
-  Tables with transition length values, the first and last year of the transition, the resulting agriculture type, and the number of the transition cycle.
+  Tables with conversion length values, the first and last year of the conversion, the resulting agriculture type, and the number of the conversion cycle.
   They also contain the unique id of each pixel (related to the table above);
 \item
-  Table with the LULC classes of all years within the transition, and also the first 5 years after the transition.
+  Table with the LULC classes of all years within the conversion, and also the first 5 years after the conversion.
 \end{itemize}
@@
 
-The transition calculations show that between 1985 and 2021, 64,874 square kilometers of forests were converted to agriculture in the Brazilian Amazon biome.
-The length of the transitions can go from 0 to 35 years, in which transitions closer to 0 years are considered as fast transitions, and transitions closer to 35 years are considered as slow transitions.
-Transitions of 0 years are considered "direct" transitions, where there was no presence of pasture before the establishment of agriculture.
-Our estimations show that around 9.2\% of the transitions were "direct".
+The conversion calculations show that between 1985 and 2021, 64,874 square kilometers of forests were converted to agriculture in the Brazilian Amazon biome.
+The length of the conversions can go from 0 to 35 years, in which conversions closer to 0 years are considered as fast conversions, and conversions closer to 35 years are considered as slow conversions.
+Transitions of 0 years are considered ``direct'' conversions, where there was no presence of pasture before the establishment of agriculture.
+Our estimations show that around 9.2\% of the conversions were ``direct''.
 
-Our transitions calculations found pixels that presented up to six transitions from forest to agriculture since 1985, as these numbers are unlikely to happen (they are distributed as sparse pixels, and do not show patterns of real transitions, being common at borders), we proceeded the results analysis only in the first transitions found in each of these pixels.
+Our conversions calculations found pixels that presented up to six conversions from forest to agriculture since 1985, as these numbers are unlikely to happen (they are distributed as sparse pixels, and do not show patterns of real conversions, being common at borders), we proceeded the results analysis only in the first conversions found in each of these pixels.
 
@@
 \begin{figure}[ht]
-\includegraphics[width=17cm]{figs/map} \caption{Map of distribution of transitions from forest to agriculture in the Brazilian Amazon biome.
-The hexagonal cells represent the most common transition length, and do not reflect the amount of area of transitions inside a cell.
+\includegraphics[width=17cm]{figs/map} \caption{Map of distribution of conversions from forest to agriculture in the Brazilian Amazon biome.
+The hexagonal cells represent the most common conversion length, and do not reflect the amount of area of conversions inside a cell.
 Transitions are concentrated in the south (Mato Grosso state), and in the east (Pará and Maranhão states).
-The transition length ranges from 0 (blue tones) to 35 years (pink tones), and clusters of fast transitions (transitions closer to 1 year) can be discerned from clusters of slow transitions (transitions closer to 36 years).
-The histogram located in the bottom left shows that fast transitions are more common than slower transitions.
-The zoomed map in the bottom right shows the results in finer resolution, where it is possible to observe different transition lengths between properties.}\label{fig:map-plot}
+The conversion length ranges from 0 (dark blue) to 35 years (light pink), and clusters of fast conversions (conversions closer to 1 year) can be discerned from clusters of slow conversions (conversions closer to 36 years).
+The histogram located in the bottom left shows that fast conversions are more common than slower conversions.
+The zoomed map in the bottom right shows the results in finer resolution, where it is possible to observe different conversion lengths between properties.
+The red square shows the extent of the zoomed map.}\label{fig:map-plot}
 \end{figure}
 
-Well defined clusters can be observed in the map created with aggregated transition length data (Figure \ref{fig:map-plot}).
-Slow transition areas tend to concentrate in specific regions of the biome, while fast transitions seem to have a wider distribution, but also tend to form spatial clusters.
-However, this pattern does not hold completely when observing the data at its original scale (Figure \ref{fig:map-plot}), where areas with different transition lengths are mixed between each other.
-When observing at the original scale, we could not spot any well defined pattern or direction of the occurrence of faster to slower transitions (Figure \ref{fig:map-plot}).
+Well defined clusters can be observed in the map created with aggregated conversion length data (Figure \ref{fig:map-plot}).
+Slow conversion areas tend to concentrate in specific regions of the biome, while fast conversions seem to have a wider distribution, but also tend to form spatial clusters.
+However, this pattern does not hold completely when observing the data at its original scale (Figure \ref{fig:map-plot}), where areas with different conversion lengths are mixed between each other.
+When observing at the original scale, we could not spot any well defined pattern or direction of the occurrence of faster to slower conversions (Figure \ref{fig:map-plot}).
 
-Other studies investigating patterns of transitions in the Amazon also found the formation of clusters of patterns, although there is a big heterogeneity at larger scales \citep{MullerHansen2017}.
+Other studies investigating patterns of conversions in the Amazon also found the formation of clusters of patterns, although there is a big heterogeneity at larger scales \citep{MullerHansen2017}.
 The data can be analyzed year by year, and also be separated by primary and secondary forests being converted to agriculture (Figure \ref{fig:transbar-plot}).
@@
 \begin{figure}[ht]
-\includegraphics[width=17cm]{figs/trans_length_cols} \caption{Transition area from forest to agriculture, per year and transition length.
-The bars represent the total amount of area at some state of the transition for each year.
-The color gradient in each bar represents the transition length related to a deforestation or an agriculture establishment event.
-Blue tones represents fast transitions (transitions closer to 0 years), pink tones represent slow transitions (transitions closer to 35 years).
+\includegraphics[width=17cm]{figs/trans_length_cols} \caption{Transition area from forest to agriculture, per year and conversion length.
+The bars represent the total amount of area at some state of the conversion for each year.
+The color gradient in each bar represents the conversion length related to a deforestation or an agriculture establishment event.
+Blue tones represents fast conversions (closer to 0 years), pink tones represent slow conversions (closer to 35 years).
 Transition events were separated by deforestation of primary forests (a) and secondary forests (c), and the subsequent agriculture establishment of primary forests (b) and secondary forests (d).}\label{fig:transbar-plot}
@@
 The deforestation area of primary forests increased largely from 1986 to 2003, which was followed by a steep decrease until 2009, when the deforested areas reached a stable rate (Figure \ref{fig:transbar-plot}.a).
-From 1986 to approximately 1995, most of the deforested areas suffered a slow transition, mostly were higher than 10 years.
-After this period, fast transitions started to become more common, specially from 2002 to 2004 (Figure \ref{fig:transbar-plot}.a).
-The peak of deforestation in 2003, according to our results, shows that 12.6\% of the transitions were a direct conversion from forest to agriculture.
+From 1986 to approximately 1995, most of the deforested areas suffered a slow conversion, mostly were higher than 10 years.
+After this period, fast conversions started to become more common, specially from 2002 to 2004 (Figure \ref{fig:transbar-plot}.a).
+The peak of deforestation in 2003, according to our results, shows that 12.6\% of the conversions were a direct conversion from forest to agriculture.
 Previous studies estimated a proportion of 23\% \citep{Morton2006}.
-If we include fast transitions (from 0 to 2 years) from our results, the proportion jumps to 48.5 \% of all transitions.
+If we include fast conversions (from 0 to 2 years) from our results, the proportion jumps to 48.5 \% of all conversions.
 
 Agricultural establishment over areas of primary forests peaked in 2005 and 2013 (Figure \ref{fig:transbar-plot}.b).
-Despite similar agricultural establishment rates between both years, their transition lengths differ greatly, in 2005 most of the transitions were faster than 10 years, while in 2013 the great majority of transitions were slower than 10 years (Figure \ref{fig:transbar-plot}.b).
-In 2003 it was observed a shift in the transition length from forest until the establishment of agricultural areas.
-From this year onwards, most transitions happened in areas deforested at least 10 years before (Figure \ref{fig:transbar-plot}.b).
+Despite similar agricultural establishment rates between both years, their conversion lengths differ greatly, in 2005 most of the conversions were faster than 10 years, while in 2013 the great majority of conversions were slower than 10 years (Figure \ref{fig:transbar-plot}.b).
+In 2003 it was observed a shift in the conversion length from forest until the establishment of agricultural areas.
+From this year onward, most conversions happened in areas deforested at least 10 years before (Figure \ref{fig:transbar-plot}.b).
 Even after the decrease of deforestation after 2002, areas under agriculture expanded over lands where deforestation happened before 2002 (mostly over pastureland).
@@
 In the late 1990s, the development policies ``Brasil em Ação'' and ``Avança Brasil'' (1995 - 2003) accelerated the national infrastructure expansion, including in the Amazon \citep{Carvalho2002}.
-This is a period when the deforestation areas to be converted to agriculture reached the highest values in the time series (Figure \ref{fig:transbar-plot}.a), and also a peak in the establishment of agriculture in the Amazon biome, predominantly with fast transitions (Figure \ref{fig:transbar-plot}.b).
+This is a period when the deforestation areas to be converted to agriculture reached the highest values in the time series (Figure \ref{fig:transbar-plot}.a), and also a peak in the establishment of agriculture in the Amazon biome, predominantly with fast conversions (Figure \ref{fig:transbar-plot}.b).
 
@@
 In 2006, the Brazilian Association of Vegetable Oil Industries (BIOVE) and the National Association of Cereal Exporters (ANEC) committed to avoid commercialization of soybean grains harvested from areas deforested after 2008.
-Our estimates of transitions show a decrease of deforested areas to be converted to agriculture after 2008, where it reached minimal values (Figure \ref{fig:transbar-plot}.a).
+Our estimates of conversions show a decrease of deforested areas to be converted to agriculture after 2008, where it reached minimal values (Figure \ref{fig:transbar-plot}.a).
 After 2006, agriculture establishment over deforested primary forests suffered a decrease, which stayed relatively stable until 2012, where a steep increase occurred, however, the new areas being occupied by agriculture were mainly over areas that were cleared more than a decade before (therefore, before 2006) (Figure \ref{fig:transbar-plot}.b) This shows that agriculture expansion did not halt after the soy moratorium, but producers expended over areas cleared many years ago, which may have indirect impacts on deforestation increase over distant areas \citep{Arima2011}.
 However, this indirect effect is not always clearly observed when analyzing the Amazon biome as a whole, but our data shows that periods of fast agriculture expansion occurred at the same time with minimal deforestation values, according to PRODES estimates (add reference).
-Expansion of agricultural areas also occurred over cleared areas of secondary forests in 2012, but with an important amount of fast transitions (Figure \ref{fig:transbar-plot}.d).
+Expansion of agricultural areas also occurred over cleared areas of secondary forests in 2012, but with an important amount of fast conversions (Figure \ref{fig:transbar-plot}.d).
 The causes of the increase of agricultural establishment areas can be numerous, and one main driver was the approval of a new Forest Code, in 2012, which is considered to have undermined the environmental protection of forests \citep{Kroger2017, Pereira2019}.
 
-The transition length patterns across years can change significantly between different states (Figure \ref{fig:transridge-plot}).
+The conversion length patterns across years can change significantly between different states (Figure \ref{fig:transridge-plot}).
 
@@
 \includegraphics[width=17cm]{figs/trans_ridge} \caption{Transition length patterns inside states in the Amazon biome.
-Each year has a density estimate of the transition lengths, represented as colored curves.
-The peak of the curves represents transition length values with more frequency in one year of one state.
-Blue tones represents fast transitions (transitions closer to 0 years), pink tones represents slow transitions (transitions closer to 35 years)}\label{fig:transridge-plot}
+Each year has a density estimate of the conversion lengths, represented as colored curves.
+The peak of the curves represents conversion length values with more frequency in one year of one state.
+Blue tones represents fast conversions (conversions closer to 0 years), pink tones represents slow conversions (conversions closer to 35 years)}\label{fig:transridge-plot}
 \end{figure}
 
-The state of Amapá presented fast transitions along all the time series, without any significant period of slow transition.
-In contrast, transitions in Acre were majorly slow, in which only 2021 showed faster transitions.
-There are two states where the pattern of transition length are alike, Mato Grosso (MT) and Pará (PA).
-Both states underwent fast transitions from 1995 to 2005, and after this period slow transitions became predominant until the end of the time series.
-Rondônia (RO) and Tocantins (TO) presented a similar pattern of MT and PA, in which transitions decelerated , but for RO and TO since the beginning of 2000s, earlier than MA and PA.
+The state of Amapá presented fast conversions along all the time series, without any significant period of slow conversion.
+In contrast, conversions in Acre were majorly slow, in which only 2021 showed faster conversions.
+There are two states where the pattern of conversion length are alike, Mato Grosso (MT) and Pará (PA).
+Both states underwent fast conversions from 1995 to 2005, and after this period slow conversions became predominant until the end of the time series.
+Rondônia (RO) and Tocantins (TO) presented a similar pattern of MT and PA, in which conversions decelerated , but for RO and TO since the beginning of 2000s, earlier than MA and PA.
 
-The great majority of transitions are from forests to soybean and "Other Temporary Crops", which represents 95\% of the transitions in the Amazon biome (14\% of Soybean and 81\% of "Other Temporary Crops").
-When analyzing what happened after 5 years since the transition, soybean areas persistency rate was at 79\% (i.e., areas that remained as soybean), 14\% was converted to "Other Temporary Crops" and 7\% was converted to pasture.
-"Other Temporary Crops" are less persistent.
+The great majority of conversions are from forests to soybean and ``Other Temporary Crops'', which represents 95\% of the conversions in the Amazon biome (14\% of Soybean and 81\% of ``Other Temporary Crops'').
+When analyzing what happened after 5 years since the conversion, soybean areas persistence rate was at 79\% (i.e., areas that remained as soybean), 14\% was converted to ``Other Temporary Crops'' and 7\% was converted to pasture.
+``Other Temporary Crops'' are less persistent.
 Hence, 27\% of these areas remain in the same class, 56\% were converted to soybean, and 15\% to pasture.
-Conversions from soybean and "Other Temporary Uses" to other LULC classes (apart from the already cited) are negligible.
+Conversions from soybean and ``Other Temporary Uses'' to other LULC classes (apart from the already cited) are negligible.
 Sugar cane, cotton and perennial crops also appear, but in a negligible proportion.
@@
 
-From the 100 random sample points used in the results validation, 1 point (sample 21) was not considered as a transition from our estimations, however, the visual inspection pointed to a likely event of transition from forest to agriculture.
-Also, there were 6 points (sample 6, 7, 55, 56, 78, 97) which the visual inspection did not find a transition from forest to pasture, although our estimations pointed as transitions.
+From the 100 random sample points used in the results validation, 1 point (sample 21) was not considered as a conversion from our estimations, however, the visual inspection pointed to a likely event of conversion from forest to agriculture.
+Also, there were 6 points (sample 6, 7, 55, 56, 78, 97) which the visual inspection did not find a conversion from forest to pasture, although our estimations pointed as conversions.
 Therefore, 7\% of the sample were completely misclassified by our estimates, and the rest of the accuracy assessment was performed over the remaining 93 sample points.
 
-When analyzing the errors from the transition length estimates, we observed that the year of deforestation shows the least amount of errors (Figure \ref{fig:errorbar-plot}).
+When analyzing the errors from the conversion length estimates, we observed that the year of deforestation shows the least amount of errors (Figure \ref{fig:errorbar-plot}).
 The MAE of the deforestation year is 1.42 years, and the error shows a bias towards underestimation.
-The year of the agriculture establishment and the transition length estimates showed larger errors when compared with visual inspection.
+The year of the agriculture establishment and the conversion length estimates showed larger errors when compared with visual inspection.
 
@@
 Transition length values are more concentrated in smaller values, which also shows higher errors (Figure \ref{fig:errorscatter-plot}.c).
-However, this is expected since faster transitions are more common (Figure \ref{fig:map-plot}).
+However, this is expected since faster conversions are more common (Figure \ref{fig:map-plot}).
 

### 2022-12-05 scripts/figures/fig_1.R
Change fonts and more minor changes.

@@ -158,7 +158,7 @@ dist_map <- ggplot() +
     color = "#c50505"
   ) +
   ggtitle(
-    "Distribution of transitions from forest to agriculture in the Amazon"
+    "Distribution of conversion from forest to agriculture in the Amazon"
   ) +
   guides(
     fill = guide_colourbar(
@@ -171,9 +171,9 @@ dist_map <- ggplot() +
   scale_fill_scico(
     palette = "batlow",
     breaks = c(0, 17, 35),
-    name = "Transition Length (year)"
+    name = "Conversion Length (year)"
   ) +
-  theme_void() +
+  theme_void(base_family = "roboto") +
   theme(
     legend.position = c(0.14, 0.77),
     text = element_text(size = 11)
@@ -188,7 +188,7 @@ hr_map <- ggplot() +
   scale_x_continuous(expand = c(0.0004, 0.0004)) +
   scale_y_continuous(expand = c(0.0004, 0.0004)) +
   coord_fixed() +
-  theme_void() +
+  theme_void(base_family = "roboto") +
   theme(
     legend.position = "",
     panel.background = element_rect(fill = "#747171"),
@@ -205,7 +205,7 @@ hist_plot <- ggplot() +
   scale_fill_scico_d(palette = "batlow") +
   scale_x_continuous(expand = c(0, 0)) +
   scale_y_continuous(expand = c(0, 0)) +
-  theme_void() +
+  theme_void(base_family = "roboto") +
   theme(
     legend.position = ""
   )
@@ -214,7 +214,7 @@ full_map <- dist_map + theme(plot.tag.position = c(0.2, 0.8)) +
   inset_element(hr_map, 0.76, 0, 1, 0.34, align_to = "panel") +
   theme(plot.tag.position = c(0.95, 1.08)) +
   inset_element(hist_plot, 0, 0, 0.5, 0.4, align_to = "panel") +
-  theme(plot.tag.position = c(0.4, 0.4)) +
+  theme(plot.tag.position = c(0.4, 0.37)) +
   plot_annotation(tag_levels = 'a')
 
 ggsave(
@@ -225,4 +225,3 @@ ggsave(
   units = "cm",
   dpi = 300
 )
-

### 2022-12-05 manuscript/manuscript.Rmd
Minor changes in the whole text. Change order of methods.

@@
 The pixels of the Amazon biome were filtered to contain only those that were occupied by forest and agriculture at some period, and that are not considered as water, according to the Global Surface Water product provided by Copernicus [@Pekel2016].
-For the download, the data was divided in 2732 tiles, to allow local processing of a large quantity of pixels.Two sets of data were downloaded with the same spatial dimensions, one with the LULC classifications for each year, and another with binary values, representing whether that pixels was valid for processing or not.
+For the download, the data was divided in 2732 tiles, to allow local processing of a large quantity of pixels.Two sets of data were downloaded with the same spatial dimensions, one with the LULC classifications for each year, and another with binary values, representing whether that pixels were valid for processing or not (according to the filter conditions mentioned above).
 
@@
 
-The second part of the processing was to load each tile separately, convert it to a table and calculate every possible conversion from forest to agriculture.
-The conversions were calculated pixel by pixel, by performing the following steps:
+The second part of the processing was to load each tile separately, convert it to a table and calculate every possible conversion from forest to agriculture (Figure \@ref(fig:routine-plot)).
+The conversions were calculated pixel by pixel, where the LULC time series were extracted and analysed for each conversion cycle that was present for that pixel.
+The calculation of the conversion length was extremely simple, it is the year of the establishment of agriculture subtracted by the last year classified as a forest, however, we add one year to the latter, so that it represents the year of deforestation (since it is more meaningful for analysis).
+
+```{r routine-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = "Representation of the process to calculate the conversion length from forests to agriculture.
+All values presented are hiphothetical, and serves only as an ilustration of the method."}
+
+knitr::include_graphics("figs/change_lenght_routine_scheme.png")
+```
+
+The process of calculation of the conversion length (Figure \@ref(fig:routine-plot)), can be breakdown to seven steps of processing:
 
@@
 
-2.  Calculate the year of first occurrence of any agriculture class for each pixel;
+2.  Calculate the year of first occurrence of any agriculture class (annual and permanent crops) for each pixel;
 
@@
 4.  Classify rows as "before" or "after" the occurrence of the year calculated in step 3.
-    This is the identification of periods before and after the recurrence of forest after a first conversion to agriculture;
+    This is the identification of periods before and after a return of forest after a first conversion to agriculture;
 
@@
 The steps 2 to 7 are performed recursively to identify multiple conversions, in case they are present.
-In addition to the conversion length values and the years of the conversion, they are also qualified by the type of agriculture that were established after deforestation, if the conversion occurred from primary or secondary forests, and the number of recurrences of conversions at the pixel.
+In addition to the conversion length values and the years of the conversion, they are also qualified by the type of agriculture that were established after deforestation, if the conversion occurred from primary or secondary forests, and the number of occurences of conversions at the pixel.
 The LULC classes within the conversion and until 5 years after were also stored.
@@
 After the calculations, data was also converted back to raster format, with the aim to expand the accessibility of the data set.
+
+## Description of data collection
+
+After the calculation of conversions, the results are stored in three different types of tables (Figure \@ref(fig:tables-plot)), organized in a folder structure and stored as Apache Parquet files.
+
+```{r tables-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = " Representation of the tables generated by the conversion length calculation.
+All values presented are hiphothetical, and serves only as an ilustration of the method."}
+
+knitr::include_graphics("figs/tables_scheme.png")
+```
+
+The first type of table (Figure \@ref(fig:tables-plot).a) contain the spatial information (longitude, latitude and area) of each pixel, its unique identificators and the code of the municipality which contains the pixel.
+The "cell_id" column contains a unique identification for the whole data, while the "tile_cell_id" holds a unique value of each pixel for a single tile.
+Each tile is also uniquely identified in the column "tile_id".
+This table is named as "mask_cells".
+
+A second type of tables (Figure \@ref(fig:tables-plot).b) stores conversion length values, the first and last year of the conversion, the resulting agriculture type, and the number of the conversion cycle.
+They also contain the unique id of each pixel and tile (related to the table above).
+The "agri_code" column holds information about the type of agriculture the was established after deforestation.
+The "forest_year" have values that represent the year of deforestation, the "agri_year" is the year when the agriculture establishment happened, and the subtraction between both (the conversion length) is stored in the "c_length" column.
+Since it is possible that one pixels contains more than one conversion from forest to agriculture, we stored the number of the conversion in the column "c_cycle" (1 is the first conversion cycle).
+However, conversion from forest to other LULC is also possible, so the column "f_cycle" holds information about the number of the conversion from forest to any LULC.
+If the forest have suffered any conversion, it starts being considered as a secondary forest, as indicated in the column "forest_type".
+
+The third type of tables (Figure \@ref(fig:tables-plot).c) contains the LULC classes of all years within the conversion, and also the first 5 years after the conversion.
+The column "year" holds the year of the MapBiomas classification, and the "class_code" variable stores the LULC classification (according to MapBiomas) for the respective year.
+
+The three tables are related to each other and can be used altogether, and are separated by tiles.
+Another table containing the metadata of each tile is also created, and holds the spatial characteristics of the tiles.
+With this spatial information, it is possible to convert the tabular data back to spatial raster, with identical spatial properties as the original MapBiomas classification data.
+
+The tabular data was organized in a system of folders and sub folders in a hive format partitioning (Figure \@ref(fig:dataset-plot)).
+This structure allows for easily access parts of the data set, and also filter it without the necessity to read the tables.
+In the case of this data set, it is partitioned by the "tile_id" and "agri_cycle" variables.
+The data set also have a metadata file with the spatial characteristics of each file, it stores the identification of the tile, the maximum and minimum latitude and longitude, the coordinate reference system, and the number of columns, rows, and cells of each tile.
+
+```{r dataset-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = " Representation of the data set, structured in folders and sub folders."}
+
+knitr::include_graphics("figs/dataset_structure.png")
+```
 
@@
 After the completion of the analysis of the 100 sample points, we also conducted a qualitative assessment, where we compared our results with the satellite images.
-
-# Description of data collection
-
-After the calculation of conversions, the results are stored in three different types of tables, organized in a folder structure and stored as Apache Parquet files.
-
--   Tables that contain the spatial information (longitude, latitude) of each pixel, its unique id and the code of the municipality which contains the pixel.
-    It is named as "mask_cells";
-
--   Tables with conversion length values, the first and last year of the conversion, the resulting agriculture type, and the number of the conversion cycle.
-    They also contain the unique id of each pixel (related to the table above);
-
--   Table with the LULC classes of all years within the conversion, and also the first 5 years after the conversion.
-
-The three tables are related to each other and can be used altogether, and are separated by tiles.
-Another table containing the metadata of each tile is also created, and holds the spatial characteristics of the tiles.
-With this spatial information, it is possible to convert the tabular data back to spatial raster, with identical spatial properties as the original MapBiomas classification data.
 
@@
 
-The period between 2001 and 2005 showed a peak of deforestation destined to agriculture (Figure \@ref(fig:transbar-plot).a), and of croplands expansion (Figure \@ref(fig:transbar-plot).b).
+The period between 2001 and 2005 showed a peak of deforestation (Figure \@ref(fig:transbar-plot).a), and of croplands expansion (Figure \@ref(fig:transbar-plot).b).
 From 2006 to 2010, we observed a decrease in both variables, a process that was already observed between 2001 and 2010 [@Macedo2012], which points out a decoupling of soybean expansion and deforestation.
@@
 In the late 1990s, the development policies "Brasil em Ação" and "Avança Brasil" (1995 - 2003) accelerated the national infrastructure expansion, including in the Amazon [@Carvalho2002].
-This is a period when the deforestation areas to be converted to agriculture reached the highest values in the time series (Figure \@ref(fig:transbar-plot).a), and also a peak in the establishment of agriculture in the Amazon biome, predominantly with fast conversions (Figure \@ref(fig:transbar-plot).b).
+This is a period when the deforestation areas reached the highest values in the time series (Figure \@ref(fig:transbar-plot).a), and also a peak in the establishment of agriculture in the Amazon biome, predominantly with fast conversions (Figure \@ref(fig:transbar-plot).b).
 
@@
 In 2006, the Brazilian Association of Vegetable Oil Industries (BIOVE) and the National Association of Cereal Exporters (ANEC) committed to avoid commercialization of soybean grains harvested from areas deforested after 2008.
-Our estimates of conversions show a decrease of deforested areas to be converted to agriculture after 2008, where it reached minimal values (Figure \@ref(fig:transbar-plot).a).
+Our estimates of conversions show a decrease of deforested areas after 2008, where it reached minimal values (Figure \@ref(fig:transbar-plot).a).
 After 2006, agriculture establishment over deforested primary forests suffered a decrease, which stayed relatively stable until 2012, where a steep increase occurred, however, the new areas being occupied by agriculture were mainly over areas that were cleared more than a decade before (therefore, before 2006) (Figure \@ref(fig:transbar-plot).b) This shows that agriculture expansion did not halt after the soy moratorium, but producers expended over areas cleared many years ago, which may have indirect impacts on deforestation increase over distant areas [@Arima2011].
@@
 ```{r errorscatter-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = "Scatter plot between paired values of estimated and observed Deforestation and Agriculture Establishment years, and their respective Transition Length.
+The gray line represents values which there would be no errors of the estimated values in comparison with the observed values.
 Translucid white areas represent quantile ranges (5\\%, 25\\%, 75\\% and 95\\%) of the errors.
@@
 The Agriculture class presented more errors, the inclusion errors ranged from 22% to 5%, and were mostly composed by Forests pixels (forest pixels misclassified as agriculture).
-The omission errors of Agriculture ranged from 22% to 8%, and were also mostly composed of Forest pixels (agriculture pixels misclassified as Agriculture).
+The omission errors of Agriculture ranged from 22% to 8%, and were also mostly composed of Forest pixels (agriculture pixels misclassified as forest).
 
@@
 \conclusions[Conclusions]
-
-TODO: resumo ainda precisa de mais informações, por exemplo, a série temporal da análise, a escala (ex. todo o bioma Amazônia no Brasil, na América do Sul, apenas alguns estados da Amazônia).
-Falar do método, que tipo de álgebra de mapa foi usada, algum tipo de regressão...enfim.

### 2022-12-05 manuscript/manuscript.tex
Minor changes in the whole text. Change order of methods.

@@
 The pixels of the Amazon biome were filtered to contain only those that were occupied by forest and agriculture at some period, and that are not considered as water, according to the Global Surface Water product provided by Copernicus \citep{Pekel2016}.
-For the download, the data was divided in 2732 tiles, to allow local processing of a large quantity of pixels.Two sets of data were downloaded with the same spatial dimensions, one with the LULC classifications for each year, and another with binary values, representing whether that pixels was valid for processing or not.
+For the download, the data was divided in 2732 tiles, to allow local processing of a large quantity of pixels.Two sets of data were downloaded with the same spatial dimensions, one with the LULC classifications for each year, and another with binary values, representing whether that pixels were valid for processing or not (according to the filter conditions mentioned above).
 
@@
 
-The second part of the processing was to load each tile separately, convert it to a table and calculate every possible conversion from forest to agriculture.
-The conversions were calculated pixel by pixel, by performing the following steps:
+The second part of the processing was to load each tile separately, convert it to a table and calculate every possible conversion from forest to agriculture (Figure \ref{fig:routine-plot}).
+The conversions were calculated pixel by pixel, where the LULC time series were extracted and analysed for each conversion cycle that was present for that pixel.
+The calculation of the conversion length was extremely simple, it is the year of the establishment of agriculture subtracted by the last year classified as a forest, however, we add one year to the latter, so that it represents the year of deforestation (since it is more meaningful for analysis).
+
+\begin{figure}[ht]
+\includegraphics[width=17cm]{figs/change_lenght_routine_scheme} \caption{Representation of the process to calculate the conversion length from forests to agriculture.
+All values presented are hiphothetical, and serves only as an ilustration of the method.}\label{fig:routine-plot}
+\end{figure}
+
+The process of calculation of the conversion length (Figure \ref{fig:routine-plot}), can be breakdown to seven steps of processing:
 
@@
 \item
-  Calculate the year of first occurrence of any agriculture class for each pixel;
+  Calculate the year of first occurrence of any agriculture class (annual and permanent crops) for each pixel;
 \item
@@
 
+\subsection{Description of data collection}
+
+After the calculation of conversions, the results are stored in three different types of tables (Figure \ref{fig:tables-plot}), organized in a folder structure and stored as Apache Parquet files.
+
+\begin{figure}[ht]
+\includegraphics[width=17cm]{figs/tables_scheme} \caption{ Representation of the tables generated by the conversion length calculation.
+All values presented are hiphothetical, and serves only as an ilustration of the method.}\label{fig:tables-plot}
+\end{figure}
+
+The first type of table (Figure \ref{fig:tables-plot}.a) contain the spatial information (longitude, latitude and area) of each pixel, its unique identificators and the code of the municipality which contains the pixel.
+The ``cell\_id'' column contains a unique identification for the whole data, while the ``tile\_cell\_id'' holds a unique value of each pixel for a single tile.
+Each tile is also uniquely identified in the column ``tile\_id''.
+This table is named as ``mask\_cells''.
+
+A second type of tables (Figure \ref{fig:tables-plot}.b) stores conversion length values, the first and last year of the conversion, the resulting agriculture type, and the number of the conversion cycle.
+They also contain the unique id of each pixel and tile (related to the table above).
+The ``agri\_code'' column holds information about the type of agriculture the was established after deforestation.
+The ``forest\_year'' have values that represent the year of deforestation, the ``agri\_year'' is the year when the agriculture establishment happened, and the subtraction between both (the conversion length) is stored in the ``c\_length'' column.
+Since it is possible that one pixels contains more than one conversion from forest to agriculture, we stored the number of the conversion in the column ``c\_cycle'' (1 is the first conversion cycle).
+However, conversion from forest to other LULC is also possible, so the column ``f\_cycle'' holds information about the number of the conversion from forest to any LULC.
+If the forest have suffered any conversion, it starts being considered as a secondary forest, as indicated in the column ``forest\_type''.
+
+The third type of tables (Figure \ref{fig:tables-plot}.c) contains the LULC classes of all years within the conversion, and also the first 5 years after the conversion.
+The column ``year'' holds the year of the MapBiomas classification, and the ``class\_code'' variable stores the LULC classification (according to MapBiomas) for the respective year.
+
+The three tables are related to each other and can be used altogether, and are separated by tiles.
+Another table containing the metadata of each tile is also created, and holds the spatial characteristics of the tiles.
+With this spatial information, it is possible to convert the tabular data back to spatial raster, with identical spatial properties as the original MapBiomas classification data.
+
+The tabular data was organized in a system of folders and sub folders in a hive format partitioning (Figure \ref{fig:dataset-plot}).
+This structure allows for easily access parts of the data set, and also filter it without the necessity to read the tables.
+In the case of this data set, it is partitioned by the ``tile\_id'' and ``agri\_cycle'' variables.
+The data set also have a metadata file with the spatial characteristics of each file, it stores the identification of the tile, the maximum and minimum latitude and longitude, the coordinate reference system, and the number of columns, rows, and cells of each tile.
+
+\begin{figure}[ht]
+\includegraphics[width=17cm]{figs/dataset_structure} \caption{ Representation of the data set, structured in folders and sub folders.}\label{fig:dataset-plot}
+\end{figure}
+
 \subsection{Accuracy assessment}
@@
 After the completion of the analysis of the 100 sample points, we also conducted a qualitative assessment, where we compared our results with the satellite images.
-
-\section{Description of data collection}
-
-After the calculation of conversions, the results are stored in three different types of tables, organized in a folder structure and stored as Apache Parquet files.
-
-\begin{itemize}
-\item
-  Tables that contain the spatial information (longitude, latitude) of each pixel, its unique id and the code of the municipality which contains the pixel.
-  It is named as ``mask\_cells'';
-\item
-  Tables with conversion length values, the first and last year of the conversion, the resulting agriculture type, and the number of the conversion cycle.
-  They also contain the unique id of each pixel (related to the table above);
-\item
-  Table with the LULC classes of all years within the conversion, and also the first 5 years after the conversion.
-\end{itemize}
-
-The three tables are related to each other and can be used altogether, and are separated by tiles.
-Another table containing the metadata of each tile is also created, and holds the spatial characteristics of the tiles.
-With this spatial information, it is possible to convert the tabular data back to spatial raster, with identical spatial properties as the original MapBiomas classification data.
 
@@
 \includegraphics[width=17cm]{figs/error_scatter} \caption{Scatter plot between paired values of estimated and observed Deforestation and Agriculture Establishment years, and their respective Transition Length.
+The gray line represents values which there would be no errors of the estimated values in comparison with the observed values.
 Translucid white areas represent quantile ranges (5\%, 25\%, 75\% and 95\%) of the errors.
@@
 \conclusions[Conclusions]
-
-TODO: resumo ainda precisa de mais informações, por exemplo, a série temporal da análise, a escala (ex. todo o bioma Amazônia no Brasil, na América do Sul, apenas alguns estados da Amazônia).
-Falar do método, que tipo de álgebra de mapa foi usada, algum tipo de regressão\ldots enfim.
 

### 2022-12-05 manuscript/manuscript.Rmd
Many insertions to the text. Inclusion of conclusion. Improvement of results. Fix typos. Update figures.

@@
 abstract: |
-  Land-use and land-cover changes in the Amazon biome are a main process that influences the environment and societies at local, national, and global levels.
+  Land-use and land-cover changes in the Amazon biome are main processes that influence the environment and societies at local, national, and global levels.
 A large quantity of studies already relied on land-use classification to analyze this process.
@@
 Land-use and land-cover (LULC) changes are known to cause impacts on different scales.
-Deforestation and land use changes can affect hydrological processes in large basins [@Arias2018]; reduce the forest resilience to extreme events (e.g., drought) and other sources of perturbation [@Boulton2022]; deforestation and farming practices can reduce convective rainfall and increase surface temperature [@Maeda2021]; increase carbon emissions [@Gatti2021]; cause important impacts on fauna and flora diversity, soil properties and carbon stocks [@Nunes2022; @Rittl2017]; and even negatively impact public health [@Ellwanger2020].
+Deforestation and LULC changes can affect hydrological processes in large basins [@Arias2018]; reduce the forest resilience to extreme events (e.g., drought) and other sources of perturbation [@Boulton2022]; deforestation and farming practices can reduce convective rainfall and increase surface temperature [@Maeda2021]; increase carbon emissions [@Gatti2021]; cause important impacts on fauna and flora diversity, soil properties and carbon stocks [@Nunes2022; @Rittl2017]; and even negatively impact public health [@Ellwanger2020].
 
@@
 Such types of data are essential to analyze LULC change processes, make causal inferences, and understand what drives and what are the impacts of the transformation of the landscape in Brazil.
-In this context, this study aims to characterize and quantify the length of the conversions from forest formations to agriculture in the Amazon over the last 36 years.
-The availability of this data can be useful to the development of interdisciplinary research involved with LULC changes.
+In this context, this study aims to quantify and characterize the length of the conversions from forest formations to agriculture in the Amazon over the last 36 years.
+The availability of this data can be useful to the development of interdisciplinary research, since it can provide useful insights into the environmental, economic, and social impacts of land conversion.
+For example, this data can be used to develop insights into the economic costs and benefits associated with land conversion, the effects of land conversion on local ecosystems, and the impact of land conversion on livelihoods and social dynamics.
 
@@
 
-The first part of the processing was to create a table from the binary values data, and create unique identification for each pixel, their coordinates, calculate the area, and also address the municipality in which each pixel was inside.
+The first part of the processing was to create a table from the binary values data, and create unique identification for each pixel, their coordinates, calculate the area, and also the municipality in which each pixel was inside.
 In this part, a table was also generated with the spatial information of each tile, so that it is possible to convert the generated tables back to raster format.
@@
 The second part of the processing was to load each tile separately, convert it to a table and calculate every possible conversion from forest to agriculture (Figure \@ref(fig:routine-plot)).
-The conversions were calculated pixel by pixel, where the LULC time series were extracted and analysed for each conversion cycle that was present for that pixel.
-The calculation of the conversion length was extremely simple, it is the year of the establishment of agriculture subtracted by the last year classified as a forest, however, we add one year to the latter, so that it represents the year of deforestation (since it is more meaningful for analysis).
+The conversions were calculated pixel by pixel, where the LULC time series were extracted and analyzed for each conversion cycle that was present for that pixel.
+The calculation of the conversion length was simple, it is the year of the establishment of agriculture subtracted by the last year classified as a forest, however, we add one year to the latter, so that it represents the year of deforestation (since it is more meaningful for analysis).
 
-```{r routine-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = "Representation of the process to calculate the conversion length from forests to agriculture.
-All values presented are hiphothetical, and serves only as an ilustration of the method."}
+```{r routine-plot, out.width = "17cm", fig.pos = "h", echo = FALSE, fig.cap = "Representation of the process to calculate the conversion length from forests to agriculture.
+All values presented are hyphothetical, and serve only as an illustration of the method."}
 
@@
 
-The process of calculation of the conversion length (Figure \@ref(fig:routine-plot)), can be breakdown to seven steps of processing:
+The process of calculation of the conversion length (Figure \@ref(fig:routine-plot)), can be divided into seven steps of processing:
 
@@
 The steps 2 to 7 are performed recursively to identify multiple conversions, in case they are present.
-In addition to the conversion length values and the years of the conversion, they are also qualified by the type of agriculture that were established after deforestation, if the conversion occurred from primary or secondary forests, and the number of occurences of conversions at the pixel.
+In addition to the conversion length values and the years of the conversion, they are also qualified by the type of agriculture that were established after deforestation, if the conversion occurred from primary or secondary forests, and the number of occurrences of conversions at the pixel.
 The LULC classes within the conversion and until 5 years after were also stored.
@@
 
-```{r tables-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = " Representation of the tables generated by the conversion length calculation.
-All values presented are hiphothetical, and serves only as an ilustration of the method."}
+```{r tables-plot, out.width = "17cm", fig.pos = "h", echo = FALSE, fig.cap = " Representation of the tables generated by the conversion length calculation.
+All values presented are hyphothetical, and serve only as an illustration of the method."}
 
@@
 
-The first type of table (Figure \@ref(fig:tables-plot).a) contain the spatial information (longitude, latitude and area) of each pixel, its unique identificators and the code of the municipality which contains the pixel.
+The first type of table (Figure \@ref(fig:tables-plot).a) contains the spatial information (longitude, latitude and area) of each pixel, its unique identification and the code of the municipality which contains the pixel.
 The "cell_id" column contains a unique identification for the whole data, while the "tile_cell_id" holds a unique value of each pixel for a single tile.
 Each tile is also uniquely identified in the column "tile_id".
-This table is named as "mask_cells".
+This table is named "mask_cells".
 
@@
 The "forest_year" have values that represent the year of deforestation, the "agri_year" is the year when the agriculture establishment happened, and the subtraction between both (the conversion length) is stored in the "c_length" column.
-Since it is possible that one pixels contains more than one conversion from forest to agriculture, we stored the number of the conversion in the column "c_cycle" (1 is the first conversion cycle).
+Since it is possible that one pixel contains more than one conversion from forest to agriculture, we stored the number of the conversion in the column "c_cycle" (1 is the first conversion cycle).
 However, conversion from forest to other LULC is also possible, so the column "f_cycle" holds information about the number of the conversion from forest to any LULC.
-If the forest have suffered any conversion, it starts being considered as a secondary forest, as indicated in the column "forest_type".
+If the forest has suffered any conversion, it starts being considered as a secondary forest, as indicated in the column "forest_type".
 
@@
 The tabular data was organized in a system of folders and sub folders in a hive format partitioning (Figure \@ref(fig:dataset-plot)).
-This structure allows for easily access parts of the data set, and also filter it without the necessity to read the tables.
+This structure allows for easily accessing parts of the data set, and also to filter it without the necessity to read the tables.
 In the case of this data set, it is partitioned by the "tile_id" and "agri_cycle" variables.
-The data set also have a metadata file with the spatial characteristics of each file, it stores the identification of the tile, the maximum and minimum latitude and longitude, the coordinate reference system, and the number of columns, rows, and cells of each tile.
+The data set also has a metadata file with the spatial characteristics of each file, it stores the identification of the tile, the maximum and minimum latitude and longitude, the coordinate reference system, and the number of columns, rows, and cells of each tile.
 
-```{r dataset-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = " Representation of the data set, structured in folders and sub folders."}
+```{r dataset-plot, out.width = "17cm", fig.pos = "h", echo = FALSE, fig.cap = " Representation of the data set, structured in folders and sub folders."}
 
@@
 We also analyzed the results by plotting the errors as frequency bars, and scatter plots between observed and estimated values.
-After the completion of the analysis of the 100 sample points, we also conducted a qualitative assessment, where we compared our results with the satellite images.
+After the completion of the analysis of the 100 sample points, we also conducted a qualitative assessment, where we compared our results with satellite images composites.
 
@@
 
-The conversion calculations show that between 1985 and 2021, 64,874 square kilometers of forests were converted to agriculture in the Brazilian Amazon biome.
+The conversion calculations show that in the Brazilian Amazon biome, 64874 square kilometers of forests were converted to agriculture between 1985 and 2021.
 The length of the conversions can go from 0 to 35 years, in which conversions closer to 0 years are considered as fast conversions, and conversions closer to 35 years are considered as slow conversions.
@@
 
-```{r map-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = "Map of distribution of conversions from forest to agriculture in the Brazilian Amazon biome.
+```{r map-plot, out.width = "17cm", fig.pos = "h", echo = FALSE, fig.cap = "Map of distribution of conversions from forest to agriculture in the Brazilian Amazon biome.
 The hexagonal cells represent the most common conversion length, and do not reflect the amount of area of conversions inside a cell.
 Transitions are concentrated in the south (Mato Grosso state), and in the east (Pará and Maranhão states).
-The conversion length ranges from 0 (dark blue) to 35 years (light pink), and clusters of fast conversions (conversions closer to 1 year) can be discerned from clusters of slow conversions (conversions closer to 36 years).
+The conversion length ranges from 0 (dark blue) to 35 years (light pink), and clusters of fast conversions (conversions closer to 0 year) can be discerned from clusters of slow conversions (conversions closer to 35 years).
 The histogram located in the bottom left shows that fast conversions are more common than slower conversions.
@@
 When observing at the original scale, we could not spot any well defined pattern or direction of the occurrence of faster to slower conversions (Figure \@ref(fig:map-plot)).
+Other studies investigating patterns of conversions in the Amazon also found the formation of clusters of patterns, although there is a big heterogeneity at larger scales [@MullerHansen2017].
 
-Other studies investigating patterns of conversions in the Amazon also found the formation of clusters of patterns, although there is a big heterogeneity at larger scales [@MullerHansen2017].
 The data can be analyzed year by year, and also be separated by primary and secondary forests being converted to agriculture (Figure \@ref(fig:transbar-plot)).
 
-```{r transbar-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = "Transition area from forest to agriculture, per year and conversion length.
+```{r transbar-plot, out.width = "17cm", fig.pos = "h", echo = FALSE, fig.cap = "Transition area from forest to agriculture, per year and conversion length.
 The bars represent the total amount of area at some state of the conversion for each year.
-The color gradient in each bar represents the conversion length related to a deforestation or an agriculture establishment event.
+The color gradient in each bar represent the conversion length related to a deforestation or an agriculture establishment event.
 Blue tones represents fast conversions (closer to 0 years), pink tones represent slow conversions (closer to 35 years).
@@
 The deforestation area of primary forests increased largely from 1986 to 2003, which was followed by a steep decrease until 2009, when the deforested areas reached a stable rate (Figure \@ref(fig:transbar-plot).a).
-From 1986 to approximately 1995, most of the deforested areas suffered a slow conversion, mostly were higher than 10 years.
-After this period, fast conversions started to become more common, specially from 2002 to 2004 (Figure \@ref(fig:transbar-plot).a).
-The peak of deforestation in 2003, according to our results, shows that 12.6% of the conversions were a direct conversion from forest to agriculture.
-Previous studies estimated a proportion of 23% [@Morton2006].
-If we include fast conversions (from 0 to 2 years) from our results, the proportion jumps to 48.5 % of all conversions.
+From 1985 to 1995, most deforested areas suffered a slow conversion, mostly were higher than 10 years (76 %).
+After this period, fast conversions started to become more common (Figure \@ref(fig:transbar-plot).a), specially from 2002 to 2004, where 70 % of the transitions were faster than 10 years.
+Our results show that in 2003 (the peak of deforestation), 12.7% of the conversions were a direct conversion from forest to agriculture.
+Previous studies estimated a proportion of 23 % in the Mato Grosso state, for the same year [@Morton2006].
+If we include fast conversions (from 0 to 2 years) from our results, the proportion jumps to 48.1 % of all conversions in 2003.
 
 Agricultural establishment over areas of primary forests peaked in 2005 and 2013 (Figure \@ref(fig:transbar-plot).b).
-Despite similar agricultural establishment rates between both years, their conversion lengths differ greatly, in 2005 most of the conversions were faster than 10 years, while in 2013 the great majority of conversions were slower than 10 years (Figure \@ref(fig:transbar-plot).b).
-In 2003 it was observed a shift in the conversion length from forest until the establishment of agricultural areas.
-From this year onward, most conversions happened in areas deforested at least 10 years before (Figure \@ref(fig:transbar-plot).b).
+Despite similar agricultural establishment areas in both years, their conversion lengths differ greatly (Figure \@ref(fig:transbar-plot).b), in 2005 most of the conversions were faster than 10 years (78 %), while in 2013 the majority of conversions were slower than 10 years (66 %).
+In 2010, it was observed a shift in the conversion length from forest until the establishment of agricultural areas.
+From this year onward, 68 % of the conversions happened in areas deforested at least 10 years before (Figure \@ref(fig:transbar-plot).b).
 Even after the decrease of deforestation after 2002, areas under agriculture expanded over lands where deforestation happened before 2002 (mostly over pastureland).
 However, after 2019, a sudden drop of agriculture establishment rate happened (Figure \@ref(fig:transbar-plot).b).
-
-The period between 2001 and 2005 showed a peak of deforestation (Figure \@ref(fig:transbar-plot).a), and of croplands expansion (Figure \@ref(fig:transbar-plot).b).
-From 2006 to 2010, we observed a decrease in both variables, a process that was already observed between 2001 and 2010 [@Macedo2012], which points out a decoupling of soybean expansion and deforestation.
-However, our result shows that after 2011, agricultural areas rapidly expanded in the amazon biome, especially at areas deforested many years before.
+This behavior could be related to economic causes or even the scarcity of suitable areas for agricultural expansion, which seems unlikely, since the last years have been marked by an increase of deforestation rates and the advance of pastureland over the forest.
 
@@
 
-In 2006, the Brazilian Association of Vegetable Oil Industries (BIOVE) and the National Association of Cereal Exporters (ANEC) committed to avoid commercialization of soybean grains harvested from areas deforested after 2008.
+In 2006, the Brazilian Association of Vegetable Oil Industries (BIOVE) and the National Association of Cereal Exporters (ANEC) committed to avoid commercialization of soybean grains harvested from areas deforested after 2008, known as the soy moratorium.
+The soy moratorium is considered to have contributed to the decline of the deforestation directly related to soybean expansion [@Paim2021; @Amaral2021; @Heilmayr2020; @Kastens2017].
 Our estimates of conversions show a decrease of deforested areas after 2008, where it reached minimal values (Figure \@ref(fig:transbar-plot).a).
-After 2006, agriculture establishment over deforested primary forests suffered a decrease, which stayed relatively stable until 2012, where a steep increase occurred, however, the new areas being occupied by agriculture were mainly over areas that were cleared more than a decade before (therefore, before 2006) (Figure \@ref(fig:transbar-plot).b) This shows that agriculture expansion did not halt after the soy moratorium, but producers expended over areas cleared many years ago, which may have indirect impacts on deforestation increase over distant areas [@Arima2011].
-However, this indirect effect is not always clearly observed when analyzing the Amazon biome as a whole, but our data shows that periods of fast agriculture expansion occurred at the same time with minimal deforestation values, according to PRODES estimates (add reference).
+After 2006, agriculture establishment over deforested primary forests suffered a decrease, which stayed relatively stable until 2012, where a steep increase occurred, however, the new areas being occupied by agriculture were mainly over areas that were cleared before 2008 (Figure \@ref(fig:transbar-plot).b).
+The period between 2006 and 2010 is also considered as a turnover point, known as the decoupling of deforestation and soybean production, where the total production was not as dependent on agriculture expansion as before [@Macedo2012].
+Therefore, the observed increase of agricultural establishment after 2012 shows that agriculture expansion did not halt after the soy moratorium nor after periods of high soybean productivity.
+Producers started to expand over areas cleared many years ago, which may still have indirect impacts on deforestation increase over distant areas [@Arima2011; @Gollnow2018].
+However, this indirect effect is not always clearly observed when analyzing the Amazon biome as a whole, our data shows that periods of fast agriculture expansion occurred at the same time with minimal deforestation values, according to PRODES estimates [@Assis2019].
 Expansion of agricultural areas also occurred over cleared areas of secondary forests in 2012, but with an important amount of fast conversions (Figure \@ref(fig:transbar-plot).d).
-The causes of the increase of agricultural establishment areas can be numerous, and one main driver was the approval of a new Forest Code, in 2012, which is considered to have undermined the environmental protection of forests [@Kroger2017; @Pereira2019].
+The causes of the increase of agricultural establishment areas can be numerous, and one main driver was the approval of a new Forest Code, in 2012, which is considered to have undermined the environmental protection of forests [@Kroger2017; @Pereira2019], since it opened the possibility of amnesty for illegally deforested areas [@Schielein2018; @Santanna2021; @Filho2014], and the possible reduction of legal reserve area [@Freitas2018].
 
-The conversion length patterns across years can change significantly between different states (Figure \@ref(fig:transridge-plot)).
+Even though it is possible to analyze the results in the Amazon as a whole, the conversion length patterns across years can change significantly between different states (Figure \@ref(fig:transridge-plot)).
 
-```{r transridge-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = "Transition length patterns inside states in the Amazon biome.
+```{r transridge-plot, out.width = "17cm", fig.pos = "h", echo = FALSE, fig.cap = "Transition length patterns inside states in the Amazon biome.
 Each year has a density estimate of the conversion lengths, represented as colored curves.
 The peak of the curves represents conversion length values with more frequency in one year of one state.
-Blue tones represents fast conversions (conversions closer to 0 years), pink tones represents slow conversions (conversions closer to 35 years)"}
+Blue tones represent fast conversions (conversions closer to 0 years), pink tones represent slow conversions (conversions closer to 35 years)."}
 
@@
 
-```{r errorbar-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = "Bar with the count of error values (difference between observed and estimated values).
+```{r errorbar-plot, out.width = "17cm", fig.pos = "h", echo = FALSE, fig.cap = "Bar with the count of error values (difference between observed and estimated values).
 Positive values indicate underestimation of the variable (estimations were lower than observations), negative values indicate overestimation (estimations higher than observations).
@@
 
-```{r errorscatter-plot, out.width = "17cm", fig.pos = "ht", echo = FALSE, fig.cap = "Scatter plot between paired values of estimated and observed Deforestation and Agriculture Establishment years, and their respective Transition Length.
-The gray line represents values which there would be no errors of the estimated values in comparison with the observed values.
+```{r errorscatter-plot, out.width = "17cm", fig.pos = "h", echo = FALSE, fig.cap = "Scatter plot between paired values of estimated and observed Deforestation and Agriculture Establishment years, and their respective Transition Length.
+The gray line represents values where there would be no errors of the estimated values in comparison with the observed values.
 Translucid white areas represent quantile ranges (5\\%, 25\\%, 75\\% and 95\\%) of the errors.
@@
 The omission errors for Forests are also small (proportion of pixels misclassified as Forests, but the real class were not Forest), which fluctuated around 2%.
-The Agriculture class presented more errors, the inclusion errors ranged from 22% to 5%, and were mostly composed by Forests pixels (forest pixels misclassified as agriculture).
+The Agriculture class presented more errors, the inclusion errors ranged from 22% to 5%, and were mostly composed of Forests pixels (forest pixels misclassified as agriculture).
 The omission errors of Agriculture ranged from 22% to 8%, and were also mostly composed of Forest pixels (agriculture pixels misclassified as forest).
@@
 
-High heterogeneity inside agriculture plots (sample 56), caused by both year of deforestation or year of agriculture establishment.
-Roads being considered as agriculture (sample 2).
-Omissions were detected, specially agriculture plots not being considered as a conversion as a whole, parts of a plot that were clearly a conversion were not considered as a conversion (sample 16, 30, 75).
-Inclusions were detected, areas that clearly showed no agriculture were considered as conversions (sample 10, 98).
-Effect of border on results (sample 25, 56, 84, 92).
-Areas with sparse forest vegetation presented a large error of deforestation year (sample 33).
+The qualitative assessment of 100 samples was performed to analyze the results in conjunction with the error calculations and metrics.
+It allows the interpretation of results with more context, since results are compared visually with reference satellite image composites, in an area of 4 square kilometers for each sample.
 
-Patterns of agricultural areas are roughly well represented (sample 25, 43, 56, 84).
-Some locations present a better homogeneity of conversions inside agriculture plots (sample 84).
+We found high heterogeneity inside agriculture plots (sample 56), caused by both year of deforestation or year of agriculture establishment.
+This heterogeneity may be caused by smaller uncertainties within properties, where it would be expected higher homogeneity of conversion length values.
+In few samples, roads were considered as agriculture (sample 2), so there was in fact no conversion to agriculture.
+Other inclusion errors were detected, where areas that clearly showed no agriculture were considered as conversions (sample 10, 98).
+There were also omissions errors, especially agriculture plots not being considered as a conversion as a whole, where parts of a plot that were clearly a conversion to agriculture were not considered as one (sample 16, 30, 75).
+A lot of uncertainty was identified at borders of agricultural areas (sample 25, 56, 84, 92), which is a very common issue in classification of satellite imagery, since there are the presence of mixed pixels, where different objects share the space of a same pixel [@Kaur2019].
+We also observed that areas with sparse forest vegetation presented a large error of deforestation year (sample 33).
 
-Interference due to atmospheric conditions and lack of image availability clearly shaped spatial patterns of conversions at some locations.
-Quality of composites through the time series affected accuracy of results.
-Regions of the North part of the Amazon present lower quality than regions in the South of the Amazon biome.
+We also observed that in many samples, the patterns of agricultural areas were roughly well represented (sample 25, 43, 56, 84), the shape of agricultural plots were clear and it was possible to distinguish different agricultural areas.
+Some locations present a higher homogeneity of conversions inside agriculture plots (sample 84), which may be related to areas with smaller uncertainties.
 
 \conclusions[Conclusions]
+
+The conversion from forests to agriculture estimates can be a valuable source of information to support research over different disciplines related to LULC changes.
+We showed that our data relates to many historical events in Brazil, which shaped the landscape of the Brazilian Amazon biome, covering public policies, economic and environmental aspects.
+The data is structured and stored in a ready to analyze format, which is very convenient to be analyzed in conjunction with economic and demographic data.
+
+Our data also explores new aspects of the MapBiomas classification data, which was not yet analyzed, by explicitly linking deforestation with agricultural establishment.
+The accuracy assessment creates new opportunities for future improvements of LULC data, especially concerning conversions from forest to agriculture.

### 2022-12-05 manuscript/manuscript.tex
Many insertions to the text. Inclusion of conclusion. Improvement of results. Fix typos. Update figures.

@@
 \begin{abstract}
-Land-use and land-cover changes in the Amazon biome are a main process that influences the environment and societies at local, national, and global levels.
+Land-use and land-cover changes in the Amazon biome are main processes that influence the environment and societies at local, national, and global levels.
 A large quantity of studies already relied on land-use classification to analyze this process.
@@
 Land-use and land-cover (LULC) changes are known to cause impacts on different scales.
-Deforestation and land use changes can affect hydrological processes in large basins \citep{Arias2018}; reduce the forest resilience to extreme events (e.g., drought) and other sources of perturbation \citep{Boulton2022}; deforestation and farming practices can reduce convective rainfall and increase surface temperature \citep{Maeda2021}; increase carbon emissions \citep{Gatti2021}; cause important impacts on fauna and flora diversity, soil properties and carbon stocks \citep{Nunes2022, Rittl2017}; and even negatively impact public health \citep{Ellwanger2020}.
+Deforestation and LULC changes can affect hydrological processes in large basins \citep{Arias2018}; reduce the forest resilience to extreme events (e.g., drought) and other sources of perturbation \citep{Boulton2022}; deforestation and farming practices can reduce convective rainfall and increase surface temperature \citep{Maeda2021}; increase carbon emissions \citep{Gatti2021}; cause important impacts on fauna and flora diversity, soil properties and carbon stocks \citep{Nunes2022, Rittl2017}; and even negatively impact public health \citep{Ellwanger2020}.
 
@@
 The second part of the processing was to load each tile separately, convert it to a table and calculate every possible conversion from forest to agriculture (Figure \ref{fig:routine-plot}).
-The conversions were calculated pixel by pixel, where the LULC time series were extracted and analysed for each conversion cycle that was present for that pixel.
-The calculation of the conversion length was extremely simple, it is the year of the establishment of agriculture subtracted by the last year classified as a forest, however, we add one year to the latter, so that it represents the year of deforestation (since it is more meaningful for analysis).
+The conversions were calculated pixel by pixel, where the LULC time series were extracted and analyzed for each conversion cycle that was present for that pixel.
+The calculation of the conversion length was simple, it is the year of the establishment of agriculture subtracted by the last year classified as a forest, however, we add one year to the latter, so that it represents the year of deforestation (since it is more meaningful for analysis).
 
-\begin{figure}[ht]
+\begin{figure}[h]
 \includegraphics[width=17cm]{figs/change_lenght_routine_scheme} \caption{Representation of the process to calculate the conversion length from forests to agriculture.
-All values presented are hiphothetical, and serves only as an ilustration of the method.}\label{fig:routine-plot}
+All values presented are hyphothetical, and serve only as an illustration of the method.}\label{fig:routine-plot}
 \end{figure}
 
-The process of calculation of the conversion length (Figure \ref{fig:routine-plot}), can be breakdown to seven steps of processing:
+The process of calculation of the conversion length (Figure \ref{fig:routine-plot}), can be divided into seven steps of processing:
 
@@
   Classify rows as ``before'' or ``after'' the occurrence of the year calculated in step 3.
-  This is the identification of periods before and after the recurrence of forest after a first conversion to agriculture;
+  This is the identification of periods before and after a return of forest after a first conversion to agriculture;
 \item
@@
 The steps 2 to 7 are performed recursively to identify multiple conversions, in case they are present.
-In addition to the conversion length values and the years of the conversion, they are also qualified by the type of agriculture that were established after deforestation, if the conversion occurred from primary or secondary forests, and the number of recurrences of conversions at the pixel.
+In addition to the conversion length values and the years of the conversion, they are also qualified by the type of agriculture that were established after deforestation, if the conversion occurred from primary or secondary forests, and the number of occurrences of conversions at the pixel.
 The LULC classes within the conversion and until 5 years after were also stored.
@@
 
-\begin{figure}[ht]
+\begin{figure}[h]
 \includegraphics[width=17cm]{figs/tables_scheme} \caption{ Representation of the tables generated by the conversion length calculation.
-All values presented are hiphothetical, and serves only as an ilustration of the method.}\label{fig:tables-plot}
+All values presented are hyphothetical, and serve only as an illustration of the method.}\label{fig:tables-plot}
 \end{figure}
 
-The first type of table (Figure \ref{fig:tables-plot}.a) contain the spatial information (longitude, latitude and area) of each pixel, its unique identificators and the code of the municipality which contains the pixel.
+The first type of table (Figure \ref{fig:tables-plot}.a) contains the spatial information (longitude, latitude and area) of each pixel, its unique identification and the code of the municipality which contains the pixel.
 The ``cell\_id'' column contains a unique identification for the whole data, while the ``tile\_cell\_id'' holds a unique value of each pixel for a single tile.
 Each tile is also uniquely identified in the column ``tile\_id''.
-This table is named as ``mask\_cells''.
+This table is named ``mask\_cells''.
 
@@
 The ``forest\_year'' have values that represent the year of deforestation, the ``agri\_year'' is the year when the agriculture establishment happened, and the subtraction between both (the conversion length) is stored in the ``c\_length'' column.
-Since it is possible that one pixels contains more than one conversion from forest to agriculture, we stored the number of the conversion in the column ``c\_cycle'' (1 is the first conversion cycle).
+Since it is possible that one pixel contains more than one conversion from forest to agriculture, we stored the number of the conversion in the column ``c\_cycle'' (1 is the first conversion cycle).
 However, conversion from forest to other LULC is also possible, so the column ``f\_cycle'' holds information about the number of the conversion from forest to any LULC.
-If the forest have suffered any conversion, it starts being considered as a secondary forest, as indicated in the column ``forest\_type''.
+If the forest has suffered any conversion, it starts being considered as a secondary forest, as indicated in the column ``forest\_type''.
 
@@
 The tabular data was organized in a system of folders and sub folders in a hive format partitioning (Figure \ref{fig:dataset-plot}).
-This structure allows for easily access parts of the data set, and also filter it without the necessity to read the tables.
+This structure allows for easily accessing parts of the data set, and also to filter it without the necessity to read the tables.
 In the case of this data set, it is partitioned by the ``tile\_id'' and ``agri\_cycle'' variables.
-The data set also have a metadata file with the spatial characteristics of each file, it stores the identification of the tile, the maximum and minimum latitude and longitude, the coordinate reference system, and the number of columns, rows, and cells of each tile.
+The data set also has a metadata file with the spatial characteristics of each file, it stores the identification of the tile, the maximum and minimum latitude and longitude, the coordinate reference system, and the number of columns, rows, and cells of each tile.
 
-\begin{figure}[ht]
+\begin{figure}[h]
 \includegraphics[width=17cm]{figs/dataset_structure} \caption{ Representation of the data set, structured in folders and sub folders.}\label{fig:dataset-plot}
@@
 
-\begin{figure}[ht]
+\begin{figure}[h]
 \includegraphics[width=17cm]{figs/map} \caption{Map of distribution of conversions from forest to agriculture in the Brazilian Amazon biome.
@@
 Transitions are concentrated in the south (Mato Grosso state), and in the east (Pará and Maranhão states).
-The conversion length ranges from 0 (dark blue) to 35 years (light pink), and clusters of fast conversions (conversions closer to 1 year) can be discerned from clusters of slow conversions (conversions closer to 36 years).
+The conversion length ranges from 0 (dark blue) to 35 years (light pink), and clusters of fast conversions (conversions closer to 0 year) can be discerned from clusters of slow conversions (conversions closer to 35 years).
 The histogram located in the bottom left shows that fast conversions are more common than slower conversions.
@@
 When observing at the original scale, we could not spot any well defined pattern or direction of the occurrence of faster to slower conversions (Figure \ref{fig:map-plot}).
+Other studies investigating patterns of conversions in the Amazon also found the formation of clusters of patterns, although there is a big heterogeneity at larger scales \citep{MullerHansen2017}.
 
-Other studies investigating patterns of conversions in the Amazon also found the formation of clusters of patterns, although there is a big heterogeneity at larger scales \citep{MullerHansen2017}.
 The data can be analyzed year by year, and also be separated by primary and secondary forests being converted to agriculture (Figure \ref{fig:transbar-plot}).
 
-\begin{figure}[ht]
+\begin{figure}[h]
 \includegraphics[width=17cm]{figs/trans_length_cols} \caption{Transition area from forest to agriculture, per year and conversion length.
 The bars represent the total amount of area at some state of the conversion for each year.
-The color gradient in each bar represents the conversion length related to a deforestation or an agriculture establishment event.
+The color gradient in each bar represent the conversion length related to a deforestation or an agriculture establishment event.
 Blue tones represents fast conversions (closer to 0 years), pink tones represent slow conversions (closer to 35 years).
@@
 The deforestation area of primary forests increased largely from 1986 to 2003, which was followed by a steep decrease until 2009, when the deforested areas reached a stable rate (Figure \ref{fig:transbar-plot}.a).
-From 1986 to approximately 1995, most of the deforested areas suffered a slow conversion, mostly were higher than 10 years.
-After this period, fast conversions started to become more common, specially from 2002 to 2004 (Figure \ref{fig:transbar-plot}.a).
-The peak of deforestation in 2003, according to our results, shows that 12.6\% of the conversions were a direct conversion from forest to agriculture.
-Previous studies estimated a proportion of 23\% \citep{Morton2006}.
-If we include fast conversions (from 0 to 2 years) from our results, the proportion jumps to 48.5 \% of all conversions.
+From 1985 to 1995, most deforested areas suffered a slow conversion, mostly were higher than 10 years (76 \%).
+After this period, fast conversions started to become more common (Figure \ref{fig:transbar-plot}.a), specially from 2002 to 2004, where 70 \% of the transitions were faster than 10 years.
+Our results show that in 2003 (the peak of deforestation), 12.7\% of the conversions were a direct conversion from forest to agriculture.
+Previous studies estimated a proportion of 23 \% in the Mato Grosso state, for the same year \citep{Morton2006}.
+If we include fast conversions (from 0 to 2 years) from our results, the proportion jumps to 48.1 \% of all conversions in 2003.
 
 Agricultural establishment over areas of primary forests peaked in 2005 and 2013 (Figure \ref{fig:transbar-plot}.b).
-Despite similar agricultural establishment rates between both years, their conversion lengths differ greatly, in 2005 most of the conversions were faster than 10 years, while in 2013 the great majority of conversions were slower than 10 years (Figure \ref{fig:transbar-plot}.b).
-In 2003 it was observed a shift in the conversion length from forest until the establishment of agricultural areas.
-From this year onward, most conversions happened in areas deforested at least 10 years before (Figure \ref{fig:transbar-plot}.b).
+Despite similar agricultural establishment areas in both years, their conversion lengths differ greatly (Figure \ref{fig:transbar-plot}.b), in 2005 most of the conversions were faster than 10 years (78 \%), while in 2013 the majority of conversions were slower than 10 years (66 \%).
+In 2010, it was observed a shift in the conversion length from forest until the establishment of agricultural areas.
+From this year onward, 68 \% of the conversions happened in areas deforested at least 10 years before (Figure \ref{fig:transbar-plot}.b).
 Even after the decrease of deforestation after 2002, areas under agriculture expanded over lands where deforestation happened before 2002 (mostly over pastureland).
 However, after 2019, a sudden drop of agriculture establishment rate happened (Figure \ref{fig:transbar-plot}.b).
-
-The period between 2001 and 2005 showed a peak of deforestation destined to agriculture (Figure \ref{fig:transbar-plot}.a), and of croplands expansion (Figure \ref{fig:transbar-plot}.b).
-From 2006 to 2010, we observed a decrease in both variables, a process that was already observed between 2001 and 2010 \citep{Macedo2012}, which points out a decoupling of soybean expansion and deforestation.
-However, our result shows that after 2011, agricultural areas rapidly expanded in the amazon biome, especially at areas deforested many years before.
+This behavior could be related to economic causes or even the scarcity of suitable areas for agricultural expansion, which seems unlikely, since the last years have been marked by an increase of deforestation rates and the advance of pastureland over the forest.
 
@@
 In the late 1990s, the development policies ``Brasil em Ação'' and ``Avança Brasil'' (1995 - 2003) accelerated the national infrastructure expansion, including in the Amazon \citep{Carvalho2002}.
-This is a period when the deforestation areas to be converted to agriculture reached the highest values in the time series (Figure \ref{fig:transbar-plot}.a), and also a peak in the establishment of agriculture in the Amazon biome, predominantly with fast conversions (Figure \ref{fig:transbar-plot}.b).
+This is a period when the deforestation areas reached the highest values in the time series (Figure \ref{fig:transbar-plot}.a), and also a peak in the establishment of agriculture in the Amazon biome, predominantly with fast conversions (Figure \ref{fig:transbar-plot}.b).
 
@@
 
-In 2006, the Brazilian Association of Vegetable Oil Industries (BIOVE) and the National Association of Cereal Exporters (ANEC) committed to avoid commercialization of soybean grains harvested from areas deforested after 2008.
-Our estimates of conversions show a decrease of deforested areas to be converted to agriculture after 2008, where it reached minimal values (Figure \ref{fig:transbar-plot}.a).
-After 2006, agriculture establishment over deforested primary forests suffered a decrease, which stayed relatively stable until 2012, where a steep increase occurred, however, the new areas being occupied by agriculture were mainly over areas that were cleared more than a decade before (therefore, before 2006) (Figure \ref{fig:transbar-plot}.b) This shows that agriculture expansion did not halt after the soy moratorium, but producers expended over areas cleared many years ago, which may have indirect impacts on deforestation increase over distant areas \citep{Arima2011}.
-However, this indirect effect is not always clearly observed when analyzing the Amazon biome as a whole, but our data shows that periods of fast agriculture expansion occurred at the same time with minimal deforestation values, according to PRODES estimates (add reference).
+In 2006, the Brazilian Association of Vegetable Oil Industries (BIOVE) and the National Association of Cereal Exporters (ANEC) committed to avoid commercialization of soybean grains harvested from areas deforested after 2008, known as the soy moratorium.
+The soy moratorium is considered to have contributed to the decline of the deforestation directly related to soybean expansion \citep{Paim2021, Amaral2021, Heilmayr2020, Kastens2017}.
+Our estimates of conversions show a decrease of deforested areas after 2008, where it reached minimal values (Figure \ref{fig:transbar-plot}.a).
+After 2006, agriculture establishment over deforested primary forests suffered a decrease, which stayed relatively stable until 2012, where a steep increase occurred, however, the new areas being occupied by agriculture were mainly over areas that were cleared before 2008 (Figure \ref{fig:transbar-plot}.b).
+The period between 2006 and 2010 is also considered as a turnover point, known as the decoupling of deforestation and soybean production, where the total production was not as dependent on agriculture expansion as before \citep{Macedo2012}.
+Therefore, the observed increase of agricultural establishment after 2012 shows that agriculture expansion did not halt after the soy moratorium nor after periods of high soybean productivity.
+Producers started to expand over areas cleared many years ago, which may still have indirect impacts on deforestation increase over distant areas \citep{Arima2011, Gollnow2018}.
+However, this indirect effect is not always clearly observed when analyzing the Amazon biome as a whole, our data shows that periods of fast agriculture expansion occurred at the same time with minimal deforestation values, according to PRODES estimates \citep{Assis2019}.
 Expansion of agricultural areas also occurred over cleared areas of secondary forests in 2012, but with an important amount of fast conversions (Figure \ref{fig:transbar-plot}.d).
-The causes of the increase of agricultural establishment areas can be numerous, and one main driver was the approval of a new Forest Code, in 2012, which is considered to have undermined the environmental protection of forests \citep{Kroger2017, Pereira2019}.
+The causes of the increase of agricultural establishment areas can be numerous, and one main driver was the approval of a new Forest Code, in 2012, which is considered to have undermined the environmental protection of forests \citep{Kroger2017, Pereira2019}, since it opened the possibility of amnesty for illegally deforested areas \citep{Schielein2018, Santanna2021, Filho2014}, and the possible reduction of legal reserve area \citep{Freitas2018}.
 
-The conversion length patterns across years can change significantly between different states (Figure \ref{fig:transridge-plot}).
+Even though it is possible to analyze the results in the Amazon as a whole, the conversion length patterns across years can change significantly between different states (Figure \ref{fig:transridge-plot}).
 
-\begin{figure}[ht]
+\begin{figure}[h]
 \includegraphics[width=17cm]{figs/trans_ridge} \caption{Transition length patterns inside states in the Amazon biome.
@@
 The peak of the curves represents conversion length values with more frequency in one year of one state.
-Blue tones represents fast conversions (conversions closer to 0 years), pink tones represents slow conversions (conversions closer to 35 years)}\label{fig:transridge-plot}
+Blue tones represent fast conversions (conversions closer to 0 years), pink tones represent slow conversions (conversions closer to 35 years).}\label{fig:transridge-plot}
 \end{figure}
@@
 
-\begin{figure}[ht]
+\begin{figure}[h]
 \includegraphics[width=17cm]{figs/error_bars} \caption{Bar with the count of error values (difference between observed and estimated values).
@@
 
-\begin{figure}[ht]
+\begin{figure}[h]
 \includegraphics[width=17cm]{figs/error_scatter} \caption{Scatter plot between paired values of estimated and observed Deforestation and Agriculture Establishment years, and their respective Transition Length.
-The gray line represents values which there would be no errors of the estimated values in comparison with the observed values.
+The gray line represents values where there would be no errors of the estimated values in comparison with the observed values.
 Translucid white areas represent quantile ranges (5\%, 25\%, 75\% and 95\%) of the errors.
@@
 The omission errors for Forests are also small (proportion of pixels misclassified as Forests, but the real class were not Forest), which fluctuated around 2\%.
-The Agriculture class presented more errors, the inclusion errors ranged from 22\% to 5\%, and were mostly composed by Forests pixels (forest pixels misclassified as agriculture).
-The omission errors of Agriculture ranged from 22\% to 8\%, and were also mostly composed of Forest pixels (agriculture pixels misclassified as Agriculture).
+The Agriculture class presented more errors, the inclusion errors ranged from 22\% to 5\%, and were mostly composed of Forests pixels (forest pixels misclassified as agriculture).
+The omission errors of Agriculture ranged from 22\% to 8\%, and were also mostly composed of Forest pixels (agriculture pixels misclassified as forest).
 
@@
 
-High heterogeneity inside agriculture plots (sample 56), caused by both year of deforestation or year of agriculture establishment.
-Roads being considered as agriculture (sample 2).
-Omissions were detected, specially agriculture plots not being considered as a conversion as a whole, parts of a plot that were clearly a conversion were not considered as a conversion (sample 16, 30, 75).
-Inclusions were detected, areas that clearly showed no agriculture were considered as conversions (sample 10, 98).
-Effect of border on results (sample 25, 56, 84, 92).
-Areas with sparse forest vegetation presented a large error of deforestation year (sample 33).
+The qualitative assessment of 100 samples was performed to analyze the results in conjunction with the error calculations and metrics.
+It allows the interpretation of results with more context, since results are compared visually with reference satellite image composites, in an area of 4 square kilometers for each sample.
 
-Patterns of agricultural areas are roughly well represented (sample 25, 43, 56, 84).
-Some locations present a better homogeneity of conversions inside agriculture plots (sample 84).
+We found high heterogeneity inside agriculture plots (sample 56), caused by both year of deforestation or year of agriculture establishment.
+This heterogeneity may be caused by smaller uncertainties within properties, where it would be expected higher homogeneity of conversion length values.
+In few samples, roads were considered as agriculture (sample 2), so there was in fact no conversion to agriculture.
+Other inclusion errors were detected, where areas that clearly showed no agriculture were considered as conversions (sample 10, 98).
+There were also omissions errors, especially agriculture plots not being considered as a conversion as a whole, where parts of a plot that were clearly a conversion to agriculture were not considered as one (sample 16, 30, 75).
+A lot of uncertainty was identified at borders of agricultural areas (sample 25, 56, 84, 92), which is a very common issue in classification of satellite imagery, since there are the presence of mixed pixels, where different objects share the space of a same pixel \citep{Kaur2019}.
+We also observed that areas with sparse forest vegetation presented a large error of deforestation year (sample 33).
 
-Interference due to atmospheric conditions and lack of image availability clearly shaped spatial patterns of conversions at some locations.
-Quality of composites through the time series affected accuracy of results.
-Regions of the North part of the Amazon present lower quality than regions in the South of the Amazon biome.
+We also observed that in many samples, the patterns of agricultural areas were roughly well represented (sample 25, 43, 56, 84), the shape of agricultural plots were clear and it was possible to distinguish different agricultural areas.
+Some locations present a higher homogeneity of conversions inside agriculture plots (sample 84), which may be related to areas with smaller uncertainties.
 
 \conclusions[Conclusions]
+
+The conversion from forests to agriculture estimates can be a valuable source of information to support research over different disciplines related to LULC changes.
+We showed that our data relates to many historical events in Brazil, which shaped the landscape of the Brazilian Amazon biome, covering public policies, economic and environmental aspects.
+The data is structured and stored in a ready to analyze format, which is very convenient to be analyzed in conjunction with economic and demographic data.
+
+Our data also explores new aspects of the MapBiomas classification data, which was not yet analyzed, by explicitly linking deforestation with agricultural establishment.
+The accuracy assessment creates new opportunities for future improvements of LULC data, especially concerning conversions from forest to agriculture.
 

### 2022-12-21 manuscript/manuscript.Rmd
Update abstract.

@@
 abstract: |
-  Land-use and land-cover changes in the Amazon biome are main processes that influence the environment and societies at local, national, and global levels.
-A large quantity of studies already relied on land-use classification to analyze this process.
+  Land-use and land-cover changes in the Amazon biome are main processes that influence the environment and societies at local, national, and global scales.
+A large quantity of studies already relied on land thematic maps to analyze change processes.
 This study brings a new dataset created by calculating the time necessary for deforested areas to change to agriculture (annual and permanent crops) in the Brazilian Amazon biome.
-The calculations were performed over high resolution land-use classification by using simple algebra and recursion to identify every conversion from forest to agriculture, and ranges from 1985 to 2021.
+The calculations were performed over thematic maps of high resolution by using simple algebra and recursion to identify every conversion from forest to agriculture between 1985 and 2021.
 The new data can be useful in interdisciplinary studies about land-use and land-cover change analysis in Brazil, such as exploring its drivers and evaluating the effects on natural vegetation cover and ecosystem services.

### 2022-12-21 manuscript/manuscript.tex
Update abstract.

@@
 \begin{abstract}
-Land-use and land-cover changes in the Amazon biome are main processes that influence the environment and societies at local, national, and global levels.
-A large quantity of studies already relied on land-use classification to analyze this process.
+Land-use and land-cover changes in the Amazon biome are main processes that influence the environment and societies at local, national, and global scales.
+A large quantity of studies already relied on land thematic maps to analyze change processes.
 This study brings a new dataset created by calculating the time necessary for deforested areas to change to agriculture (annual and permanent crops) in the Brazilian Amazon biome.
-The calculations were performed over high resolution land-use classification by using simple algebra and recursion to identify every conversion from forest to agriculture, and ranges from 1985 to 2021.
+The calculations were performed over thematic maps of high resolution by using simple algebra and recursion to identify every conversion from forest to agriculture between 1985 and 2021.
 The new data can be useful in interdisciplinary studies about land-use and land-cover change analysis in Brazil, such as exploring its drivers and evaluating the effects on natural vegetation cover and ecosystem services.

### 2022-12-21 manuscript/manuscript.Rmd
Update introduction.

@@
 Brazilian policies since the 1960s were focused on the expansion of the agricultural frontier in the Amazon, specially focused in fostering economic growth and national security of the territory [@Carvalho2002; @Mcdonald2003; @Banerjee2009].
-Such a development model led to the construction of thousands of kilometers of roads, and the settlement of large scale crop and livestock farms [@Carvalho2002; @Banerjee2009].
-This period was marked with high rates of deforestation of mature forests [@Fearnside2005].
-At the same time, cattle and soybean production started to move from Southern Brazil into the Midwest and Northern parts of the country, in which soybean cultivation areas followed cattle expansion (i.e., pastureland expansion) over the Amazon forest [@Simon2005; @Barona2010; @Arima2011].
+Such a development model led to the construction of thousands of kilometers of roads, and the establishment of large scale crop and livestock farms [@Carvalho2002; @Banerjee2009].
+This period was marked by high rates of deforestation of mature forests [@Fearnside2005].
+At the same time, cattle and soybean production started to move from Southern Brazil into the Midwest and Northern parts of the country [@Silva2020a], in which soybean cultivation areas followed cattle expansion (i.e., pastureland expansion) over the Amazon forest [@Simon2005; @Barona2010; @Arima2011].
 The pattern of extensive livestock production after deforestation, followed by the establishment of annual crops after a given period is typical in the Amazon biome [@Barona2010], and persists nowadays.
@@
 Initiatives such as MapBiomas [@Souza2020], launched in 2015, provides high resolution, annual classifications for the Brazilian territory.
-Such types of data are essential to analyze LULC change processes, make causal inferences, and understand what drives and what are the impacts of the transformation of the landscape in Brazil.
-In this context, this study aims to quantify and characterize the length of the conversions from forest formations to agriculture in the Amazon over the last 36 years.
-The availability of this data can be useful to the development of interdisciplinary research, since it can provide useful insights into the environmental, economic, and social impacts of land conversion.
+Such types of data are essential to analyze LULC change processes [@Silva2020b], make causal inferences [@Capmourteres2018; @Gelo2022; @Odongo2014; @Larned2020], and to understand what drives and what are the impacts of the transformation of the landscapes in Brazil [@Nunes2022].
+In this context, this study aims to quantify and characterize the length (time span) of the conversions from forest formations to agriculture in the Amazon over the last 36 years.
+This this data can be useful to the development of interdisciplinary research, since it can provide useful insights into the environmental, economic, and social impacts of land conversion.
 For example, this data can be used to develop insights into the economic costs and benefits associated with land conversion, the effects of land conversion on local ecosystems, and the impact of land conversion on livelihoods and social dynamics.

### 2022-12-21 manuscript/manuscript.tex
Update introduction.

@@
 Brazilian policies since the 1960s were focused on the expansion of the agricultural frontier in the Amazon, specially focused in fostering economic growth and national security of the territory \citep{Carvalho2002, Mcdonald2003, Banerjee2009}.
-Such a development model led to the construction of thousands of kilometers of roads, and the settlement of large scale crop and livestock farms \citep{Carvalho2002, Banerjee2009}.
-This period was marked with high rates of deforestation of mature forests \citep{Fearnside2005}.
-At the same time, cattle and soybean production started to move from Southern Brazil into the Midwest and Northern parts of the country, in which soybean cultivation areas followed cattle expansion (i.e., pastureland expansion) over the Amazon forest \citep{Simon2005, Barona2010, Arima2011}.
+Such a development model led to the construction of thousands of kilometers of roads, and the establishment of large scale crop and livestock farms \citep{Carvalho2002, Banerjee2009}.
+This period was marked by high rates of deforestation of mature forests \citep{Fearnside2005}.
+At the same time, cattle and soybean production started to move from Southern Brazil into the Midwest and Northern parts of the country \citep{Silva2020a}, in which soybean cultivation areas followed cattle expansion (i.e., pastureland expansion) over the Amazon forest \citep{Simon2005, Barona2010, Arima2011}.
 The pattern of extensive livestock production after deforestation, followed by the establishment of annual crops after a given period is typical in the Amazon biome \citep{Barona2010}, and persists nowadays.
@@
 Initiatives such as MapBiomas \citep{Souza2020}, launched in 2015, provides high resolution, annual classifications for the Brazilian territory.
-Such types of data are essential to analyze LULC change processes, make causal inferences, and understand what drives and what are the impacts of the transformation of the landscape in Brazil.
-In this context, this study aims to quantify and characterize the length of the conversions from forest formations to agriculture in the Amazon over the last 36 years.
-The availability of this data can be useful to the development of interdisciplinary research, since it can provide useful insights into the environmental, economic, and social impacts of land conversion.
+Such types of data are essential to analyze LULC change processes \citep{Silva2020b}, make causal inferences \citep{Capmourteres2018, Gelo2022, Odongo2014, Larned2020}, and to understand what drives and what are the impacts of the transformation of the landscapes in Brazil \citep{Nunes2022}.
+In this context, this study aims to quantify and characterize the length (time span) of the conversions from forest formations to agriculture in the Amazon over the last 36 years.
+This this data can be useful to the development of interdisciplinary research, since it can provide useful insights into the environmental, economic, and social impacts of land conversion.
 For example, this data can be used to develop insights into the economic costs and benefits associated with land conversion, the effects of land conversion on local ecosystems, and the impact of land conversion on livelihoods and social dynamics.

### 2022-12-22 manuscript/manuscript.Rmd
Update methods. Add some information about method. Fixed some sentences.

@@
 
-The estimations of conversion from forest to agriculture were performed for the Amazon biome region (biome limits defined by the Brazilian Institute of Geography and Statistics (IBGE), in 2019).
-Transitions were calculated using LULC classification data from MapBiomas, collection 7, which ranges from 1985 to 2021, at a spatial resolution of approximately 30 meters, and global accuracy at 91.3%.
+The estimations of conversion from forest to agriculture were performed for the Amazon biome (biome limits defined by the Brazilian Institute of Geography and Statistics (IBGE), in 2019).
+Conversions were calculated using LULC classification data from MapBiomas, collection 7, which ranges from 1985 to 2021, at a spatial resolution of approximately 30 meters, and global accuracy at 91.3%.
+The calculations considers forests as the "Forest Formation" (class code 3) class from MapBiomas, and does not include other types of natural vegetation.
+The agriculture is composed by eight classes from MapBiomas, "Soybean", "Sugar Cane", "Rice", "Cotton", "Other Temporary Crops", "Coffee", "Citrus", "Other Perennial Crops" (class code 39, 20, 40, 62, 41, 46, 47 and 48, respectively).
+The conversion length (time span) was considered as how many years there were between the deforestation and the establishment of agriculture in the same area.
 
 The MapBiomas data was downloaded from Google Earth Engine platform [@Gorelick2017].
-The pixels of the Amazon biome were filtered to contain only those that were occupied by forest and agriculture at some period, and that are not considered as water, according to the Global Surface Water product provided by Copernicus [@Pekel2016].
+The pixels within the Amazon biome were filtered to contain only those that were occupied by forest and agriculture in the period, and that are not considered as water, according to the Global Surface Water product provided by Copernicus [@Pekel2016].
 For the download, the data was divided in 2732 tiles, to allow local processing of a large quantity of pixels.Two sets of data were downloaded with the same spatial dimensions, one with the LULC classifications for each year, and another with binary values, representing whether that pixels were valid for processing or not (according to the filter conditions mentioned above).
 
-The first part of the processing was to create a table from the binary values data, and create unique identification for each pixel, their coordinates, calculate the area, and also the municipality in which each pixel was inside.
-In this part, a table was also generated with the spatial information of each tile, so that it is possible to convert the generated tables back to raster format.
+The first part of the processing was to create a table (*mask_cells*) from the binary data, and set a unique identification for each pixel, along with their coordinates, area (square meters), and also the respective municipality of each pixel.
+In this part, an additional metadata table (*tiles_metadata*) was generated with the spatial information of each tile, so that it is possible to convert the generated tables back to raster format.
 
-The second part of the processing was to load each tile separately, convert it to a table and calculate every possible conversion from forest to agriculture (Figure \@ref(fig:routine-plot)).
-The conversions were calculated pixel by pixel, where the LULC time series were extracted and analyzed for each conversion cycle that was present for that pixel.
-The calculation of the conversion length was simple, it is the year of the establishment of agriculture subtracted by the last year classified as a forest, however, we add one year to the latter, so that it represents the year of deforestation (since it is more meaningful for analysis).
+The second part of the processing was to load each raster tile separately, convert it to a new table (*lulc_table*) and calculate every possible conversion from forest to agriculture (Figure \@ref(fig:routine-plot)).
+The calculation of the conversion length (time span) was simple.
+It consisted of the year of the agriculture establishment subtracted by the last year classified as forest formation.
+Than, we add one year to the latter, so that it represents the year of deforestation (since it is more meaningful for analysis).
+For example, considering a conversion where the last forest year was 1986, and than turned to agriculture in 1989.
+In this case, we would add one year to the last forest year (1986 + 1) and than subtract the agriculture establishment year (1989) by it, the result would be 1989 subtracted by 1987, which returns a conversion length of 2 years (Figure \@ref(fig:routine-plot)).
 
 ```{r routine-plot, out.width = "17cm", fig.pos = "h", echo = FALSE, fig.cap = "Representation of the process to calculate the conversion length from forests to agriculture.
+Numbers inside cricles represent the setp of the process to calculate the conversion length (1 to 7).
 All values presented are hyphothetical, and serve only as an illustration of the method."}
@@
 
-2.  Calculate the year of first occurrence of any agriculture class (annual and permanent crops) for each pixel;
+2.  Calculate the year of first occurrence of any anthropic class for each pixel;
 
@@
 
-The steps 2 to 7 are performed recursively to identify multiple conversions, in case they are present.
-In addition to the conversion length values and the years of the conversion, they are also qualified by the type of agriculture that were established after deforestation, if the conversion occurred from primary or secondary forests, and the number of occurrences of conversions at the pixel.
-The LULC classes within the conversion and until 5 years after were also stored.
+The steps 2 to 7 are performed recursively to identify multiple cycles of conversions (e.g., forest -\> agriculture -\> forest -\> agriculture), in case they are present.
+In addition to the conversion length values and the years of the conversion, they are also qualified by the agriculture class that were established after deforestation, if the conversion occurred from primary or secondary forests, and the number of cycles of conversions at the pixel.
+The LULC classes between the deforestation and the agriculture establishment, and 5 years after the conversion, were also stored.
 
@@
 Each tile is also uniquely identified in the column "tile_id".
-This table is named "mask_cells".
+This table is named "*mask_cells*".
 
 A second type of tables (Figure \@ref(fig:tables-plot).b) stores conversion length values, the first and last year of the conversion, the resulting agriculture type, and the number of the conversion cycle.
-They also contain the unique id of each pixel and tile (related to the table above).
+They also contain the unique id of each pixel and tile (same as the table above).
 The "agri_code" column holds information about the type of agriculture the was established after deforestation.
@@
 Since it is possible that one pixel contains more than one conversion from forest to agriculture, we stored the number of the conversion in the column "c_cycle" (1 is the first conversion cycle).
-However, conversion from forest to other LULC is also possible, so the column "f_cycle" holds information about the number of the conversion from forest to any LULC.
+However, conversion from forest to other LULC is also possible, so the column "f_cycle" holds information about the number of the conversion from forest to any LULC except agriculture.
 If the forest has suffered any conversion, it starts being considered as a secondary forest, as indicated in the column "forest_type".
+This table is named "*c_length*".
 
-The third type of tables (Figure \@ref(fig:tables-plot).c) contains the LULC classes of all years within the conversion, and also the first 5 years after the conversion.
+The third type of tables (Figure \@ref(fig:tables-plot).c) contain the LULC classes of all years between the deforestation and the agriculture establishment, and also the first 5 years after the conversion.
 The column "year" holds the year of the MapBiomas classification, and the "class_code" variable stores the LULC classification (according to MapBiomas) for the respective year.
+This table is named "*c_classes*".
 

### 2022-12-22 manuscript/manuscript.tex
Update methods. Add some information about method. Fixed some sentences.

@@
 
-The estimations of conversion from forest to agriculture were performed for the Amazon biome region (biome limits defined by the Brazilian Institute of Geography and Statistics (IBGE), in 2019).
-Transitions were calculated using LULC classification data from MapBiomas, collection 7, which ranges from 1985 to 2021, at a spatial resolution of approximately 30 meters, and global accuracy at 91.3\%.
+The estimations of conversion from forest to agriculture were performed for the Amazon biome (biome limits defined by the Brazilian Institute of Geography and Statistics (IBGE), in 2019).
+Conversions were calculated using LULC classification data from MapBiomas, collection 7, which ranges from 1985 to 2021, at a spatial resolution of approximately 30 meters, and global accuracy at 91.3\%.
+The calculations considers forests as the ``Forest Formation'' (class code 3) class from MapBiomas, and does not include other types of natural vegetation.
+The agriculture is composed by eight classes from MapBiomas, ``Soybean'', ``Sugar Cane'', ``Rice'', ``Cotton'', ``Other Temporary Crops'', ``Coffee'', ``Citrus'', ``Other Perennial Crops'' (class code 39, 20, 40, 62, 41, 46, 47 and 48, respectively).
+The conversion length (time span) was considered as how many years there were between the deforestation and the establishment of agriculture in the same area.
 
 The MapBiomas data was downloaded from Google Earth Engine platform \citep{Gorelick2017}.
-The pixels of the Amazon biome were filtered to contain only those that were occupied by forest and agriculture at some period, and that are not considered as water, according to the Global Surface Water product provided by Copernicus \citep{Pekel2016}.
+The pixels within the Amazon biome were filtered to contain only those that were occupied by forest and agriculture in the period, and that are not considered as water, according to the Global Surface Water product provided by Copernicus \citep{Pekel2016}.
 For the download, the data was divided in 2732 tiles, to allow local processing of a large quantity of pixels.Two sets of data were downloaded with the same spatial dimensions, one with the LULC classifications for each year, and another with binary values, representing whether that pixels were valid for processing or not (according to the filter conditions mentioned above).
 
-The first part of the processing was to create a table from the binary values data, and create unique identification for each pixel, their coordinates, calculate the area, and also the municipality in which each pixel was inside.
-In this part, a table was also generated with the spatial information of each tile, so that it is possible to convert the generated tables back to raster format.
+The first part of the processing was to create a table (\emph{mask\_cells}) from the binary data, and set a unique identification for each pixel, along with their coordinates, area (square meters), and also the respective municipality of each pixel.
+In this part, an additional metadata table (\emph{tiles\_metadata}) was generated with the spatial information of each tile, so that it is possible to convert the generated tables back to raster format.
 
-The second part of the processing was to load each tile separately, convert it to a table and calculate every possible conversion from forest to agriculture (Figure \ref{fig:routine-plot}).
-The conversions were calculated pixel by pixel, where the LULC time series were extracted and analyzed for each conversion cycle that was present for that pixel.
-The calculation of the conversion length was simple, it is the year of the establishment of agriculture subtracted by the last year classified as a forest, however, we add one year to the latter, so that it represents the year of deforestation (since it is more meaningful for analysis).
+The second part of the processing was to load each raster tile separately, convert it to a new table (\emph{lulc\_table}) and calculate every possible conversion from forest to agriculture (Figure \ref{fig:routine-plot}).
+The calculation of the conversion length (time span) was simple.
+It consisted of the year of the agriculture establishment subtracted by the last year classified as forest formation.
+Than, we add one year to the latter, so that it represents the year of deforestation (since it is more meaningful for analysis).
+For example, considering a conversion where the last forest year was 1986, and than turned to agriculture in 1989.
+In this case, we would add one year to the last forest year (1986 + 1) and than subtract the agriculture establishment year (1989) by it, the result would be 1989 subtracted by 1987, which returns a conversion length of 2 years (Figure \ref{fig:routine-plot}).
 
@@
 \includegraphics[width=17cm]{figs/change_lenght_routine_scheme} \caption{Representation of the process to calculate the conversion length from forests to agriculture.
+Numbers inside cricles represent the setp of the process to calculate the conversion length (1 to 7).
 All values presented are hyphothetical, and serve only as an illustration of the method.}\label{fig:routine-plot}
@@
 \item
-  Calculate the year of first occurrence of any agriculture class (annual and permanent crops) for each pixel;
+  Calculate the year of first occurrence of any anthropic class for each pixel;
 \item
@@
 
-The steps 2 to 7 are performed recursively to identify multiple conversions, in case they are present.
-In addition to the conversion length values and the years of the conversion, they are also qualified by the type of agriculture that were established after deforestation, if the conversion occurred from primary or secondary forests, and the number of occurrences of conversions at the pixel.
-The LULC classes within the conversion and until 5 years after were also stored.
+The steps 2 to 7 are performed recursively to identify multiple cycles of conversions (e.g., forest -\textgreater{} agriculture -\textgreater{} forest -\textgreater{} agriculture), in case they are present.
+In addition to the conversion length values and the years of the conversion, they are also qualified by the agriculture class that were established after deforestation, if the conversion occurred from primary or secondary forests, and the number of cycles of conversions at the pixel.
+The LULC classes between the deforestation and the agriculture establishment, and 5 years after the conversion, were also stored.
 
@@
 Each tile is also uniquely identified in the column ``tile\_id''.
-This table is named ``mask\_cells''.
+This table is named ``\emph{mask\_cells}''.
 
 A second type of tables (Figure \ref{fig:tables-plot}.b) stores conversion length values, the first and last year of the conversion, the resulting agriculture type, and the number of the conversion cycle.
-They also contain the unique id of each pixel and tile (related to the table above).
+They also contain the unique id of each pixel and tile (same as the table above).
 The ``agri\_code'' column holds information about the type of agriculture the was established after deforestation.
@@
 Since it is possible that one pixel contains more than one conversion from forest to agriculture, we stored the number of the conversion in the column ``c\_cycle'' (1 is the first conversion cycle).
-However, conversion from forest to other LULC is also possible, so the column ``f\_cycle'' holds information about the number of the conversion from forest to any LULC.
+However, conversion from forest to other LULC is also possible, so the column ``f\_cycle'' holds information about the number of the conversion from forest to any LULC except agriculture.
 If the forest has suffered any conversion, it starts being considered as a secondary forest, as indicated in the column ``forest\_type''.
+This table is named ``\emph{c\_length}''.
 
-The third type of tables (Figure \ref{fig:tables-plot}.c) contains the LULC classes of all years within the conversion, and also the first 5 years after the conversion.
+The third type of tables (Figure \ref{fig:tables-plot}.c) contain the LULC classes of all years between the deforestation and the agriculture establishment, and also the first 5 years after the conversion.
 The column ``year'' holds the year of the MapBiomas classification, and the ``class\_code'' variable stores the LULC classification (according to MapBiomas) for the respective year.
+This table is named ``\emph{c\_classes}''.
 

### 2022-12-22 manuscript/manuscript.Rmd
Update methods. Fix typo.
Update results. Fix confuse sentences. Add some information. Add references.

@@
 They also contain the unique id of each pixel and tile (same as the table above).
-The "agri_code" column holds information about the type of agriculture the was established after deforestation.
+The "agri_code" column holds information about the type of agriculture that was established after deforestation.
 The "forest_year" have values that represent the year of deforestation, the "agri_year" is the year when the agriculture establishment happened, and the subtraction between both (the conversion length) is stored in the "c_length" column.
@@
 The conversion calculations show that in the Brazilian Amazon biome, 64874 square kilometers of forests were converted to agriculture between 1985 and 2021.
-The length of the conversions can go from 0 to 35 years, in which conversions closer to 0 years are considered as fast conversions, and conversions closer to 35 years are considered as slow conversions.
-Transitions of 0 years are considered "direct" conversions, where there was no presence of pasture before the establishment of agriculture.
+The length of the conversions can go from 0 to 35 years (time span), in which conversions closer to 0 years are considered as fast conversions, and conversions closer to 35 years are considered as slow conversions.
+Conversions of 0 years are considered "direct" conversions, where there was no presence of pasture before the establishment of agriculture.
 Our estimations show that around 9.2% of the conversions were "direct".
 
-Our conversions calculations found pixels that presented up to six conversions from forest to agriculture since 1985, as these numbers are unlikely to happen (they are distributed as sparse pixels, and do not show patterns of real conversions, being common at borders), we proceeded the results analysis only in the first conversions found in each of these pixels.
+Our results found pixels that presented up to six conversions (i.e., six cycles) from forest to agriculture since 1985.
+However, since these cycles are unlikely to happen (they are distributed as sparse pixels, and do not show patterns of real conversions, being common at borders), we proceeded the analysis considering only the first conversions found in each of these pixels (these pixels with multiple conversions represented 0.5% of the total converted areas in our study).
 
@@
 
-## Transition patterns
+## Conversion patterns
 
-Transitions from forest to agriculture can be found in almost every region in the Amazon, but is mostly concentrated in clusters, specially in the south and east of the biome, in the states of Mato Grosso, Pará and Maranhão (Figure \@ref(fig:map-plot)).
+Cnversions from forest to agriculture can be found in almost every region in the Amazon, but is mostly concentrated in clusters, specially in the south and east of the biome, in the states of Mato Grosso, Pará and Maranhão (Figure \@ref(fig:map-plot)).
 
@@
 The hexagonal cells represent the most common conversion length, and do not reflect the amount of area of conversions inside a cell.
-Transitions are concentrated in the south (Mato Grosso state), and in the east (Pará and Maranhão states).
+Conversions are concentrated in the south (Mato Grosso state), and in the east (Pará and Maranhão states).
 The conversion length ranges from 0 (dark blue) to 35 years (light pink), and clusters of fast conversions (conversions closer to 0 year) can be discerned from clusters of slow conversions (conversions closer to 35 years).
@@
 
-```{r transbar-plot, out.width = "17cm", fig.pos = "h", echo = FALSE, fig.cap = "Transition area from forest to agriculture, per year and conversion length.
+```{r transbar-plot, out.width = "17cm", fig.pos = "h", echo = FALSE, fig.cap = "Conversion area from forest to agriculture, per year and conversion length.
 The bars represent the total amount of area at some state of the conversion for each year.
@@
 Blue tones represents fast conversions (closer to 0 years), pink tones represent slow conversions (closer to 35 years).
-Transition events were separated by deforestation of primary forests (a) and secondary forests (c), and the subsequent agriculture establishment of primary forests (b) and secondary forests (d)."}
+Conversion events were separated by deforestation of primary forests (a) and secondary forests (c), and the subsequent agriculture establishment of primary forests (b) and secondary forests (d)."}
 
@@
 The deforestation area of primary forests increased largely from 1986 to 2003, which was followed by a steep decrease until 2009, when the deforested areas reached a stable rate (Figure \@ref(fig:transbar-plot).a).
-From 1985 to 1995, most deforested areas suffered a slow conversion, mostly were higher than 10 years (76 %).
-After this period, fast conversions started to become more common (Figure \@ref(fig:transbar-plot).a), specially from 2002 to 2004, where 70 % of the transitions were faster than 10 years.
-Our results show that in 2003 (the peak of deforestation), 12.7% of the conversions were a direct conversion from forest to agriculture.
+From 1985 to 1995, most deforested areas had a slow conversion to agriculture, mostly higher than 10 years (76 %).
+After this period (1985\--1995), fast conversions started to become more common (Figure \@ref(fig:transbar-plot).a), specially from 2002 to 2004, where 70 % of the conversions were faster than 10 years.
+Our results show that in 2003 (the peak of deforestation), 12.7 % of the conversions were a direct conversion from forest to agriculture.
 Previous studies estimated a proportion of 23 % in the Mato Grosso state, for the same year [@Morton2006].
-If we include fast conversions (from 0 to 2 years) from our results, the proportion jumps to 48.1 % of all conversions in 2003.
+If we include fast conversions (from 0 to 2 years), the proportion jumps to 48.1 % of all conversions in 2003.
 
 Agricultural establishment over areas of primary forests peaked in 2005 and 2013 (Figure \@ref(fig:transbar-plot).b).
-Despite similar agricultural establishment areas in both years, their conversion lengths differ greatly (Figure \@ref(fig:transbar-plot).b), in 2005 most of the conversions were faster than 10 years (78 %), while in 2013 the majority of conversions were slower than 10 years (66 %).
+Despite similar agricultural establishment areas in both years, their conversion lengths differ greatly (Figure \@ref(fig:transbar-plot).b).
+In 2005 most of the conversions were faster than 10 years (78 %), while in 2013 the majority of conversions took more than 10 years (66 %).
 In 2010, it was observed a shift in the conversion length from forest until the establishment of agricultural areas.
@@
 However, after 2019, a sudden drop of agriculture establishment rate happened (Figure \@ref(fig:transbar-plot).b).
-This behavior could be related to economic causes or even the scarcity of suitable areas for agricultural expansion, which seems unlikely, since the last years have been marked by an increase of deforestation rates and the advance of pastureland over the forest.
+This behavior could be related to economic causes [@Morton2006; @Gusso2017; @Hargrave2013] or even the scarcity of suitable areas for agricultural expansion, which seems unlikely, since the last years have been marked by an increase of deforestation rates and the advance of pastureland over the forest.
 
@@
 The PPCDAm was considered as a successful policy to slow deforestation rates in Brazil, with international recognition.
-Our calculations from MapBiomas data reinforces the correlation of the PPCDAm with the reduction of deforestation after 2004 (Figure \@ref(fig:transbar-plot).a), and reduction of the agriculture establishment after 2005 (Figure \@ref(fig:transbar-plot).b).
+Our calculations from MapBiomas data reinforces the correlation of the PPCDAm with the reduction of deforestation after 2004 [@West2021; @Arima2014] (Figure \@ref(fig:transbar-plot).a), and reduction of the agriculture establishment after 2005 (Figure \@ref(fig:transbar-plot).b).
 
@@
 The soy moratorium is considered to have contributed to the decline of the deforestation directly related to soybean expansion [@Paim2021; @Amaral2021; @Heilmayr2020; @Kastens2017].
-Our estimates of conversions show a decrease of deforested areas after 2008, where it reached minimal values (Figure \@ref(fig:transbar-plot).a).
-After 2006, agriculture establishment over deforested primary forests suffered a decrease, which stayed relatively stable until 2012, where a steep increase occurred, however, the new areas being occupied by agriculture were mainly over areas that were cleared before 2008 (Figure \@ref(fig:transbar-plot).b).
-The period between 2006 and 2010 is also considered as a turnover point, known as the decoupling of deforestation and soybean production, where the total production was not as dependent on agriculture expansion as before [@Macedo2012].
-Therefore, the observed increase of agricultural establishment after 2012 shows that agriculture expansion did not halt after the soy moratorium nor after periods of high soybean productivity.
-Producers started to expand over areas cleared many years ago, which may still have indirect impacts on deforestation increase over distant areas [@Arima2011; @Gollnow2018].
-However, this indirect effect is not always clearly observed when analyzing the Amazon biome as a whole, our data shows that periods of fast agriculture expansion occurred at the same time with minimal deforestation values, according to PRODES estimates [@Assis2019].
-Expansion of agricultural areas also occurred over cleared areas of secondary forests in 2012, but with an important amount of fast conversions (Figure \@ref(fig:transbar-plot).d).
-The causes of the increase of agricultural establishment areas can be numerous, and one main driver was the approval of a new Forest Code, in 2012, which is considered to have undermined the environmental protection of forests [@Kroger2017; @Pereira2019], since it opened the possibility of amnesty for illegally deforested areas [@Schielein2018; @Santanna2021; @Filho2014], and the possible reduction of legal reserve area [@Freitas2018].
+Our estimates of conversions show a decrease of deforested areas after 2008, where it reached minimal values (Figure @ref(fig:transbar-plot).a).
+After 2006, agriculture establishment over deforested primary forests suffered a decrease, the low rates of agriculture establishment stayed relatively stable until 2012, where a steep increase occurred.
+However, the new primary forest areas being occupied by agriculture were mainly over areas that were cleared before 2008 (Figure @ref(fig:transbar-plot).b), which is the last year of deforestation tolerated by the soy moratorium.
+The period between 2006 and 2010 is also considered as a turnover point, known as the decoupling of deforestation and soybean production, where the total production (total weight of grains) was not as dependent on agriculture expansion as before [@Macedo2012].
+However, the observed increase of soybean establishment after 2012 shows that soybean expansion did not halt after the soy moratorium nor after periods of high soybean productivity.
+Instead of halt expansion, producers started to expand over areas cleared many years ago (70 % of the conversions took more than 10 years), which may still have indirect impacts on deforestation increase over distant areas [@Arima2011; @Gollnow2018; @Silva2021].
+Expansion of agricultural areas also occurred over cleared areas of secondary forests in 2012, but with an important amount of fast conversions (Figure @ref(fig:transbar-plot).d).
+The causes of the increase of agricultural establishment areas over primary and secondary forests can be numerous, and one main driver was the approval of a new Forest Code, in 2012, which is considered to have undermined the environmental protection of forests [@Kroger2017; @Pereira2019], since it opened the possibility of amnesty for illegally deforested areas [@Schielein2018; @Santanna2021; @Filho2014], and the possible reduction of legal reserve area [@Freitas2018].
 
@@
 
-```{r transridge-plot, out.width = "17cm", fig.pos = "h", echo = FALSE, fig.cap = "Transition length patterns inside states in the Amazon biome.
+```{r transridge-plot, out.width = "17cm", fig.pos = "h", echo = FALSE, fig.cap = "Conversion length patterns inside states in the Amazon biome.
 Each year has a density estimate of the conversion lengths, represented as colored curves.
@@
 The dispersion of observed and estimated values shows no clear pattern of errors (Figure \@ref(fig:errorscatter-plot)).
-Transition length values are more concentrated in smaller values, which also shows higher errors (Figure \@ref(fig:errorscatter-plot).c).
+Conversion length values are more concentrated in smaller values, which also shows higher errors (Figure \@ref(fig:errorscatter-plot).c).
 However, this is expected since faster conversions are more common (Figure \@ref(fig:map-plot)).
 
-```{r errorscatter-plot, out.width = "17cm", fig.pos = "h", echo = FALSE, fig.cap = "Scatter plot between paired values of estimated and observed Deforestation and Agriculture Establishment years, and their respective Transition Length.
+```{r errorscatter-plot, out.width = "17cm", fig.pos = "h", echo = FALSE, fig.cap = "Scatter plot between paired values of estimated and observed Deforestation and Agriculture Establishment years, and their respective Conversion Length.
 The gray line represents values where there would be no errors of the estimated values in comparison with the observed values.
@@
 Other inclusion errors were detected, where areas that clearly showed no agriculture were considered as conversions (sample 10, 98).
-There were also omissions errors, especially agriculture plots not being considered as a conversion as a whole, where parts of a plot that were clearly a conversion to agriculture were not considered as one (sample 16, 30, 75).
+There were also omission errors, especially agriculture plots not being considered as a conversion as a whole, where parts of a plot that were clearly a conversion to agriculture were not considered as one (sample 16, 30, 75).
 A lot of uncertainty was identified at borders of agricultural areas (sample 25, 56, 84, 92), which is a very common issue in classification of satellite imagery, since there are the presence of mixed pixels, where different objects share the space of a same pixel [@Kaur2019].
@@
 
-We also observed that in many samples, the patterns of agricultural areas were roughly well represented (sample 25, 43, 56, 84), the shape of agricultural plots were clear and it was possible to distinguish different agricultural areas.
+We also observed that in many samples, the agricultural areas were roughly well represented (sample 25, 43, 56, 84), the shape of agricultural plots were clear and it was possible to distinguish different agricultural areas.
 Some locations present a higher homogeneity of conversions inside agriculture plots (sample 84), which may be related to areas with smaller uncertainties.

### 2022-12-22 manuscript/manuscript.tex
Update methods. Fix typo.
Update results. Fix confuse sentences. Add some information. Add references.

@@
 They also contain the unique id of each pixel and tile (same as the table above).
-The ``agri\_code'' column holds information about the type of agriculture the was established after deforestation.
+The ``agri\_code'' column holds information about the type of agriculture that was established after deforestation.
 The ``forest\_year'' have values that represent the year of deforestation, the ``agri\_year'' is the year when the agriculture establishment happened, and the subtraction between both (the conversion length) is stored in the ``c\_length'' column.
@@
 The conversion calculations show that in the Brazilian Amazon biome, 64874 square kilometers of forests were converted to agriculture between 1985 and 2021.
-The length of the conversions can go from 0 to 35 years, in which conversions closer to 0 years are considered as fast conversions, and conversions closer to 35 years are considered as slow conversions.
-Transitions of 0 years are considered ``direct'' conversions, where there was no presence of pasture before the establishment of agriculture.
+The length of the conversions can go from 0 to 35 years (time span), in which conversions closer to 0 years are considered as fast conversions, and conversions closer to 35 years are considered as slow conversions.
+Conversions of 0 years are considered ``direct'' conversions, where there was no presence of pasture before the establishment of agriculture.
 Our estimations show that around 9.2\% of the conversions were ``direct''.
 
-Our conversions calculations found pixels that presented up to six conversions from forest to agriculture since 1985, as these numbers are unlikely to happen (they are distributed as sparse pixels, and do not show patterns of real conversions, being common at borders), we proceeded the results analysis only in the first conversions found in each of these pixels.
+Our results found pixels that presented up to six conversions (i.e., six cycles) from forest to agriculture since 1985.
+However, since these cycles are unlikely to happen (they are distributed as sparse pixels, and do not show patterns of real conversions, being common at borders), we proceeded the analysis considering only the first conversions found in each of these pixels (these pixels with multiple conversions represented 0.5\% of the total converted areas in our study).
 
@@
 
-\subsection{Transition patterns}
+\subsection{Conversion patterns}
 
-Transitions from forest to agriculture can be found in almost every region in the Amazon, but is mostly concentrated in clusters, specially in the south and east of the biome, in the states of Mato Grosso, Pará and Maranhão (Figure \ref{fig:map-plot}).
+Cnversions from forest to agriculture can be found in almost every region in the Amazon, but is mostly concentrated in clusters, specially in the south and east of the biome, in the states of Mato Grosso, Pará and Maranhão (Figure \ref{fig:map-plot}).
 
@@
 \begin{figure}[h]
-\includegraphics[width=17cm]{figs/trans_length_cols} \caption{Transition area from forest to agriculture, per year and conversion length.
+\includegraphics[width=17cm]{figs/trans_length_cols} \caption{Conversion area from forest to agriculture, per year and conversion length.
 The bars represent the total amount of area at some state of the conversion for each year.
@@
 Blue tones represents fast conversions (closer to 0 years), pink tones represent slow conversions (closer to 35 years).
-Transition events were separated by deforestation of primary forests (a) and secondary forests (c), and the subsequent agriculture establishment of primary forests (b) and secondary forests (d).}\label{fig:transbar-plot}
+Conversion events were separated by deforestation of primary forests (a) and secondary forests (c), and the subsequent agriculture establishment of primary forests (b) and secondary forests (d).}\label{fig:transbar-plot}
 \end{figure}
@@
 The deforestation area of primary forests increased largely from 1986 to 2003, which was followed by a steep decrease until 2009, when the deforested areas reached a stable rate (Figure \ref{fig:transbar-plot}.a).
-From 1985 to 1995, most deforested areas suffered a slow conversion, mostly were higher than 10 years (76 \%).
-After this period, fast conversions started to become more common (Figure \ref{fig:transbar-plot}.a), specially from 2002 to 2004, where 70 \% of the transitions were faster than 10 years.
-Our results show that in 2003 (the peak of deforestation), 12.7\% of the conversions were a direct conversion from forest to agriculture.
+From 1985 to 1995, most deforested areas had a slow conversion to agriculture, mostly higher than 10 years (76 \%).
+After this period (1985-\/-1995), fast conversions started to become more common (Figure \ref{fig:transbar-plot}.a), specially from 2002 to 2004, where 70 \% of the conversions were faster than 10 years.
+Our results show that in 2003 (the peak of deforestation), 12.7 \% of the conversions were a direct conversion from forest to agriculture.
 Previous studies estimated a proportion of 23 \% in the Mato Grosso state, for the same year \citep{Morton2006}.
-If we include fast conversions (from 0 to 2 years) from our results, the proportion jumps to 48.1 \% of all conversions in 2003.
+If we include fast conversions (from 0 to 2 years), the proportion jumps to 48.1 \% of all conversions in 2003.
 
 Agricultural establishment over areas of primary forests peaked in 2005 and 2013 (Figure \ref{fig:transbar-plot}.b).
-Despite similar agricultural establishment areas in both years, their conversion lengths differ greatly (Figure \ref{fig:transbar-plot}.b), in 2005 most of the conversions were faster than 10 years (78 \%), while in 2013 the majority of conversions were slower than 10 years (66 \%).
+Despite similar agricultural establishment areas in both years, their conversion lengths differ greatly (Figure \ref{fig:transbar-plot}.b).
+In 2005 most of the conversions were faster than 10 years (78 \%), while in 2013 the majority of conversions took more than 10 years (66 \%).
 In 2010, it was observed a shift in the conversion length from forest until the establishment of agricultural areas.
@@
 However, after 2019, a sudden drop of agriculture establishment rate happened (Figure \ref{fig:transbar-plot}.b).
-This behavior could be related to economic causes or even the scarcity of suitable areas for agricultural expansion, which seems unlikely, since the last years have been marked by an increase of deforestation rates and the advance of pastureland over the forest.
+This behavior could be related to economic causes \citep{Morton2006, Gusso2017, Hargrave2013} or even the scarcity of suitable areas for agricultural expansion, which seems unlikely, since the last years have been marked by an increase of deforestation rates and the advance of pastureland over the forest.
 
@@
 The PPCDAm was considered as a successful policy to slow deforestation rates in Brazil, with international recognition.
-Our calculations from MapBiomas data reinforces the correlation of the PPCDAm with the reduction of deforestation after 2004 (Figure \ref{fig:transbar-plot}.a), and reduction of the agriculture establishment after 2005 (Figure \ref{fig:transbar-plot}.b).
+Our calculations from MapBiomas data reinforces the correlation of the PPCDAm with the reduction of deforestation after 2004 \citep{West2021, Arima2014} (Figure \ref{fig:transbar-plot}.a), and reduction of the agriculture establishment after 2005 (Figure \ref{fig:transbar-plot}.b).
 
@@
 The soy moratorium is considered to have contributed to the decline of the deforestation directly related to soybean expansion \citep{Paim2021, Amaral2021, Heilmayr2020, Kastens2017}.
-Our estimates of conversions show a decrease of deforested areas after 2008, where it reached minimal values (Figure \ref{fig:transbar-plot}.a).
-After 2006, agriculture establishment over deforested primary forests suffered a decrease, which stayed relatively stable until 2012, where a steep increase occurred, however, the new areas being occupied by agriculture were mainly over areas that were cleared before 2008 (Figure \ref{fig:transbar-plot}.b).
-The period between 2006 and 2010 is also considered as a turnover point, known as the decoupling of deforestation and soybean production, where the total production was not as dependent on agriculture expansion as before \citep{Macedo2012}.
-Therefore, the observed increase of agricultural establishment after 2012 shows that agriculture expansion did not halt after the soy moratorium nor after periods of high soybean productivity.
-Producers started to expand over areas cleared many years ago, which may still have indirect impacts on deforestation increase over distant areas \citep{Arima2011, Gollnow2018}.
-However, this indirect effect is not always clearly observed when analyzing the Amazon biome as a whole, our data shows that periods of fast agriculture expansion occurred at the same time with minimal deforestation values, according to PRODES estimates \citep{Assis2019}.
-Expansion of agricultural areas also occurred over cleared areas of secondary forests in 2012, but with an important amount of fast conversions (Figure \ref{fig:transbar-plot}.d).
-The causes of the increase of agricultural establishment areas can be numerous, and one main driver was the approval of a new Forest Code, in 2012, which is considered to have undermined the environmental protection of forests \citep{Kroger2017, Pereira2019}, since it opened the possibility of amnesty for illegally deforested areas \citep{Schielein2018, Santanna2021, Filho2014}, and the possible reduction of legal reserve area \citep{Freitas2018}.
+Our estimates of conversions show a decrease of deforested areas after 2008, where it reached minimal values (Figure \citet{ref}(fig:transbar-plot).a).
+After 2006, agriculture establishment over deforested primary forests suffered a decrease, the low rates of agriculture establishment stayed relatively stable until 2012, where a steep increase occurred.
+However, the new primary forest areas being occupied by agriculture were mainly over areas that were cleared before 2008 (Figure \citet{ref}(fig:transbar-plot).b), which is the last year of deforestation tolerated by the soy moratorium.
+The period between 2006 and 2010 is also considered as a turnover point, known as the decoupling of deforestation and soybean production, where the total production (total weight of grains) was not as dependent on agriculture expansion as before \citep{Macedo2012}.
+However, the observed increase of soybean establishment after 2012 shows that soybean expansion did not halt after the soy moratorium nor after periods of high soybean productivity.
+Instead of halt expansion, producers started to expand over areas cleared many years ago (70 \% of the conversions took more than 10 years), which may still have indirect impacts on deforestation increase over distant areas \citep{Arima2011, Gollnow2018, Silva2021}.
+Expansion of agricultural areas also occurred over cleared areas of secondary forests in 2012, but with an important amount of fast conversions (Figure \citet{ref}(fig:transbar-plot).d).
+The causes of the increase of agricultural establishment areas over primary and secondary forests can be numerous, and one main driver was the approval of a new Forest Code, in 2012, which is considered to have undermined the environmental protection of forests \citep{Kroger2017, Pereira2019}, since it opened the possibility of amnesty for illegally deforested areas \citep{Schielein2018, Santanna2021, Filho2014}, and the possible reduction of legal reserve area \citep{Freitas2018}.
 
@@
 \begin{figure}[h]
-\includegraphics[width=17cm]{figs/trans_ridge} \caption{Transition length patterns inside states in the Amazon biome.
+\includegraphics[width=17cm]{figs/trans_ridge} \caption{Conversion length patterns inside states in the Amazon biome.
 Each year has a density estimate of the conversion lengths, represented as colored curves.
@@
 The dispersion of observed and estimated values shows no clear pattern of errors (Figure \ref{fig:errorscatter-plot}).
-Transition length values are more concentrated in smaller values, which also shows higher errors (Figure \ref{fig:errorscatter-plot}.c).
+Conversion length values are more concentrated in smaller values, which also shows higher errors (Figure \ref{fig:errorscatter-plot}.c).
 However, this is expected since faster conversions are more common (Figure \ref{fig:map-plot}).
@@
 \begin{figure}[h]
-\includegraphics[width=17cm]{figs/error_scatter} \caption{Scatter plot between paired values of estimated and observed Deforestation and Agriculture Establishment years, and their respective Transition Length.
+\includegraphics[width=17cm]{figs/error_scatter} \caption{Scatter plot between paired values of estimated and observed Deforestation and Agriculture Establishment years, and their respective Conversion Length.
 The gray line represents values where there would be no errors of the estimated values in comparison with the observed values.
@@
 Other inclusion errors were detected, where areas that clearly showed no agriculture were considered as conversions (sample 10, 98).
-There were also omissions errors, especially agriculture plots not being considered as a conversion as a whole, where parts of a plot that were clearly a conversion to agriculture were not considered as one (sample 16, 30, 75).
+There were also omission errors, especially agriculture plots not being considered as a conversion as a whole, where parts of a plot that were clearly a conversion to agriculture were not considered as one (sample 16, 30, 75).
 A lot of uncertainty was identified at borders of agricultural areas (sample 25, 56, 84, 92), which is a very common issue in classification of satellite imagery, since there are the presence of mixed pixels, where different objects share the space of a same pixel \citep{Kaur2019}.

### 2022-12-22 manuscript/manuscript.Rmd
Small fixes.

@@
 
-Cnversions from forest to agriculture can be found in almost every region in the Amazon, but is mostly concentrated in clusters, specially in the south and east of the biome, in the states of Mato Grosso, Pará and Maranhão (Figure \@ref(fig:map-plot)).
+Conversions from forest to agriculture can be found in almost every region in the Amazon, but is mostly concentrated in clusters, specially in the south and east of the biome, in the states of Mato Grosso, Pará and Maranhão (Figure \@ref(fig:map-plot)).
 
@@
 The soy moratorium is considered to have contributed to the decline of the deforestation directly related to soybean expansion [@Paim2021; @Amaral2021; @Heilmayr2020; @Kastens2017].
-Our estimates of conversions show a decrease of deforested areas after 2008, where it reached minimal values (Figure @ref(fig:transbar-plot).a).
+Our estimates of conversions show a decrease of deforested areas after 2008, where it reached minimal values (Figure \@ref(fig:transbar-plot).a).
 After 2006, agriculture establishment over deforested primary forests suffered a decrease, the low rates of agriculture establishment stayed relatively stable until 2012, where a steep increase occurred.
-However, the new primary forest areas being occupied by agriculture were mainly over areas that were cleared before 2008 (Figure @ref(fig:transbar-plot).b), which is the last year of deforestation tolerated by the soy moratorium.
+However, the new primary forest areas being occupied by agriculture were mainly over areas that were cleared before 2008 (Figure \@ref(fig:transbar-plot).b), which is the last year of deforestation tolerated by the soy moratorium.
 The period between 2006 and 2010 is also considered as a turnover point, known as the decoupling of deforestation and soybean production, where the total production (total weight of grains) was not as dependent on agriculture expansion as before [@Macedo2012].
@@
 Instead of halt expansion, producers started to expand over areas cleared many years ago (70 % of the conversions took more than 10 years), which may still have indirect impacts on deforestation increase over distant areas [@Arima2011; @Gollnow2018; @Silva2021].
-Expansion of agricultural areas also occurred over cleared areas of secondary forests in 2012, but with an important amount of fast conversions (Figure @ref(fig:transbar-plot).d).
+Expansion of agricultural areas also occurred over cleared areas of secondary forests in 2012, but with an important amount of fast conversions (Figure \@ref(fig:transbar-plot).d).
 The causes of the increase of agricultural establishment areas over primary and secondary forests can be numerous, and one main driver was the approval of a new Forest Code, in 2012, which is considered to have undermined the environmental protection of forests [@Kroger2017; @Pereira2019], since it opened the possibility of amnesty for illegally deforested areas [@Schielein2018; @Santanna2021; @Filho2014], and the possible reduction of legal reserve area [@Freitas2018].

### 2022-12-22 manuscript/manuscript.tex
Small fixes.

@@
 
-Cnversions from forest to agriculture can be found in almost every region in the Amazon, but is mostly concentrated in clusters, specially in the south and east of the biome, in the states of Mato Grosso, Pará and Maranhão (Figure \ref{fig:map-plot}).
+Conversions from forest to agriculture can be found in almost every region in the Amazon, but is mostly concentrated in clusters, specially in the south and east of the biome, in the states of Mato Grosso, Pará and Maranhão (Figure \ref{fig:map-plot}).
 
@@
 The soy moratorium is considered to have contributed to the decline of the deforestation directly related to soybean expansion \citep{Paim2021, Amaral2021, Heilmayr2020, Kastens2017}.
-Our estimates of conversions show a decrease of deforested areas after 2008, where it reached minimal values (Figure \citet{ref}(fig:transbar-plot).a).
+Our estimates of conversions show a decrease of deforested areas after 2008, where it reached minimal values (Figure \ref{fig:transbar-plot}.a).
 After 2006, agriculture establishment over deforested primary forests suffered a decrease, the low rates of agriculture establishment stayed relatively stable until 2012, where a steep increase occurred.
-However, the new primary forest areas being occupied by agriculture were mainly over areas that were cleared before 2008 (Figure \citet{ref}(fig:transbar-plot).b), which is the last year of deforestation tolerated by the soy moratorium.
+However, the new primary forest areas being occupied by agriculture were mainly over areas that were cleared before 2008 (Figure \ref{fig:transbar-plot}.b), which is the last year of deforestation tolerated by the soy moratorium.
 The period between 2006 and 2010 is also considered as a turnover point, known as the decoupling of deforestation and soybean production, where the total production (total weight of grains) was not as dependent on agriculture expansion as before \citep{Macedo2012}.
@@
 Instead of halt expansion, producers started to expand over areas cleared many years ago (70 \% of the conversions took more than 10 years), which may still have indirect impacts on deforestation increase over distant areas \citep{Arima2011, Gollnow2018, Silva2021}.
-Expansion of agricultural areas also occurred over cleared areas of secondary forests in 2012, but with an important amount of fast conversions (Figure \citet{ref}(fig:transbar-plot).d).
+Expansion of agricultural areas also occurred over cleared areas of secondary forests in 2012, but with an important amount of fast conversions (Figure \ref{fig:transbar-plot}.d).
 The causes of the increase of agricultural establishment areas over primary and secondary forests can be numerous, and one main driver was the approval of a new Forest Code, in 2012, which is considered to have undermined the environmental protection of forests \citep{Kroger2017, Pereira2019}, since it opened the possibility of amnesty for illegally deforested areas \citep{Schielein2018, Santanna2021, Filho2014}, and the possible reduction of legal reserve area \citep{Freitas2018}.

### 2022-12-23 scripts/figures/fig_1.R
Rename figures scripts.

@@ -1,227 +1,0 @@
-# HEADER ----------------------------------------------------------------------
-#
-# Title:        Create figure 1
-# Description:  The objective of this routine is to explore the dataset by
-#               filtering and aggregating data in Spark. Results are supposed
-#               to give us a overall view of the data by plots and tables.
-#
-# Author:       Hugo Tameirao Seixas
-# Contact:      tameirao.hugo@gmail.com
-# Date:         2020-11-10
-#
-# Notes:        In order to run this routine, you will need to successfully
-#               install and connect to Spark using the {sparklyr} package.
-#               Installation and use guide can be found in:
-#               https://github.com/sparklyr/sparklyr
-#
-# LIBRARIES -------------------------------------------------------------------
-#
-library(sf)
-library(geobr)
-library(terra)
-library(magrittr)
-library(sparklyr)
-library(dplyr)
-library(tibble)
-library(readr)
-library(tidyr)
-library(forcats)
-library(ggplot2)
-library(ggfx)
-library(patchwork)
-library(scico)
-#
-# OPTIONS ---------------------------------------------------------------------
-#
-source('conf/spark_config.R')
-#
-# LOAD AUXILIARY DATA ---------------------------------------------------------
-
-amazon <- read_biomes() %>%
-  filter(code_biome == 1) %>%
-  st_transform("EPSG:4326")
-
-states <- read_state() %>%
-  st_transform("EPSG:4326") %>%
-  st_intersection(amazon)
-
-# LOAD DATASETS ---------------------------------------------------------------
-
-## Load transition length raster ----
-mosaic <-
-  rast("data/trans_raster_mosaic/mb_mosaic_cycle_1.tif")[["trans_length"]]
-
-mosaic_agg <- terra::aggregate(mosaic, fact = 10, fun = "modal")
-
-## Create hex grid ----
-grid <- vect(st_make_grid(mosaic_agg, cellsize = 0.1, square = FALSE))
-
-## Connect with Spark -----
-sc <- spark_connect(master = "local", config = config, version = "3.3.0")
-
-## Load trans_length table ----
-trans_length <-
-  spark_read_parquet(
[…]

### 2022-12-23 scripts/figures/fig_2.R

@@ -1,230 +1,0 @@
-# HEADER ----------------------------------------------------------------------
-#
-# Title:        Create figure 2
-# Description:  The objective of this routine is to explore the dataset by
-#               filtering and aggregating data in Spark. Results are supposed
-#               to give us a overall view of the data by plots and tables.
-#
-# Author:       Hugo Tameirao Seixas
-# Contact:      tameirao.hugo@gmail.com
-# Date:         2020-11-10
-#
-# Notes:        In order to run this routine, you will need to successfully
-#               install and connect to Spark using the {sparklyr} package.
-#               Installation and use guide can be found in:
-#               https://github.com/sparklyr/sparklyr
-#
-# LIBRARIES -------------------------------------------------------------------
-#
-library(sparklyr)
-library(magrittr)
-library(lubridate)
-library(scales)
-library(dplyr)
-library(tibble)
-library(readr)
-library(tidyr)
-library(forcats)
-library(ggplot2)
-library(ggtext)
-library(scico)
-#
-# OPTIONS ---------------------------------------------------------------------
-#
-source('conf/spark_config.R')
-#
-# SET LULC NAMES --------------------------------------------------------------
-
-## Get the names for each LULC class from MapBiomas ----
-mb_dict <- read_csv("data/mb_class_dictionary.csv") %>%
-  select(class_code, class_name) %>%
-  rename(agri_code = class_code)
-
-# LOAD DATASETS ---------------------------------------------------------------
-
-## Connect with Spark -----
-sc <- spark_connect(master = "local", config = config, version = "3.3.0")
-
-## Load mask_cells table ----
-mask_cells <-
-  spark_read_parquet(
-    sc,
-    name = "mask_cells",
-    path = "data/trans_tabular_dataset/mask_cells/",
-    memory = FALSE
-  )
-
-## Load trans_length table ----
-trans_length <-
-  spark_read_parquet(
-    sc,
-    name = "trans_length",
-    path = "data/trans_tabular_dataset/trans_length/",
-    memory = FALSE
-  )
-
-# FILTER AND AGGREGATE DATA ---------------------------------------------------
-
[…]

### 2022-12-23 scripts/figures/fig_4.R

@@ -1,188 +1,227 @@
 # HEADER ----------------------------------------------------------------------
 #
-# Title:        Validate transitions against visual inspection
-# Description:  The objective of this routine is compare the transition length
-#               results, comparing them with visual observations.
-#
+# Title:        Create figure 4
+# Description:  The objective of this routine is to explore the dataset by
+#               filtering and aggregating data in Spark. Results are supposed
+#               to give us a overall view of the data by plots and tables.
 #
 # Author:       Hugo Tameirao Seixas
 # Contact:      tameirao.hugo@gmail.com
-# Date:         2022-08-03
+# Date:         2020-11-10
 #
-# Notes:
-#
-#
-#
+# Notes:        In order to run this routine, you will need to successfully
+#               install and connect to Spark using the {sparklyr} package.
+#               Installation and use guide can be found in:
+#               https://github.com/sparklyr/sparklyr
 #
 # LIBRARIES -------------------------------------------------------------------
 #
-library(arrow)
+library(sf)
+library(geobr)
+library(terra)
 library(magrittr)
-library(lubridate)
-library(glue)
+library(sparklyr)
 library(dplyr)
-library(units)
 library(tibble)
 library(readr)
 library(tidyr)
 library(forcats)
-library(purrr)
-library(scales)
 library(ggplot2)
-library(ggtext)
+library(ggfx)
+library(patchwork)
+library(scico)
 #
 # OPTIONS ---------------------------------------------------------------------
 #
+source('conf/spark_config.R')
 #
-# LOAD DATA -------------------------------------------------------------------
+# LOAD AUXILIARY DATA ---------------------------------------------------------
 
-## Load observed data ----
-observed <- read_csv("data/validation/results.csv") %>%
-  mutate( # Remove points where there were no transition
-    forest_year = if_else(forest_year <= 0, NA_real_, forest_year),
-    agri_year = if_else(agri_year <= 0, NA_real_, agri_year),
[…]

### 2022-12-23 scripts/figures/fig_5.R

@@ -1,186 +1,207 @@
 # HEADER ----------------------------------------------------------------------
 #
-# Title:        Validate transitions against visual inspection
-# Description:  The objective of this routine is compare the transition length
-#               results, comparing them with visual observations.
-#
+# Title:        Create figure 5
+# Description:  The objective of this routine is to explore the dataset by
+#               filtering and aggregating data in Spark. Results are supposed
+#               to give us a overall view of the data by plots and tables.
 #
 # Author:       Hugo Tameirao Seixas
 # Contact:      tameirao.hugo@gmail.com
-# Date:         2022-08-03
+# Date:         2020-11-10
 #
-# Notes:
-#
-#
-#
+# Notes:        In order to run this routine, you will need to successfully
+#               install and connect to Spark using the {sparklyr} package.
+#               Installation and use guide can be found in:
+#               https://github.com/sparklyr/sparklyr
 #
 # LIBRARIES -------------------------------------------------------------------
 #
-library(arrow)
+library(sparklyr)
 library(magrittr)
 library(lubridate)
-library(glue)
+library(scales)
 library(dplyr)
-library(units)
 library(tibble)
 library(readr)
 library(tidyr)
 library(forcats)
-library(purrr)
-library(scales)
 library(ggplot2)
 library(ggtext)
+library(scico)
 #
 # OPTIONS ---------------------------------------------------------------------
 #
+source('conf/spark_config.R')
 #
-# LOAD DATA -------------------------------------------------------------------
+# SET LULC NAMES --------------------------------------------------------------
 
-## Load observed data ----
-observed <- read_csv("data/validation/results.csv") %>%
-  mutate( # Remove points where there were no transition
-    forest_year = if_else(forest_year <= 0, NA_real_, forest_year),
-    agri_year = if_else(agri_year <= 0, NA_real_, agri_year),
-    trans_length = if_else(agri_year <= 0, NA_real_, trans_length),
-    trans_length = if_else(forest_year <= 0, NA_real_, trans_length)
[…]

### 2022-12-23 scripts/figures/fig_6.R

@@ -1,6 +1,6 @@
 # HEADER ----------------------------------------------------------------------
 #
-# Title:        Create figure 3
+# Title:        Create figure 6
 # Description:  The objective of this routine is to explore the dataset by
 #               filtering and aggregating data in Spark. Results are supposed
 #               to give us a overall view of the data by plots and tables.

### 2022-12-23 scripts/figures/fig_7.R

@@ -1,0 +1,188 @@
+# HEADER ----------------------------------------------------------------------
+#
+# Title:        Create figure 7
+# Description:  The objective of this routine is compare the transition length
+#               results, comparing them with visual observations.
+#
+#
+# Author:       Hugo Tameirao Seixas
+# Contact:      tameirao.hugo@gmail.com
+# Date:         2022-08-03
+#
+# Notes:
+#
+#
+#
+#
+# LIBRARIES -------------------------------------------------------------------
+#
+library(arrow)
+library(magrittr)
+library(lubridate)
+library(glue)
+library(dplyr)
+library(units)
+library(tibble)
+library(readr)
+library(tidyr)
+library(forcats)
+library(purrr)
+library(scales)
+library(ggplot2)
+library(ggtext)
+#
+# OPTIONS ---------------------------------------------------------------------
+#
+#
+# LOAD DATA -------------------------------------------------------------------
+
+## Load observed data ----
+observed <- read_csv("data/validation/results.csv") %>%
+  mutate( # Remove points where there were no transition
+    forest_year = if_else(forest_year <= 0, NA_real_, forest_year),
+    agri_year = if_else(agri_year <= 0, NA_real_, agri_year),
+    trans_length = if_else(agri_year <= 0, NA_real_, trans_length),
+    trans_length = if_else(forest_year <= 0, NA_real_, trans_length)
+  ) %>%
+  mutate( # Fix dates to match the estimates
+    forest_year = forest_year,
+    agri_year = agri_year
+  ) %>%
+  pivot_longer(
+    cols =  c(forest_year, agri_year, trans_length),
+    names_to = "var",
+    values_to = "observed"
+  ) %>%
+  arrange(cell_id)
+
+## Load estimates data ----
+estimated <- open_dataset("data/trans_tabular_dataset/trans_length/")
+
+estimated <- estimated %>%
+  filter(
+    cell_id %in% observed$cell_id
+  ) %>%
+  collect() %>%
+  pivot_longer(
+    cols =  c(forest_year, agri_year, trans_length),
+    names_to = "var",
+    values_to = "estimated"
+  ) %>%
+  arrange(cell_id)
+
+# Check if there is any difference between tables
[…]

### 2022-12-23 scripts/figures/fig_8.R

@@ -1,0 +1,207 @@
+# HEADER ----------------------------------------------------------------------
+#
+# Title:        Create figure 8
+# Description:  The objective of this routine is compare the transition length
+#               results, comparing them with visual observations.
+#
+#
+# Author:       Hugo Tameirao Seixas
+# Contact:      tameirao.hugo@gmail.com
+# Date:         2022-08-03
+#
+# Notes:
+#
+#
+#
+#
+# LIBRARIES -------------------------------------------------------------------
+#
+library(arrow)
+library(magrittr)
+library(lubridate)
+library(glue)
+library(dplyr)
+library(units)
+library(tibble)
+library(readr)
+library(tidyr)
+library(forcats)
+library(purrr)
+library(scales)
+library(ggplot2)
+library(ggtext)
+#
+# OPTIONS ---------------------------------------------------------------------
+#
+#
+# LOAD DATA -------------------------------------------------------------------
+
+## Load observed data ----
+observed <- read_csv("data/validation/results.csv") %>%
+  mutate( # Remove points where there were no transition
+    forest_year = if_else(forest_year <= 0, NA_real_, forest_year),
+    agri_year = if_else(agri_year <= 0, NA_real_, agri_year),
+    trans_length = if_else(agri_year <= 0, NA_real_, trans_length),
+    trans_length = if_else(forest_year <= 0, NA_real_, trans_length)
+  ) %>%
+  mutate(across(forest_year:agri_year, ~ .x - 1984)) %>%
+  pivot_longer(
+    cols =  c(forest_year, agri_year, trans_length),
+    names_to = "var",
+    values_to = "observed"
+  ) %>%
+  arrange(cell_id)
+
+## Load estimates data ----
+estimated <- open_dataset("data/trans_tabular_dataset/trans_length/")
+
+estimated %<>%
+  filter(
+    cell_id %in% observed$cell_id
+  ) %>%
+  collect() %>%
+  mutate(across(forest_year:agri_year, ~ .x - 1984)) %>%
+  pivot_longer(
+    cols =  c(forest_year, agri_year, trans_length),
+    names_to = "var",
+    values_to = "estimated"
+  ) %>%
+  arrange(cell_id)
+
+# Check if there is any difference between tables
[…]

### 2022-12-23 conf/config.R
Fix variables names.

@@ -1,7 +1,8 @@
 # CONFIG ----------------------------------------------------------------------
 #
 # This is a configuration file that sets options and variables that will be
-# used in the "01_download_mapbiomas.R" and "02_extract_transition.R".
+# used in the "01_download_mapbiomas.R", "02_extract_mask" and
+# "03_extract_conversion.R".
 # The options "gee_email", "clear_driver_folder" and "view_map" are safe
 # to be changed ("gee_email" is actually essential to run the download script).
 # However, it is not guaranteed that changes in the other options will generate
@@ -52,13 +53,13 @@ tile_dim <- 1536L
 # Default:
 #   tile_dim <- 1536L
 
-# The natural class of the transition -----------------------------------------
+# The natural class of the conversion -----------------------------------------
 trans_nat <- c(3)
 # Default value is "Forest Formation":
 #   trans_nat <- c(3)
 # Check https://mapbiomas.org/en/codigos-de-legenda?cama_set_language=en
 
-# The non natural class of the transition -------------------------------------
+# The non natural class of the conversion -------------------------------------
 trans_ant <- c(18, 19, 39, 20, 40, 62, 41, 36, 46, 47, 48)
 # Default value are agriculture classes:
 #   trans_ant <- c(18, 19, 39, 20, 40, 62, 41, 36, 46, 47, 48)

### 2022-12-23 scripts/01_download.R
Update variable names.

@@ -8,7 +8,7 @@
 #               (https://mapbiomas.org/en?cama_set_language=en)
 #
 # Author:       Hugo Tameirao Seixas
-# Contact:      tameirao.hugo@gmail.com
+# Contact:      seixas.hugo@protonmail.com
 # Date:         2020-09-11
 #
 # Notes:        In order to run this routine, you will need to have access
@@ -164,7 +164,7 @@ time_code <- as.integer(Sys.time())
 
 # Parent folder
 drive_mkdir(
-  name = glue("mb_transition-{time_code}"),
+  name = glue("mb_conversion-{time_code}"),
   path = "~/",
   overwrite = FALSE
 )
@@ -176,7 +176,7 @@ walk(
 
     drive_mkdir(
       name = glue("{folder_name}-{time_code}"),
-      path = glue("~/mb_transition-{time_code}/"),
+      path = glue("~/mb_conversion-{time_code}/"),
       overwrite = FALSE
     )
 
@@ -232,7 +232,7 @@ ee_monitoring(download_mb, task_time = 600)
 drive_files <-
   map_dfr(
     1:10,
-    ~ drive_ls(glue("~/mb_transition-{time_code}/"), recursive = TRUE)
+    ~ drive_ls(glue("~/mb_conversion-{time_code}/"), recursive = TRUE)
   ) %>%
   distinct(name, .keep_all = TRUE) %>%
   filter(

### 2022-12-23 scripts/02_extract_mask.R

@@ -5,7 +5,7 @@
 #               are saved in .parquet format.
 #
 # Author:       Hugo Tameirao Seixas
-# Contact:      tameirao.hugo@gmail.com
+# Contact:      seixas.hugo@protonmail.com
 # Date:         2021-26-02
 #
 # Notes:        Do not change the names of files and directories of the
@@ -168,7 +168,7 @@ walk(
       ## Create directories ----
       dir_create(
         path = glue(
-          "data/trans_tabular_dataset/mask_cells/",
+          "data/c_tabular_dataset/mask_cells/",
           "tile_id={tile}/"
         )
       )
@@ -177,7 +177,7 @@ walk(
       write_parquet(
         x = table,
         sink = glue(
-          "data/trans_tabular_dataset/mask_cells/",
+          "data/c_tabular_dataset/mask_cells/",
           "tile_id={tile}/",
           "mask_cells-{file_list$tile_code[tile]}.parquet"
         ),
@@ -192,7 +192,7 @@ walk(
 ## Save raster metadata file ----
 write_parquet(
   x = tiles_metadata,
-  sink = "data/trans_tabular_dataset/tiles_metadata.parquet",
+  sink = "data/c_tabular_dataset/tiles_metadata.parquet",
   version = "2.6"
 )
 

### 2022-12-23 scripts/03_extract_conversion.R

@@ -1,10 +1,10 @@
 # HEADER ----------------------------------------------------------------------
 #
-# Title:        Extract transition from Forest to Agriculture
-# Description:  This code extracts values of the transition between Forest and
+# Title:        Extract conversion from Forest to Agriculture
+# Description:  This code extracts values of the conversion between Forest and
 #               Agriculture from MapBiomas raster files. The process is applied
 #               in each file, where a number of iterations are performed to
-#               extract values from different transition cycles. The results
+#               extract values from different conversion cycles. The results
 #               are saved in .parquet format.
 #
 # Author:       Hugo Tameirao Seixas
@@ -39,7 +39,7 @@ source("R/terra_as_tibble.R")
 mb_dict <- read_delim('data/mb_class_dictionary.csv', delim = ',')
 
 ## Load mask cells from parquet dataset  ----
-mask_ds <- open_dataset("data/trans_tabular_dataset/mask_cells/")
+mask_ds <- open_dataset("data/c_tabular_dataset/mask_cells/")
 
 # LIST RASTER FILES -----------------------------------------------------------
 
@@ -57,9 +57,9 @@ file_list <-
   relocate(tile_id, tile_code) %>%
   arrange(tile_code)
 
-# EXTRACT LULC TRANSITION VALUES ----------------------------------------------
+# EXTRACT LULC CONVERSION VALUES ----------------------------------------------
 
-## Iterate trough files to extract transition values ----
+## Iterate trough files to extract conversion values ----
 walk(
   .x = pull(.data = file_list, var = tile_id),
   function(tile) {
@@ -107,32 +107,32 @@ walk(
 
     } else {
 
-      ### Create empty tibble to store transition values ----
-      trans_length_all <-
+      ### Create empty tibble to store conversion values ----
+      c_length_all <-
         tibble(
           cell_id = integer(),
           agri_code = integer(),
           forest_year = integer(),
           agri_year = integer(),
-          trans_length = integer(),
-          trans_cycle = integer()
+          c_length = integer(),
+          c_cycle = integer()
         )
 
       ### Create empty tibble to store lulc time series ----
-      trans_classes_all <-
+      c_classes_all <-
         tibble(
           cell_id = integer(),
           year = integer(),
           class_code = integer(),
-          trans_cycle = integer()
+          c_cycle = integer()
         )
 
-      ### Start transition cycles counting  ----
[…]

### 2022-12-23 scripts/04_rasterize.R

@@ -1,17 +1,17 @@
 # HEADER ----------------------------------------------------------------------
 #
-# Title:        Create transition raster files
-# Description:  This codes takes values from the "trans_length" dataset and
+# Title:        Create conversion raster files
+# Description:  This codes takes values from the "c_length" dataset and
 #               metadata information to recreate raster tiles with the
-#               transition values. The new raster files share the same spatial
+#               conversion values. The new raster files share the same spatial
 #               characteristics as the downloaded tiles (extent, resolution...)
 #               Each tile will be composed by a set of raster files, one file
-#               for each forest-agriculture transition cycle, and the bands of
+#               for each forest-agriculture conversion cycle, and the bands of
 #               the new files carry values of the variables stored in the
 #               parquet dataset.
 #
 # Author:       Hugo Tameirao Seixas
-# Contact:      tameirao.hugo@gmail.com
+# Contact:      seixas.hugo@protonmail.com
 # Date:         2020-09-11
 #
 # Notes:        Do not change the names of files and directories of the
@@ -36,18 +36,18 @@ library(purrr)
 
 ## Get tiles list from metadata file ----
 tiles_metadata <-
-  read_parquet("data/trans_tabular_dataset/tiles_metadata.parquet")
+  read_parquet("data/c_tabular_dataset/tiles_metadata.parquet")
 
 # RASTERIZE TABLES ------------------------------------------------------------
 
 ## Create directory to store new raster files  ----
-dir_create("data/trans_raster_tiles/")
+dir_create("data/c_raster_tiles/")
 
-## Load transition_length values from parquet dataset  ----
-trans_ds <- open_dataset("data/trans_tabular_dataset/trans_length/")
+## Load conversion length values from parquet dataset  ----
+c_ds <- open_dataset("data/c_tabular_dataset/c_length/")
 
 ## Load mask cells from parquet dataset  ----
-mask_ds <- open_dataset("data/trans_tabular_dataset/mask_cells/")
+mask_ds <- open_dataset("data/c_tabular_dataset/mask_cells/")
 
 ## Create set of raster files for each tile ----
 walk(
@@ -70,9 +70,9 @@ walk(
       tiles_metadata %>%
       filter(tile_id == tile)
 
-    # Get transition values ----
-    trans_table <-
-      trans_ds %>%
+    # Get conversion values ----
+    c_table <-
+      c_ds %>%
[…]

### 2022-12-23 scripts/05_merge_raster.R

@@ -2,11 +2,11 @@
 #
 # Title:        Merge raster tiles
 # Description:  This script merges the raster tiles created in the
-#               "03_rasterize.R" script into a single mosaic. Each mosaic is
-#               created for each forest-agriculture transition cycle.
+#               "04_rasterize.R" script into a single mosaic. Each mosaic is
+#               created for each forest-agriculture conversion cycle.
 #
 # Author:       Hugo Tameirao Seixas
-# Contact:      tameirao.hugo@gmail.com
+# Contact:      seixas.hugo@protonmail.com
 # Date:         2020-09-11
 #
 # Notes:        Do not change the names of files and directories of the
@@ -31,12 +31,12 @@ library(stringr)
 # PREPARE FILES AND DIRECTORIES -----------------------------------------------
 
 ## Create directory to store new raster files ----
-dir_create("data/trans_raster_mosaic/")
+dir_create("data/c_raster_mosaic/")
 
-## List raster files with transition values ----
-file_list <- dir_ls("data/trans_raster_tiles/")
+## List raster files with conversion values ----
+file_list <- dir_ls("data/c_raster_tiles/")
 
-# LIST FOREST-AGRICULTURE TRANSITION CYCLES -----------------------------------
+# LIST FOREST-AGRICULTURE CONVERSION CYCLES -----------------------------------
 
 ## Get number of cycles ----
 file_list %<>%
@@ -65,7 +65,7 @@ walk(
     # Merge raster as a mosaic using GDAL ----
     gdalbuildvrt(
       gdalfile = raster_path,
-      glue("data/trans_raster_mosaic/gdal_mosaic_cycle_{cycle_num}.vrt"),
+      glue("data/c_raster_mosaic/gdal_mosaic_cycle_{cycle_num}.vrt"),
       hidenodata = FALSE,
       overwrite = TRUE
     )
@@ -73,7 +73,7 @@ walk(
     # Open virtual raster to SpatRaster object ----
     raster_mosaic <-
       rast(
-        glue("data/trans_raster_mosaic/gdal_mosaic_cycle_{cycle_num}.vrt")
+        glue("data/c_raster_mosaic/gdal_mosaic_cycle_{cycle_num}.vrt")
       )
 
     # Rename bands ----
@@ -86,7 +86,7 @@ walk(
     writeRaster(
       x = raster_mosaic,
       filename = glue(
[…]

### 2022-12-23 scripts/06_validation_data.R

@@ -39,7 +39,7 @@ source('conf/config.R')
 #
 # CREATE SAMPLES --------------------------------------------------------------
 
-mask_cells <- open_dataset("data/trans_tabular_dataset/mask_cells/")
+mask_cells <- open_dataset("data/c_tabular_dataset/mask_cells/")
 
 cells <- mask_cells %>%
   select(cell_id) %>%

### 2022-12-23 scripts/validation_app.R

@@ -33,7 +33,7 @@ results <-
     cell_id = NA_real_,
     forest_year = NA_real_,
     agri_year = NA_real_,
-    trans_length = NA_real_
+    c_length = NA_real_
   ) %>%
   drop_na()
 
@@ -95,7 +95,7 @@ ui <- grid_page(
       width = "auto"
     ),
     numericInput(
-      inputId = "transBegin",
+      inputId = "cBegin",
       label = "Forest to Pasture",
       value = 1985L,
       min = 1985L,
@@ -104,7 +104,7 @@ ui <- grid_page(
       width = "auto"
     ),
     numericInput(
-      inputId = "transEnd",
+      inputId = "cEnd",
       label = "Pasture to Agriculture",
       value = 1985L,
       min = 1985L,
@@ -119,7 +119,7 @@ ui <- grid_page(
   ),
   grid_card_text(
     area = "header",
-    content = "Forest to agriculture transition: accuracy assessment",
+    content = "Forest to agriculture conversion: accuracy assessment",
     alignment = "center",
     is_title = FALSE
   ),
@@ -315,7 +315,7 @@ server <-
         crop(
           rast(
             glue(
-              "{proj_path}/data/trans_raster_mosaic/mb_mosaic_cycle_1.tif"
+              "{proj_path}/data/c_raster_mosaic/mb_mosaic_cycle_1.tif"
             )
           )[[2]],
           bbox
@@ -360,7 +360,7 @@ server <-
         crop(
           rast(
             glue(
-              "{proj_path}/data/trans_raster_mosaic/mb_mosaic_cycle_1.tif"
+              "{proj_path}/data/c_raster_mosaic/mb_mosaic_cycle_1.tif"
             )
           )[[3]],
           bbox
@@ -405,7 +405,7 @@ server <-
         crop(
           rast(
             glue(
-              "{proj_path}/data/trans_raster_mosaic/mb_mosaic_cycle_1.tif"
+              "{proj_path}/data/c_raster_mosaic/mb_mosaic_cycle_1.tif"
             )
           )[[4]],
           bbox
@@ -419,7 +419,7 @@ server <-
       ggplot() +
         geom_raster(
           data = img_df,
-          aes(x = x, y = y, fill = trans_length)
+          aes(x = x, y = y, fill = c_length)
         ) +
         geom_sf(
           data = point,
@@ -427,7 +427,7 @@ server <-
         ) +
[…]

### 2022-12-23 scripts/figures/fig_4.R
Update email. Update variable names.

@@ -6,7 +6,7 @@
 #               to give us a overall view of the data by plots and tables.
 #
 # Author:       Hugo Tameirao Seixas
-# Contact:      tameirao.hugo@gmail.com
+# Contact:      seixas.hugo@protonmail.com
 # Date:         2020-11-10
 #
 # Notes:        In order to run this routine, you will need to successfully
@@ -47,9 +47,9 @@ states <- read_state() %>%
 
 # LOAD DATASETS ---------------------------------------------------------------
 
-## Load transition length raster ----
+## Load conversion length raster ----
 mosaic <-
-  rast("data/trans_raster_mosaic/mb_mosaic_cycle_1.tif")[["trans_length"]]
+  rast("data/c_raster_mosaic/mb_mosaic_cycle_1.tif")[["c_length"]]
 
 mosaic_agg <- terra::aggregate(mosaic, fact = 10, fun = "modal")
 
@@ -59,26 +59,26 @@ grid <- vect(st_make_grid(mosaic_agg, cellsize = 0.1, square = FALSE))
 ## Connect with Spark -----
 sc <- spark_connect(master = "local", config = config, version = "3.3.0")
 
-## Load trans_length table ----
-trans_length <-
+## Load c_length table ----
+c_length <-
   spark_read_parquet(
     sc,
-    name = "trans_length",
-    path = "data/trans_tabular_dataset/trans_length/",
+    name = "c_length",
+    path = "data/c_tabular_dataset/c_length/",
     memory = FALSE
   )
 
 # FILTER AND AGGREGATE DATA ---------------------------------------------------
 
-## Extract trans_length values from raster into hex grid ----
+## Extract c_length values from raster into hex grid ----
 extract_values <- terra::extract(mosaic_agg, grid) %>%
   drop_na() %>%
-  group_by(ID, trans_length) %>%
+  group_by(ID, c_length) %>%
   count() %>%
   group_by(ID) %>%
   filter(n == max(n))
 
-trans_hex <- grid %>%
+c_hex <- grid %>%
   st_as_sf() %>%
   mutate(ID := seq_len(nrow(.))) %>%
   left_join(., extract_values, by = "ID") %>%
@@ -110,17 +110,17 @@ hr_raster <-
   ) %>%
   as_tibble()
 
-## Get the frequency of trans_lenght ----
-frequency <- trans_length %>%
[…]

### 2022-12-23 scripts/figures/fig_5.R

@@ -6,7 +6,7 @@
 #               to give us a overall view of the data by plots and tables.
 #
 # Author:       Hugo Tameirao Seixas
-# Contact:      tameirao.hugo@gmail.com
+# Contact:      seixas.hugo@protonmail.com
 # Date:         2020-11-10
 #
 # Notes:        In order to run this routine, you will need to successfully
@@ -50,24 +50,24 @@ mask_cells <-
   spark_read_parquet(
     sc,
     name = "mask_cells",
-    path = "data/trans_tabular_dataset/mask_cells/",
+    path = "data/c_tabular_dataset/mask_cells/",
     memory = FALSE
   )
 
-## Load trans_length table ----
-trans_length <-
+## Load c_length table ----
+c_length <-
   spark_read_parquet(
     sc,
-    name = "trans_length",
-    path = "data/trans_tabular_dataset/trans_length/",
+    name = "c_length",
+    path = "data/c_tabular_dataset/c_length/",
     memory = FALSE
   )
 
 # FILTER AND AGGREGATE DATA ---------------------------------------------------
 
 ## Sum deforestation and agriculture establishment along years ----
-trans_subset <-
-  trans_length %>%
+c_subset <-
+  c_length %>%
   filter(agri_cycle == 1) %>%
   select(cell_id:forest_type) %>%
   left_join(
@@ -76,33 +76,33 @@ trans_subset <-
   ) %>%
   group_by(
     forest_type, agri_year, forest_year,
-    agri_code, trans_length
+    agri_code, c_length
   ) %>%
   summarise(total_area = sum(area, na.rm = TRUE), .groups = "drop") %>%
   collect()
 
 ### Mutate columns and convert to long format ----
-trans_subset %<>%
+c_subset %<>%
   left_join(mb_dict, by = "agri_code") %>%
   pivot_longer(
     agri_year:forest_year,
-    names_to = "transition",
+    names_to = "conversion",
     values_to = "year") %>%
   mutate(
-    transition = fct_rev(factor(transition)),
+    conversion = fct_rev(factor(conversion)),
     year = ymd(year, truncated = 2L),
     forest_type = as.character(forest_type)
   )
 
 ### Summarise data ----
-trans_subset %<>%
-  group_by(transition, forest_type, trans_length, year) %>%
+c_subset %<>%
+  group_by(conversion, forest_type, c_length, year) %>%
[…]

### 2022-12-23 scripts/figures/fig_6.R

@@ -6,7 +6,7 @@
 #               to give us a overall view of the data by plots and tables.
 #
 # Author:       Hugo Tameirao Seixas
-# Contact:      tameirao.hugo@gmail.com
+# Contact:      seixas.hugo@protonmail.com
 # Date:         2020-11-10
 #
 # Notes:        In order to run this routine, you will need to successfully
@@ -50,24 +50,24 @@ mask_cells <-
   spark_read_parquet(
     sc,
     name = "mask_cells",
-    path = "data/trans_tabular_dataset/mask_cells/",
+    path = "data/c_tabular_dataset/mask_cells/",
     memory = FALSE
   )
 
-## Load trans_length table ----
-trans_length <-
+## Load c_length table ----
+c_length <-
   spark_read_parquet(
     sc,
-    name = "trans_length",
-    path = "data/trans_tabular_dataset/trans_length/",
+    name = "c_length",
+    path = "data/c_tabular_dataset/c_length/",
     memory = FALSE
   )
 
 # FILTER AND AGGREGATE DATA ---------------------------------------------------
 
 ## Sample rows to create ridge plot ----
-trans_ridges <-
-  trans_length %>%
+c_ridges <-
+  c_length %>%
   filter(agri_cycle == 1) %>%
   select(cell_id:forest_type) %>%
   sdf_sample(fraction = 0.2, replacement = FALSE) %>%
@@ -87,12 +87,12 @@ spark_disconnect(sc)
 # CREATE PLOTS ----------------------------------------------------------------
 
 ## Ridge plot ----
-rp <- trans_ridges %>%
+rp <- c_ridges %>%
   ggplot() +
   facet_wrap( ~ name_state, ncol = 2) +
   geom_density_ridges_gradient(
     aes(
-      x = trans_length,
+      x = c_length,
       y = agri_year,
       group = agri_year,
       fill = after_stat(x)
@@ -102,11 +102,11 @@ rp <- trans_ridges %>%
   geom_text(
     data = tibble(
       label = c("(j)", "(g)", "(b)", "(i)", "(h)", "(e)", "(a)", "(d)", "(c)"),
-      name_state = unique(trans_ridges$name_state),
-      trans_length = rep(34, 9),
+      name_state = unique(c_ridges$name_state),
+      c_length = rep(34, 9),
       agri_year = rep(ymd("1995-01-01"), 9)
     ),
-    aes(x = trans_length, y = agri_year, label = label),
[…]

### 2022-12-23 scripts/figures/fig_7.R

@@ -1,12 +1,12 @@
 # HEADER ----------------------------------------------------------------------
 #
 # Title:        Create figure 7
-# Description:  The objective of this routine is compare the transition length
+# Description:  The objective of this routine is compare the conversion length
 #               results, comparing them with visual observations.
 #
 #
 # Author:       Hugo Tameirao Seixas
-# Contact:      tameirao.hugo@gmail.com
+# Contact:      seixas.hugo@protonmail.com
 # Date:         2022-08-03
 #
 # Notes:
@@ -38,25 +38,25 @@ library(ggtext)
 
 ## Load observed data ----
 observed <- read_csv("data/validation/results.csv") %>%
-  mutate( # Remove points where there were no transition
+  mutate( # Remove points where there were no conversion
     forest_year = if_else(forest_year <= 0, NA_real_, forest_year),
     agri_year = if_else(agri_year <= 0, NA_real_, agri_year),
-    trans_length = if_else(agri_year <= 0, NA_real_, trans_length),
-    trans_length = if_else(forest_year <= 0, NA_real_, trans_length)
+    c_length = if_else(agri_year <= 0, NA_real_, c_length),
+    c_length = if_else(forest_year <= 0, NA_real_, c_length)
   ) %>%
   mutate( # Fix dates to match the estimates
     forest_year = forest_year,
     agri_year = agri_year
   ) %>%
   pivot_longer(
-    cols =  c(forest_year, agri_year, trans_length),
+    cols =  c(forest_year, agri_year, c_length),
     names_to = "var",
     values_to = "observed"
   ) %>%
   arrange(cell_id)
 
 ## Load estimates data ----
-estimated <- open_dataset("data/trans_tabular_dataset/trans_length/")
+estimated <- open_dataset("data/c_tabular_dataset/c_length/")
 
 estimated <- estimated %>%
   filter(
@@ -64,7 +64,7 @@ estimated <- estimated %>%
   ) %>%
   collect() %>%
   pivot_longer(
-    cols =  c(forest_year, agri_year, trans_length),
+    cols =  c(forest_year, agri_year, c_length),
     names_to = "var",
     values_to = "estimated"
   ) %>%
@@ -80,9 +80,9 @@ comparison <- estimated %>%
     difference = observed - estimated,
     var = factor(
       var,
-      levels = c("forest_year", "agri_year", "trans_length"),
+      levels = c("forest_year", "agri_year", "c_length"),
[…]

### 2022-12-23 scripts/figures/fig_8.R

@@ -1,12 +1,12 @@
 # HEADER ----------------------------------------------------------------------
 #
 # Title:        Create figure 8
-# Description:  The objective of this routine is compare the transition length
+# Description:  The objective of this routine is compare the conversion length
 #               results, comparing them with visual observations.
 #
 #
 # Author:       Hugo Tameirao Seixas
-# Contact:      tameirao.hugo@gmail.com
+# Contact:      seixas.hugo@protonmail.com
 # Date:         2022-08-03
 #
 # Notes:
@@ -38,22 +38,22 @@ library(ggtext)
 
 ## Load observed data ----
 observed <- read_csv("data/validation/results.csv") %>%
-  mutate( # Remove points where there were no transition
+  mutate( # Remove points where there were no conversion
     forest_year = if_else(forest_year <= 0, NA_real_, forest_year),
     agri_year = if_else(agri_year <= 0, NA_real_, agri_year),
-    trans_length = if_else(agri_year <= 0, NA_real_, trans_length),
-    trans_length = if_else(forest_year <= 0, NA_real_, trans_length)
+    c_length = if_else(agri_year <= 0, NA_real_, c_length),
+    c_length = if_else(forest_year <= 0, NA_real_, c_length)
   ) %>%
   mutate(across(forest_year:agri_year, ~ .x - 1984)) %>%
   pivot_longer(
-    cols =  c(forest_year, agri_year, trans_length),
+    cols =  c(forest_year, agri_year, c_length),
     names_to = "var",
     values_to = "observed"
   ) %>%
   arrange(cell_id)
 
 ## Load estimates data ----
-estimated <- open_dataset("data/trans_tabular_dataset/trans_length/")
+estimated <- open_dataset("data/c_tabular_dataset/c_length/")
 
 estimated %<>%
   filter(
@@ -62,7 +62,7 @@ estimated %<>%
   collect() %>%
   mutate(across(forest_year:agri_year, ~ .x - 1984)) %>%
   pivot_longer(
-    cols =  c(forest_year, agri_year, trans_length),
+    cols =  c(forest_year, agri_year, c_length),
     names_to = "var",
     values_to = "estimated"
   ) %>%
@@ -78,9 +78,9 @@ comparison <- estimated %>%
     difference = observed - estimated,
     var = factor(
       var,
-      levels = c("forest_year", "agri_year", "trans_length"),
+      levels = c("forest_year", "agri_year", "c_length"),
       labels = c(
[…]

### 2022-12-23 scripts/03_extract_conversion.R
Remove deprecated function.

@@ -31,7 +31,6 @@ library(dplyr)
 # OPTIONS ---------------------------------------------------------------------
 #
 source("conf/config.R")
-source("R/terra_as_tibble.R")
 #
 # LOAD AUXILIARY DATA ---------------------------------------------------------
 

### 2022-12-26 scripts/figures/fig_4.R
Update output tables names.

@@ -118,13 +118,13 @@ frequency <- c_length %>%
 
 ## Save plot data ----
 # Hex grid data
-write_csv(c_hex, "data/figures/fig_1_hex.csv")
+write_csv(c_hex, "data/figures/fig_4_hex.csv")
 
 # Conversion length frequency
-write_csv(frequency, "data/figures/fig_1_freq.csv")
+write_csv(frequency, "data/figures/fig_4_freq.csv")
 
 # High resolution raster values
-write_csv(hr_raster, "data/figures/fig_1_hr.csv")
+write_csv(hr_raster, "data/figures/fig_4_hr.csv")
 
 ## Disconnect from Spark after all queries ----
 spark_disconnect(sc)

### 2022-12-26 scripts/figures/fig_5.R

@@ -103,7 +103,7 @@ c_subset %<>%
 ### Save plot data ----
 write_csv(
   c_subset,
-  "data/figures/fig_2.csv"
+  "data/figures/fig_5.csv"
 )
 
 ## Disconnect from Spark after all queries ----

### 2022-12-26 scripts/figures/fig_6.R

@@ -38,7 +38,9 @@ source('conf/spark_config.R')
 # LOAD AUXILIARY DATA ---------------------------------------------------------
 
 ## Load municipality data ----
-municip <- read_municipality(year = "2019")
+municip <- read_municipality(year = "2019") %>%
+  as_tibble() %>%
+  select(code_muni, name_state)
 
 # LOAD DATASETS ---------------------------------------------------------------
 
@@ -70,16 +72,24 @@ c_ridges <-
   c_length %>%
   filter(agri_cycle == 1) %>%
   select(cell_id:forest_type) %>%
-  sdf_sample(fraction = 0.2, replacement = FALSE) %>%
+  sdf_sample(fraction = 0.01, replacement = FALSE, seed = 1) %>%
   left_join(
     mask_cells %>% select(cell_id, area, code_muni),
     by = "cell_id"
   ) %>%
+  select(agri_year, c_length, code_muni) %>%
   collect() %>%
   left_join(municip, by = "code_muni") %>%
   drop_na() %>%
   mutate(agri_year = ymd(agri_year, truncated = 2L)) %>%
-  filter(year(agri_year) >= 1995)
+  filter(year(agri_year) >= 1995) %>%
+  select(-code_muni)
+
+### Save plot data ----
+write_csv(
+  c_ridges,
+  "data/figures/fig_6.csv"
+)
 
 ## Disconnect from Spark after all queries ----
 spark_disconnect(sc)

### 2022-12-26 scripts/figures/fig_7.R

@@ -110,10 +110,10 @@ metrics <- mae %>%
 ## Save plots data ----
 
 # Comparison table
-write_csv(comparison, "data/figures/fig_4_5.csv")
+write_csv(comparison, "data/figures/fig_7_8.csv")
 
 # Error metrics
-write_csv(metrics, "data/figures/fig_4_metrics.csv")
+write_csv(metrics, "data/figures/fig_7_metrics.csv")
 
 # ANALYSE RESULTS -------------------------------------------------------------
 

### 2022-12-26 scripts/figures/fig_8.R

@@ -112,10 +112,10 @@ quants <- do.call("rbind", replicate(38, quants, simplify = FALSE)) %>%
 ## Save plots data ----
 
 # Comparison table
-write_csv(comparison, "data/figures/fig_4_5.csv")
+write_csv(comparison, "data/figures/fig_7_8.csv")
 
 # Quantiles
-write_csv(quants, "data/figures/fig_5_quants.csv")
+write_csv(quants, "data/figures/fig_8_quants.csv")
 
 # ANALYSE RESULTS -------------------------------------------------------------
 

### 2022-12-26 manuscript/manuscript.tex
Update figures.

@@
 \begin{figure}[h]
-\includegraphics[width=17cm]{figs/trans_length_cols} \caption{Conversion area from forest to agriculture, per year and conversion length.
+\includegraphics[width=17cm]{figs/c_length_cols} \caption{Conversion area from forest to agriculture, per year and conversion length.
 The bars represent the total amount of area at some state of the conversion for each year.
@@
 \begin{figure}[h]
-\includegraphics[width=17cm]{figs/trans_ridge} \caption{Conversion length patterns inside states in the Amazon biome.
+\includegraphics[width=17cm]{figs/c_ridge} \caption{Conversion length patterns inside states in the Amazon biome.
 Each year has a density estimate of the conversion lengths, represented as colored curves.
